---
title: "Sea Ice Edge"
author: "Melanie Frazier (UCSB, NCEAS, OHI)"
date: "June 22, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---
```{r download data}

library(rgdal)
library(raster)
library(rasterVis)

source("https://raw.githubusercontent.com/OHI-Science/ohiprep_v2018/master/src/R/spatial_common.R")

```
Goals:

1. Get a sea ice raster that reflects the location of the sea ice boundary
2. Create a permanent sea ice raster layer

### Sea ice edge
The habitat layer indicates the location of sea ice edge because this is the critical habitat.

We also create a sea ice layer to add as a mask to indicate regions with high uncertainty in scores.


### Data
We use: Sea Ice Concentrations from Nimbus-7 SMMR and DMSP SSM/I-SSMIS Passive Microwave Data

Downloaded from: http://nsidc.org/data/NSIDC-0051 (July 3 2018)

### Notes about data

Data Value   | Description
------------- | ---------------------
0-250       | Sea ice concentration (fractional coverage scaled by 250)
251         | Circular mask used in the Arctic to cover the irregularly-shaped data gap around the pole (caused by the orbit inclination and instrument swath)
252         | Unused
253         | Coastlines
254         | Superimposed land mask
255         | Missing data

#### Methods
Habitat is based on sea ice data from 1981-2010, which is used by the NOAA Sea Ice Index.  

This is a good compromise between where sea ice currently is and it's location in the recent past.  

We classify monthly data with 15-80% sea ice concentrations as sea ice.  

We then determine the proportion of time sea ice edge is present across all years and months.  

#### Download, format, and summarize rasters
The following portion of the script:

1. Downloads monthly sea ice concentration data from ftp site for years 1981-2010 for north and south poles and saves as a raster stack.
2. Converts 250 values to sea ice concentration by dividing by 250. Values >250 indicate other features, such as shoreline or land, and these were converted to NA.
2. Classifies 15-80% ice concentration as sea ice edge (raster value 1)
3. Average of all month/year rasters to determine the proportion of time a raster is sea ice edge.
4. Also creates a "hole" raster layer (hole has value of 1) and land raster layer (land and shoreline regions are classified as 1).

```{r download data}

pixel = 25000 # pixel dimension in meters for both x and y
prj.n = "+proj=stere +lat_0=90 +lat_ts=70 +lon_0=-45 +k=1 +x_0=0 +y_0=0 +a=6378273 +b=6356889.449 +units=m +no_defs"
prj.s = "+proj=stere +lat_0=-90 +lat_ts=-70 +lon_0=0 +k=1 +x_0=0 +y_0=0 +a=6378273 +b=6356889.449 +units=m +no_defs"
prj.mol = "+proj=moll +lon_0=0 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"

# URL base (ub), filename format for final monthly data is nt_YYYYMM_SSS_vVV_R.bin
ub.n = "ftp://sidads.colorado.edu/pub/DATASETS/nsidc0051_gsfc_nasateam_seaice/final-gsfc/north/monthly"
ub.s = "ftp://sidads.colorado.edu/pub/DATASETS/nsidc0051_gsfc_nasateam_seaice/final-gsfc/south/monthly"

poles = c("n","s")
years = c(1981:2010) 
months = 1:12
n.pym = length(poles)*length(years)*length(months)
i.pym = 0  # starting count


t0 = Sys.time()

for(p in poles){
#p="s" # testing
missing_data_count <- NA   
  ######################################################################################################################
  ## Create an empty raster stack with appropriate dimensions and CRS
  ######################################################################################################################
  

  ## extents from NSIDC (http://nsidc.org/data/polar_stereo/ps_grids.html
  
  if (p == "n"){
    xMin = -3850000; yMin = -5350000; nr = 448; nc = 304; prj = prj.n; ub = ub.n
  } else if (p == "s"){
    xMin = -3950000; yMin = -3950000; nr = 332; nc = 316; prj = prj.s; ub = ub.s
  }
  xMax = xMin + (pixel*nc); yMax = yMin + (pixel*nr)
  
  r <- raster(nrow = nr, ncol = nc, xmn = xMin, xmx = xMax, ymn = yMin, ymx = yMax)
  projection(r) <- prj
  s <- stack(r)

  ######################################################################################################################
  ## Download data
  ######################################################################################################################
  
  for (yr in years){
    for (mo in months){ 
      
      # yr=1981; mo=1 # testing
      
      ## get proper ftp (file transfer protocol) site based on time of data collection    
      i.pym <- i.pym + 1 
      ym <- yr*100 + mo
      y.m <- sprintf("%d-%02d", yr, mo)
      p.y.m <- sprintf("%s%d%02d", p, yr, mo)
      
      if (ym < 198709){
        ss = "n07"
      } else if (ym >= 198709 & ym < 199201){
        ss = "f08"
      } else if (ym >= 199201 & ym < 199510){
        ss = "f11"
      } else if (ym >= 199510 & ym < 200801){
        ss = "f13"
      } else if (ym >= 200801){
        ss = "f17"
      }
      
      ## retrieving the data using ftp
      min.done <- as.numeric(difftime(Sys.time(), t0, units="mins"))
      min.togo <- (n.pym - i.pym) * min.done/i.pym
      print(sprintf("Retrieving %s (%d of %d). Minutes done=%0.1f, to go=%0.1f",
                    p.y.m,i.pym,n.pym,min.done,min.togo)) # time remaining for data download
      
      u <- sprintf("%s/nt_%d_%s_v1.1_%s.bin", ub, ym, ss, p)
      con <- file(u, "rb")  # "rb" = "open for reading in binary mode"
      x <- readBin(con, "raw", 300)
      x <- readBin(con,"int", size = 1, signed = FALSE, 150000)
      close(con)    
      
      ## place result in raster framework
      edge <- setValues(r, x)
      
      ## count of missing data (need to do something about these if they are present)
      missing <- sum(edge[edge==255])
      missing_data_count <- c(missing_data_count, missing)
      
      ## get a land/shoreline raster
      land_shoreline <- edge
      land_shoreline[land_shoreline < 253] <- NA
      land_shoreline[land_shoreline %in% c(253, 254)] <- 1
      
      ## get a land/shoreline raster
      hole <- edge
      hole[hole < 251] <- NA
      hole[hole > 251] <- NA
      hole[hole ==251] <- 1
      
      # these are rasters that are not ice
      edge[edge>250] <- NA
      edge <- edge/250

      edge[(edge<0.15 | edge > 0.8) ] <- 0 
      edge[edge>0] <- 1
        

       ## add raster data (r) to the stack (s) and name the layer: pole.year.month (e.g. s197901)
    s.names <- names(s)
      
      if (nlayers(s) == 0){ 
        s = stack(edge)
        names(s) = p.y.m
        
        hole_s = stack(hole)
        names(hole_s) = p.y.m
        
        land_shoreline_s = stack(land_shoreline)
        names(land_shoreline_s) = p.y.m
      } else {
        s = stack(s, edge)
        names(s) = c(s.names, p.y.m)   
        
        hole_s = stack(hole_s, hole)
        names(hole_s) = c(s.names, p.y.m)   

        land_shoreline_s = stack(land_shoreline_s, land_shoreline)
        names(land_shoreline_s) = c(s.names, p.y.m)
      }
    }
  }
  
  ## visualize
  # animate(s)
  # library(rasterVis)
  # rasterVis::levelplot(s[[361:372]], col.regions=c('gray', "red"), colorkey=FALSE)
    
if(p=="n"){the_pole = "north"} else
{the_pole = "south"}
    
  n.mean <- mean(s)
  #plot(n.mean)
  writeRaster(n.mean, sprintf("habitats/seaice/seaice_%s_mean_sea_ice_edge.tif", the_pole), overwrite=TRUE)

  hole.mean <- mean(hole_s, na.rm=TRUE)
  #plot(hole.mean)
  writeRaster(hole.mean, sprintf("habitats/seaice/seaice_%s_hole.tif", the_pole), overwrite=TRUE)
  
  land.mean <- mean(land_shoreline_s, na.rm=TRUE)
  #plot(land.mean)
  writeRaster(land.mean, sprintf("habitats/seaice/seaice_%s_land.tif", the_pole), overwrite=TRUE)
  
  cat(sprintf("missing data count for %s pole", the_pole))
  table(missing_data_count)
  }
  
```


#### Gapfill North Pole hole
We make the hole 0 sea ice edge (this isn't exactly correct, but it is close enough).


```{r hole}

north <- raster("habitats/seaice/seaice_north_mean_sea_ice_edge.tif")
land <- raster("habitats/seaice/seaice_north_land.tif")


north[is.na(north)] <- 0 
north_gf_hole <- raster::mask(north_gf_hole, land, inverse=TRUE)
writeRaster(north_gf_hole, "habitats/seaice/seaice_north_mean_edge_gapfill_hole.tif")

```

#### Combine North and South pole data

```{r north and south}

raster("habitats/seaice/seaice_north_mean_edge_gapfill_hole.tif") %>%
  projectRaster(ocean, over=TRUE, method="ngb",
                filename = file.path(dir_M, "git-annex/impact_acceleration/seaice/north_sea_ice_edge.tif"),
                progress="text", overwrite=TRUE)


raster("habitats/seaice/seaice_south_mean_sea_ice_edge.tif") %>%
  projectRaster(ocean, over=TRUE, method="ngb",
                filename = file.path(dir_M, "git-annex/impact_acceleration/seaice/south_sea_ice_edge.tif"),
                progress="text", overwrite=TRUE)

n_rast <- raster(file.path(dir_M, "git-annex/impact_acceleration/seaice/north_sea_ice_edge.tif"))
s_rast <- raster(file.path(dir_M, "git-annex/impact_acceleration/seaice/south_sea_ice_edge.tif"))


n_and_s <- merge(n_rast, s_rast, progress="text",
                 filename=file.path(dir_M, "git-annex/impact_acceleration/seaice/north_south_sea_ice_edge.tif"), overwrite=TRUE)


combined <- raster(file.path(dir_M, "git-annex/impact_acceleration/seaice/north_south_sea_ice_edge.tif"))

plot(combined)

```


#### Gapfill to coast

```{r coast_gf}


combined <- raster(file.path(dir_M, "git-annex/impact_acceleration/seaice/north_south_sea_ice_edge.tif"))

plot(combined)

```
