---
title: "Annual Artisanal Fisheries Catch"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '../../../ohiprep_v2018/src/templates/ohi_hdr.html'
pdf_document:
  toc: true
---

This script combines spatial catch data from [Watson (2017)](https://www.nature.com/articles/sdata201739) with species & gear specific categories defined in [`watson_gear_matching`](https://github.com/OHI-Science/impact_acceleration/blob/master/stressors/comm_fish/watson_gear_matching.Rmd) to create annual catch rasters for artisanal fishing.

## Overview
- Raw catch data is combined with `taxa_gear_types.csv` as defined in `watson_gear_matching.Rmd`
- Data is filtered for every year and catch type, rasterized to half degree cells, and saved as a GeoTIFF

```{r setup, message=F,warning=F}

knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',message = FALSE, warning = FALSE)


library(seaaroundus)
library(raster)
library(readr)
library(dplyr)
library(data.table)
library(foreach)
library(doParallel)
library(RColorBrewer)
library(stringr)

#setwd("stressors/art_fish")

source("https://raw.githubusercontent.com/OHI-Science/ohiprep_v2018/master/src/R/spatial_common.R")

cols = rev(colorRampPalette(brewer.pal(11, 'Spectral'))(255)) # rainbow color scheme

# parallel processing
cl<-makeCluster(5)
registerDoParallel(cl)

```

#Load Data

Load annual catch data

```{r}
annual_files <- list.files(file.path(dir_M, 'git-annex/impact_acceleration/stressors/comm_fish_v3.0/int/catch_annual_data'), full.names=TRUE)
```

#Map annual catch

## Get base raster

Using the [searoundus R package](https://github.com/ropensci/seaaroundus), we create a raster with cell values set equal to cell IDs in the data, which are defined in the column `Seq`.

```{r}

  #get template raster for SAUP data with cell values equal to CellID
  
   saup_cells <- getcells("POLYGON ((-180 90, 180 90, 180 -90, -180 -90, -180 90))")

   saup_rast <- raster(ncol=720, nrow=360)
   saup_rast[] <- saup_cells
   
   plot(saup_rast, col=cols, main = "SAUP Cell IDs")

```

## Create a buffer to gapfill shoreline NA values

## Create a buffer to gapfill shoreline NA values

The ocean mask is used to create a buffer to identify shoreline values that will be gapfilled, but first our ocean raster needs to be resampled and reprojected to have the same resolution and CRS as the fisheries data.  

The following creates an ocean buffer (1 cell inland and 2 cells offshore) at resolution that matches fisheries data.  Cells that fall in this buffer and have no recorded catch records are gapfilled, these are the cells that were cut due to mismatches in resolution and ocean boundaries.  The following saves a file indicating the location of gapfilled cells.
```{r ocean_low_res, eval=FALSE}
 
saup_rast_mol <- projectRaster(saup_rast, crs=crs(ocean), over=TRUE)

saup_rast_mol[saup_rast_mol>=0] <- 1

plot(saup_rast_mol)
 
ocean_low_res = resample(ocean, saup_rast_mol, method="ngb")

# this creates a buffer that extends one raster cell inland and 2 raster cells offshore
buffer1 <- boundaries(ocean_low_res, type = "inner", asNA = TRUE, progress = "text") 
buffer2 <- boundaries(buffer1, type = "outer", asNA = FALSE, progress = "text") 
buffer2[buffer2==0] <- 1 
coastal_buffer <- subs(buffer2, data.frame(id=c(NA,1), v=c(0,NA)))
plot(coastal_buffer)
#zoom(coastal_buffer)


file <-  annual_files[15]

data <- readRDS(file) %>%
        rowwise() %>%
        group_by(Cell) %>%  #group by cell number
        summarise(cell_catch = sum(TotalCatch)) %>%
        data.frame()#summing up all types of catch rates (units: tonnes)

  #rasterize catch by swapping cell ids with 

catch_gf <- raster::subs(saup_rast, data, by = 1, which = 2) %>%
              projectRaster(crs=crs(ocean), method="ngb", over = TRUE)

catch_gf[is.na(catch_gf)] <- 9999 
plot(catch_gf)

# limit gapfilling to shoreline
coastal_catch_gf <- overlay(catch_gf, coastal_buffer, fun = function(x,y){ifelse(x %in% 9999 & is.na(y), 9999, NA)})

## save gf raster
catch_gf_mol <- resample(coastal_catch_gf, ocean, method = "ngb")

catch_gf_mol_mask <- mask(catch_gf_mol, ocean)  

writeRaster(catch_gf_mol_mask, file.path(dir_M,'/git-annex/impact_acceleration/stressors/art_fish_v3.0/int/art_fish_gf.tif'), overwrite=TRUE)
```

#### Create a function to gapfill missing catch data located along the coastline

Cells that should be gapfilled are identified and then estimated using the average of nearest neighbors. 
```{r, eval=FALSE}
fish_gf <- function(catch_path, file_name, yr){

rast <-  raster(file.path(dir_M, catch_path, paste0(file_name, "_", yr, ".tif"))) 
#plot(rast)

# convert NA values to 0
rast[is.na(rast)] <- 0

# for cells in the coastal buffer with no catch data, convert from zero to NA values
fis_gf <- overlay(rast, coastal_catch_gf, fun = function(x,y){ifelse(y %in% 9999, NA, x)})
#plot(fis_gf)

# now make land NA (don't want to gapfill using land 0 values)
fis_gf <- overlay(fis_gf, ocean_low_res, fun = function(x,y){ifelse(x %in% 0 & is.na(y), NA, x)})
# plot(fis_gf)
# fis_gf

## Function to gapfill NA cells using mean of surrounding cells
gf_raster <- function(x){focal(x, w = matrix(1,3,3), fun = mean, na.rm=TRUE, pad = TRUE, NAonly=TRUE)}

## Iterative gapfilling for each year of NPP data
## Repeat until critical NA cells are filled
i <- 0

while (i <= 50){
fis_gf <- gf_raster(fis_gf)
i <- i + 1
#print(i)
}

#plot(fis_gf)
writeRaster(fis_gf, file.path(dir_M, catch_path, paste0(file_name, "_", yr, "_gf", ".tif")), overwrite=TRUE)

}

```

## Map annual artisanal catch

Each annual stressor layer is created by aggregate all artisanal fishing catch per cell per year. The catch is identified by the column `SSF_CR` (small scale fishing rate)

```{r map,eval=F}
foreach(file = annual_files, .packages="dplyr") %dopar% {  #file = annual_files[14]
  
  yr <- stringr::str_sub(file, -8, -5)
  
  data <- readRDS(file) %>%
    dplyr::filter(IsArtisanal) %>% 
    dplyr::group_by(Cell) %>%  #group by cell number
    dplyr::summarise(cell_catch = sum(TotalCatch)) %>%
    data.frame()#summing up all types of catch rates (remember these are in tonnes/km2 so we don't need to multiply by Ocean Area since we have 1km2 resolution)
  
  #rasterize catch by swapping cell ids with 
  raster::subs(saup_rast, data, by = 1, which = 2) %>%
    raster::projectRaster(crs=raster::crs(ocean), method = "ngb", over = TRUE,
                     filename = file.path(dir_M, sprintf('git-annex/impact_acceleration/stressors/art_fish_v3.0/int/catch_annual_rasters/art_fishing_%s.tif', yr)), overwrite=TRUE)
#tmp <- raster(file.path(dir_M, 'git-annex/impact_acceleration/stressors/art_fish/int/catch_annual_rasters/art_fishing_2014.tif'))
#plot(tmp)
}

## apply gapfill function to each of the above rasters
for(year_catch in 2000:2014){ # year_catch = 2014
fish_gf(catch_path="git-annex/impact_acceleration/stressors/art_fish_v3.0/int/catch_annual_rasters",
        file_name = "art_fishing",
        yr = year_catch)
}

# old <- raster(file.path(dir_M, 'git-annex/impact_acceleration/stressors/art_fish/int/catch_annual_rasters/art_fishing_2014.tif'))
# 
# plot(log(old+1))
# 
# new <- raster(file.path(dir_M, 'git-annex/impact_acceleration/stressors/art_fish_v3.0/int/catch_annual_rasters/art_fishing_2014.tif'))
# 
# plot(log(new+1))
```


