---
title: "Creating artisanal fishing pressure layers"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '../../../ohiprep_v2018/src/templates/ohi_hdr.html'
pdf_document:
  toc: true
---

This script standardizes the annual catch rasters by net primary production.

```{r setup, message=F,warning=F}

knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',message = FALSE, warning = FALSE)

library(seaaroundus)
library(raster)
library(readr)
library(stringr)
library(dplyr)
library(data.table)
library(parallel)
library(foreach)
library(doParallel)
library(RColorBrewer)

source("https://raw.githubusercontent.com/OHI-Science/ohiprep_v2018/master/src/R/spatial_common.R")

years_of_data <- 2003:2014

cols = rev(colorRampPalette(brewer.pal(11, 'Spectral'))(255)) # rainbow color scheme

# setwd("stressors/art_fish")

# paralell processing
cl<-makeCluster(10)
registerDoParallel(cl)

```


# Primary production data

Load NPP data

```{r,eval=F}

npp_files_mol <- list.files(file.path(dir_M,'git-annex/impact_acceleration/stressors/comm_fish/int/npp_annual'), full.names=TRUE, pattern = "gf")

npp_files_mol <- npp_files_mol[!(str_detect(npp_files_mol, "npp_gf"))]

npp_files_mol <- npp_files_mol[grep(paste(years_of_data, collapse="|"), npp_files_mol)]
```

# Standardize catch

`catch_npp_fun` is a function that reads in a catch raster, reprojects & resamples it then divides by npp for the same year and saves the output.

```{r, eval=F}

#grab years 2003 - 2014
years_filter <-  paste(years_of_data, collapse="|")

art_catch_rasts <- list.files(file.path(dir_M,'git-annex/impact_acceleration/stressors/art_fish_v3.0/int/catch_annual_rasters'), full.names=TRUE, pattern = "gf")
art_catch_rasts <- art_catch_rasts[grep(years_filter, art_catch_rasts)]


for(file in art_catch_rasts) { # file = art_catch_rasts[1]
  
  catch <- raster(file)
  
  yr <- substr(file, nchar(file)-10, nchar(file)-7)
  
  npp <- npp_files_mol[str_detect(npp_files_mol, yr)] %>% 
    raster()
  
  c <- catch %>%
        resample(npp, method = 'ngb') %>%
       overlay(., npp, fun=function(x,y){x/y})

writeRaster(c, filename = file.path(dir_M, sprintf('git-annex/impact_acceleration/stressors/art_fish_v3.0/int/catch_annual_npp_rasters/art_fish_%s.tif', yr)), overwrite=TRUE) 
   
}

```


# Rescale

Create an ocean raster with same scale as the fishery/npp rasters to cut land values.
```{r, eval=FALSE}

catch_npp <- list.files(file.path(dir_M, 'git-annex/impact_acceleration/stressors/art_fish_v3.0/int/catch_annual_npp_rasters'), full.names = TRUE)
catch_npp <- raster(catch_npp[1])

ocean_low_res <- ocean %>%
  resample(catch_npp, method="ngb")

```


```{r,eval=F}

#list of lists
all_files <- list.files(file.path(dir_M, 'git-annex/impact_acceleration/stressors/art_fish_v3.0/int/catch_annual_npp_rasters'), full.names=TRUE)

values = stack(all_files) %>%
      mask(ocean_low_res) %>% 
      getValues()

values_vector <- as.vector(as.matrix(values))
values_vector <- na.omit(values_vector)

quant <-  quantile(values_vector, prob=0.9999, na.rm=TRUE)

tmp <- data.frame(cat="art_fishing", quant_9999=quant)
write.csv(tmp, file.path(dir_M, 'git-annex/impact_acceleration/stressors/art_fish_v3.0/int/quantile_9999.csv'))

```


Rescale the catch/npp data by the reference point and save.
```{r rescale, eval=FALSE}

quant <- read.csv(file.path(dir_M, 'git-annex/impact_acceleration/stressors/art_fish_v3.0/int/quantile_9999.csv'))[["quant_9999"]]

#for(file in all_files[11:12]){
foreach(file = all_files,.packages="dplyr") %dopar% {    # file = all_files[1]
  
    yr = substr(file, nchar(file)-7, nchar(file)-4)
    
    
     raster::raster(file) %>%
         raster::calc(fun=function(x){ifelse(x<0, 0,
                                   ifelse(x>quant, 1, x/quant))}) %>%
          raster::resample(ocean, method = 'ngb') %>%
            raster::mask(ocean,
            filename = file.path(dir_M, sprintf('git-annex/impact_acceleration/stressors/art_fish_v3.0/final/art_fish_%s_rescaled_mol.tif', yr)),                       
            overwrite=TRUE)  

# check:
# tmp <- raster(file.path(dir_M, "git-annex/impact_acceleration/stressors/art_fish/final/art_fish_2003_rescaled_mol.tif"))
# plot(tmp)
}

```

# Check Results

```{r}

results <- list.files(file.path(dir_M,'git-annex/impact_acceleration/stressors/art_fish_v3.0/final'), full.names = TRUE)[12]
tmp <- raster::raster(results)

plot(tmp, col=cols, axes=FALSE)

```