---
title: "Calc_impacts"
output: html_document
editor_options: 
  chunk_output_type: console
---

Caculate impact for each habitat by summing the stressor x habitat x vulnerability rasters for each habitat.

## Loading packages

```{r}
#libraries
library(raster)
library(RColorBrewer)
library(sf)

library(stringr)
library(dplyr)

library(doParallel)
library(foreach)
library(parallel)

source("https://raw.githubusercontent.com/OHI-Science/ohiprep_v2018/master/src/R/spatial_common.R")


```


## List of combos
```{r}


years <- 2003:2013

years_subset <- paste(years, collapse="|")

pressures_to_use <- list.files(file.path(dir_M, "git-annex/impact_acceleration/hab_stressor_combo"))
complete_yr <- table(str_sub(pressures_to_use, 1, -11))
complete_yr <- complete_yr[complete_yr == max(complete_yr)]
complete_yr <- paste(names(complete_yr), collapse="|")

combos <- list.files(file.path(dir_M, "git-annex/impact_acceleration/hab_stressor_combo"), full=TRUE)
combos <- grep(complete_yr, combos, value=TRUE)
combos <- grep(years_subset, combos, value=TRUE)

```


## Create a list of habitat/year combinations

```{r}
years <- 2003:2013

## Get a list of the habitat/year combinations:
hab_files <- list.files(file.path(dir_M, "git-annex/impact_acceleration/habitats"))
hab_files <- hab_files[-(grep(".vat.dbf|.xml|.ovr", hab_files))]
hab_files <- gsub(".tif", "", hab_files)
  
habs <- expand.grid(hab = basename(hab_files), year=years)
  
```

## Check stressor data
All stressor/years should have 21 rasters, one for each habitat.

```{r}

check_hab_count <- data.frame(N=rep(NA, dim(habs)[1]), pressure_year=rep(NA, dim(habs)[1]))

for(row in 1:dim(habs)[1]){ # row=1

  year_n <- habs$year[row]
  hab_n <- habs$hab[row]
  
  tmp <- grep(year_n, combos, value=TRUE)
  tmp <- grep(hab_n, tmp, value =TRUE)
  
  check_hab_count$N[row] = length(tmp)
  check_hab_count$pressure_year[row] = paste(hab_n, year_n, sep="_")

}

check_hab_count
summary(check_hab_count)

```

## Create raster that sums all stressors x vulnerabilty for a habitat/year
For each habitat and year, identify the stressor x habitat x vulnerability rasters.  Sum these to get the cumulative impact on the habitats.  

```{r}

registerDoParallel(5)

foreach(row = 1:dim(habs)[1]) %dopar%{ # row=230

  year_n <- habs$year[row]
  hab_n <- habs$hab[row]
  
  tmp <- grep(year_n, combos, value=TRUE)
  tmp <- grep(hab_n, tmp, value =TRUE)
  
  hab_stack <- raster::stack(tmp)
  
  raster::calc(hab_stack, fun=sum, na.rm=TRUE,
      filename=file.path(dir_M, sprintf("git-annex/impact_acceleration/habitat_impact/hab_impact/%s_%s.tif", hab_n, year_n)), 
      overwrite=TRUE)

}

plot_tmp <- raster(file.path(dir_M, "git-annex/impact_acceleration/habitat_impact/hab_impact/cum_impact_surface_waters_2013.tif"))
plot(plot_tmp)

```

## Calculate trend for each habitat

```{r}

# list of pressure x habitat combos

impacts <- c("sst", "oa", "slr", 
            "shipping",
            "nutrient", "organic", "direct_human", "light",
            "pel_hb", "pel_lb", "dem_dest", "dem_nondest_hb", "dem_nondest_lb", "art_fish")

## Get a list of the habitat/year combinations:
hab_files <- list.files(file.path(dir_M, "git-annex/impact_acceleration/habitats"))
hab_files <- hab_files[-(grep(".vat.dbf|.xml|.ovr", hab_files))]
hab_files <- gsub(".tif", "", hab_files)
  
hab_x_stress <- expand.grid(habs = hab_files, impacts=impacts)

combo_files   <- list.files(file.path(dir_M, "git-annex/impact_acceleration/hab_stressor_combo"), full=TRUE)


registerDoParallel(5)


foreach (row = 1:dim(hab_x_stress)[1]) %dopar%{ # row=1

  impacts_n <- hab_x_stress$impacts[row]
  hab_n <- hab_x_stress$hab[row]
  
  tmp <- grep(impacts_n, combo_files, value=TRUE)
  tmp <- grep(hab_n, tmp, value =TRUE)
  
  years <-   as.numeric(stringr::str_sub(basename(tmp), -8, -5))
  
  # add 1 for a model with an intercept
X <- cbind(1, years)
invXtX <- solve(t(X) %*% X) %*% t(X)
quickfun <- function(y) (invXtX %*% y)[2]

  
  hab_stack <- raster::stack(tmp)

  raster::calc(hab_stack, quickfun, filename = file.path(dir_M, sprintf("git-annex/impact_acceleration/habitat_impact/hab_trend/%s_%s_trend.tif", hab_n, impacts_n)),
       overwrite=TRUE)
  
}

plot(raster(file.path(dir_M, "git-annex/impact_acceleration/habitat_impact/hab_trend/beach_sst_trend.tif")))



```


## Ocean mask for each layer created above.
```{r}
rescaled_data <- list.files(file.path(dir_M, sprintf("git-annex/impact_acceleration/tmp")), full=TRUE)
rescaled_data <- rescaled_data[grep("rescaled", rescaled_data)]

for(rescale in rescaled_data) { # rescale=rescaled_data[1]
  
 rescaled <- raster::raster(rescale)

 # get stressor and year from file name
stressor <- basename(rescale)
 stressor <- gsub("_rescaled.tif", "", stressor)
 stressor <- gsub("summed_raster_", "", stressor)
 stressor <- data.frame(combo=stressor)
 
 combo_data <- stressor %>% 
   tidyr::separate(combo, c("stressor", "year"), sep=-4) %>%
   dplyr::mutate(stressor = substr(stressor, 1, nchar(stressor)-1))
 
 # mask data and save
 raster::mask(rescaled, ocean,
                 filename=file.path(dir_M, sprintf("git-annex/impact_acceleration/impact/stressor_impact/%s_%s.tif", 
                                                   combo_data$stressor, combo_data$year)),
                 overwrite=TRUE, progress="text")
 
}
