---
title: "Calc_impacts"
output: html_document
---

Caculate impact for each stressor by summing the stressor x habitat x vulnerability rasters for each stressor, and then dividing by the number of habitats.

## Loading packages

```{r}
#libraries
library(raster)
library(RColorBrewer)
library(sf)
library(dplyr)
library(doParallel)
library(foreach)
library(parallel)

source("https://raw.githubusercontent.com/OHI-Science/ohiprep_v2018/gh-pages/src/R/spatial_common.R")


```

## Habitat number raster
Create a raster that describes the number of habitats in each raster cell.

```{r}

habs <- list.files(file.path(dir_M, "git-annex/impact_acceleration/habitats"), full=TRUE)
habs <- habs[-(grep(".vat.dbf|.xml|.ovr", habs))]
hab_stack <- stack(habs)
calc(hab_stack, fun = sum, na.rm=TRUE, progress="text",
               filename=file.path(dir_M, "git-annex/impact_acceleration/habitat_number/habitat_num.tif"),
               overwrite=TRUE)

```

## Create a list of stressor/year combinations

```{r}
years <- 2003:2013

years_subset <- paste(years, collapse="|")

## Get a list of the stressor/year combinations:
stress_files <- list.files(file.path(dir_M, "git-annex/impact_acceleration/stressors"), recursive = TRUE, 
                           pattern = ".tif", full="TRUE")
stress_files <- grep("/final/", stress_files, value=TRUE)
stress_files <- grep("/archive/", stress_files, value=TRUE, invert=TRUE)

stress_files <- basename(stress_files)
stress_files <- gsub("_rescaled_mol.tif", "", stress_files)


#filter to relevant years
stress_files <- grep(years_subset, stress_files, value=TRUE)
sum(duplicated(stress_files)) # should be zero

stress_combos <- data.frame(files = stress_files) %>%
  tidyr::separate(files, c("stressor", "year"), sep=-4) %>%
  mutate(stressor=substr(stressor, 1, nchar(stressor)-1)) 

```

## Check stressor data
All stressor/years should have 21 rasters, one for each habitat.

```{r}

hab_num <- raster(file.path(dir_M, "git-annex/impact_acceleration/habitat_number/habitat_num.tif"))

combos <- list.files(file.path(dir_M, "git-annex/impact_acceleration/hab_stressor_combo"), full=TRUE)

check_hab_count <- data.frame(N=rep(NA, dim(stress_combos)[1]), pressure_year=rep(NA, dim(stress_combos)[1]))

for(row in 1:dim(stress_combos)[1]){ # row=1

  year <- stress_combos$year[row]
  stress <- stress_combos$stressor[row]
  
  tmp <- grep(year, combos, value=TRUE)
  tmp <- grep(stress, tmp, value =TRUE)
  
  check_hab_count$N[row] = length(tmp)
  check_hab_count$pressure_year[row] = paste(stress, year, sep="_")

}

check_hab_count
summary(check_hab_count)# should all have 21
table(check_hab_count$pressure_year) #should all have 1 record

```

## Create raster that sums all habitat x impacts for a stressor/year
For each stressor and year, identify the stressor x habitat x vulnerability rasters.  Sum these and then divide by the number of habitats.  

```{r}
hab_num <- raster(file.path(dir_M, "git-annex/impact_acceleration/habitat_number/habitat_num.tif"))

combos <- list.files(file.path(dir_M, "git-annex/impact_acceleration/hab_stressor_combo"), full=TRUE)

registerDoParallel(4)
foreach(row = 1:dim(stress_combos)[1]) %dopar%{ # row=1

  year <- stress_combos$year[row]
  stress <- stress_combos$stressor[row]
  
  tmp <- grep(year, combos, value=TRUE)
  tmp <- grep(stress, tmp, value =TRUE)
  
  stress_stack <- raster::stack(tmp)
  
  raster::calc(stress_stack, fun=sum, na.rm=TRUE,
      filename=file.path(dir_M, sprintf("git-annex/impact_acceleration/tmp/summed_raster_%s_%s.tif", stress, year)), 
      overwrite=TRUE)
  
  summed_rast <- raster::raster(file.path(dir_M, sprintf("git-annex/impact_acceleration/tmp/summed_raster_%s_%s.tif", stress, year)))
  
  raster::overlay(summed_rast, hab_num, fun=function(x,y){x/y}, 
                  filename = file.path(dir_M, sprintf("git-annex/impact_acceleration/tmp/summed_raster_%s_%s_rescaled.tif", stress, year)), overwrite=TRUE)

}

```

## Check
Were all the rasters created?

```{r}
# identify rasters that have been created:
combos_obs <- list.files(file.path(dir_M, sprintf("git-annex/impact_acceleration/tmp")))
combos_obs <- combos_obs[grep("_rescaled", combos_obs)]
combos_obs <- gsub("summed_raster_", "", combos_obs)
combos_obs <- gsub("_rescaled", "", combos_obs)

# compare them with the complete list
not_done <- setdiff(paste0(stress_combos$stressor, "_", stress_combos$year, ".tif"), combos_obs)
sort(not_done)
not_done <- gsub(".tif", "", not_done)
not_done <- data.frame(combo=not_done) %>%
  mutate(combo = as.character(combo))
not_done <- not_done %>%
  tidyr::separate(combo, c("stressor", "year"), sep=-4) %>%
  dplyr::mutate(stressor = substr(stressor, 1, nchar(stressor)-1))


```


## Ocean mask for each layer created above.
```{r}
rescaled_data <- list.files(file.path(dir_M, sprintf("git-annex/impact_acceleration/tmp")), full=TRUE)
rescaled_data <- rescaled_data[grep("rescaled", rescaled_data)]

for(rescale in rescaled_data) { # rescale=rescaled_data[1]
  
 rescaled <- raster::raster(rescale)

 # get stressor and year from file name
stressor <- basename(rescale)
 stressor <- gsub("_rescaled.tif", "", stressor)
 stressor <- gsub("summed_raster_", "", stressor)
 stressor <- data.frame(combo=stressor)
 
 combo_data <- stressor %>% 
   tidyr::separate(combo, c("stressor", "year"), sep=-4) %>%
   dplyr::mutate(stressor = substr(stressor, 1, nchar(stressor)-1))
 
 # mask data and save
 raster::mask(rescaled, ocean,
                 filename=file.path(dir_M, sprintf("git-annex/impact_acceleration/impact/stressor_impact/%s_%s.tif", 
                                                   combo_data$stressor, combo_data$year)),
                 overwrite=TRUE, progress="text")
 
}

```
