---
  title: 'EEZ (full and 3nm) data extraction'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
  code_folding: show
toc: true
toc_depth: 3
toc_float: yes
number_sections: false
theme: cerulean
highlight: haddock
includes: 
  in_header: '../../../ohiprep_v2018/src/templates/ohi_hdr.html'
pdf_document:
  toc: true
editor_options: 
  chunk_output_type: console
---

This script extracts chi and trend data for region EEZ boundaries (full eez and 3nm).

  
## set up
  
```{r setup, message=FALSE, warning=FALSE, verbose=FALSE}

library(dplyr)
library(tidyr)
library(googleVis)
library(raster)
library(ggplot2)
library(plotly)
library(htmlwidgets)
library(RColorBrewer)
library(ohicore)

source("https://raw.githubusercontent.com/OHI-Science/ohiprep_v2018/master/src/R/spatial_common.R")
```

## EEZ regions: CHI

```{r}

rgns_global <- rgns_global %>%
  filter(type_w_ant == "eez") %>%
  dplyr::select(rgn_id = rgn_ant_id, rgn_name)


chi <- list.files(file.path(dir_M, "git-annex/impact_acceleration/impact/cumulative_impact"), full=TRUE,
                  pattern="chi")


chi_stack <- stack(chi)

chi_data <- zonal(chi_stack, zones, fun="mean", progress="text", na.rm=TRUE)

chi_data_df <- data.frame(chi_data) %>%
  gather("pressure", "value", -1) %>%
  rename("rgn_id" = zone) %>%
  mutate(year = as.numeric(substring(pressure, 5, 8))) %>%
  mutate(pressure = "chi") %>%
  inner_join(rgns_global, by="rgn_id")

write.csv(chi_data_df, "impacts/zonal_data_eez/eez_chi.csv", row.names=FALSE)

## Motion plot to explore data
plot_data <- read.csv("impacts/zonal_data_eez/eez_chi.csv") %>%
  dplyr::select(rgn_name, year, value)
Motion=gvisMotionChart(plot_data, 
                       idvar="rgn_name", 
                       timevar="year")
plot(Motion)
print(Motion, file="impacts/zonal_data_eez/eez_chi.html")


```

## eez regions: chi trend
```{r}
trend <- raster(file.path(dir_M, "git-annex/impact_acceleration/impact/trend/chi_slope.tif"))

trend_data <- zonal(trend, zones, fun="mean", progress="text", na.rm=TRUE)

trend_data_df <- data.frame(trend_data) %>%
  rename("rgn_id" = zone, "value"=mean) %>%
  mutate(pressure = "chi_trend") %>%
  inner_join(rgns_global, by="rgn_id")

write.csv(trend_data_df, "impacts/zonal_data_eez/eez_chi_trend.csv", row.names=FALSE)


```

## eez regions: impact trend

```{r}
impact_trends <- list.files(file.path(dir_M, "git-annex/impact_acceleration/impact/trend/impacts"), full=TRUE)
impact_trends_stack <- stack(impact_trends)

trend_data <- zonal(impact_trends_stack, zones, fun="mean", progress="text", na.rm=TRUE)

trend_data_df <- data.frame(trend_data) %>%
  tidyr::gather("pressure", "value", -1) %>%
  dplyr::rename("rgn_id" = zone) %>%
  dplyr::mutate(pressure = stringr::str_sub(pressure, 1, stringr::str_length(pressure)-6)) %>%
  inner_join(rgns_global, by="rgn_id")


write.csv(trend_data_df, "impacts/zonal_data_eez/eez_impacts_trend.csv", row.names=FALSE)

```


## eez regions: impacts

```{r}

impact_files <- list.files(file.path(dir_M, "git-annex/impact_acceleration/impact/stressor_impact"), recursive = TRUE, 
                           pattern = ".tif", full="TRUE")

registerDoParallel(3)

foreach (year = 2003:2013) %dopar%{ # year = 2003
  
impact_year_files <- grep(year, impact_files, value=TRUE)

impact_stack <- raster::stack(impact_year_files)

impact_data <- raster::zonal(impact_stack, zones, fun="mean", progress="text", na.rm=TRUE)

impact_data_df <- data.frame(impact_data) %>%
  tidyr::gather("pressure", "value", -1) %>%
  dplyr::rename("rgn_id" = zone) %>%
  dplyr::mutate(year = as.numeric(stringr::str_sub(pressure, -4,-1))) %>%
  dplyr::mutate(pressure = stringr::str_sub(pressure, 1, stringr::str_length(pressure)-5)) %>%
  dplyr::inner_join(rgns_global, by="rgn_id")

write.csv(impact_data_df, sprintf("impacts/zonal_data_eez/eez_%s_impacts.csv", year), row.names=FALSE)

}

```

## eez regions: stressors
I haven't actually run this.  Not sure that I need this, but will keep code in case.

```{r}
years <- 2003:2013

years_subset <- paste(years, collapse="|")


####### Stressors
stress_files <- list.files(file.path(dir_M, "git-annex/impact_acceleration/stressors"), recursive = TRUE,
                           pattern = ".tif", full="TRUE")
stress_files <- grep("/final/", stress_files, value=TRUE)

#filter to relevant years
stress_files <- grep(years_subset, stress_files, value=TRUE)

stress_stack <- stack(stress_files)

hab_data <- raster::zonal(stress_stack, zones, fun="mean", progress="text", na.rm=TRUE)

  hab_data_df <- data.frame(hab_data) %>%
    tidyr::gather("pressure", "value", -1) %>%
    dplyr::rename("rgn_id" = zone) %>%
    dplyr::mutate(pressure = gsub("_rescaled_mol", "", pressure)) %>%
    dplyr::mutate(year = as.numeric(stringr::str_sub(pressure, -4,-1))) %>%
    dplyr::mutate(pressure = stringr::str_sub(pressure, 1, stringr::str_length(pressure)-5)) %>%
    inner_join(rgns_global, by="rgn_id")

  write.csv(hab_data_df, "impacts/zonal_data_eez/eez_pressure.csv", row.names=FALSE)

```


## 3nm eez regions: CHI (2003-2013)
I don't think we are using this view of the data.  But I am preserving this code in case we change our mind.

```{r}

rgns_3nm <- raster::raster(file.path(dir_M, "git-annex/globalprep/spatial/v2018/rgns_3nm_offshore_mol.tif"))
plot(rgns_3nm)

### chi
chi <- list.files(file.path(dir_M, "git-annex/impact_acceleration/impact/cumulative_impact"), full=TRUE,
                  pattern = "chi")

chi_stack <- stack(chi)

chi_data <- zonal(chi_stack, rgns_3nm, fun="mean", progress="text", na.rm=TRUE)

chi_data_df <- data.frame(chi_data) %>%
  gather("pressure", "value", -1) %>%
  rename("rgn_id" = zone) %>%
  mutate(year = as.numeric(substring(pressure, 5, 8))) %>%
  mutate(pressure = "chi") %>%
  inner_join(rgns_global, by="rgn_id")

write.csv(chi_data_df, "impacts/zonal_data_eez/eez_3nm_chi.csv", row.names=FALSE)

## motion chart for inspection
plot_data <- read.csv("impacts/zonal_data_eez/eez_3nm_chi.csv") %>%
  dplyr::select(rgn_name, year, value)

Motion=gvisMotionChart(plot_data, 
                       idvar="rgn_name", 
                       timevar="year")
plot(Motion)
print(Motion, file="impacts/zonal_data_eez/eez_3nm_chi.html")

```


## 3nm eez regions: chi trend


```{r}
trend <- raster(file.path(dir_M, "git-annex/impact_acceleration/impact/trend/chi_slope.tif"))

trend_data <- zonal(trend, rgns_3nm, fun="mean", progress="text", na.rm=TRUE)

trend_data_df <- data.frame(trend_data) %>%
  rename("rgn_id" = zone, "value"=mean) %>%
  mutate(pressure = "chi_trend") %>%
  inner_join(rgns_global, by="rgn_id")

write.csv(trend_data_df, "impacts/zonal_data_eez/eez_chi_3nm_trend.csv", row.names=FALSE)

```


## 3nm eez regions: impact trend

```{r}

rgns_3nm <- raster::raster(file.path(dir_M, "git-annex/globalprep/spatial/v2018/rgns_3nm_offshore_mol.tif"))
plot(rgns_3nm)

impact_trends <- list.files(file.path(dir_M, "git-annex/impact_acceleration/impact/trend/impacts"), full=TRUE)
impact_trends_stack <- stack(impact_trends)

trend_data <- zonal(impact_trends_stack, rgns_3nm, fun="mean", progress="text", na.rm=TRUE)

trend_data_df <- data.frame(trend_data) %>%
  tidyr::gather("pressure", "value", -1) %>%
  dplyr::rename("rgn_id" = zone) %>%
  dplyr::mutate(pressure = stringr::str_sub(pressure, 1, stringr::str_length(pressure)-6)) %>%
  inner_join(rgns_global, by="rgn_id")


write.csv(trend_data_df, "impacts/zonal_data_eez/eez_3nm_impacts_trend.csv", row.names=FALSE)

```

## 3nm global: impact trend

```{r}

rgns_3nm <- raster::raster(file.path(dir_M, "git-annex/globalprep/spatial/v2018/rgns_3nm_offshore_mol.tif"))
plot(rgns_3nm)

rgns_3nm_global <- rgns_3nm

rgns_3nm_global[!is.na(rgns_3nm_global)] <- 1

impact_trends <- list.files(file.path(dir_M, "git-annex/impact_acceleration/impact/trend/impacts"), full=TRUE)
impact_trends_stack <- stack(impact_trends)

trend_data <- zonal(impact_trends_stack, rgns_3nm_global, fun="mean", progress="text", na.rm=TRUE)

trend_data_df <- data.frame(trend_data) %>%
  tidyr::gather("pressure", "value", -1) %>%
  dplyr::rename("rgn_id" = zone) %>%
  dplyr::mutate(pressure = stringr::str_sub(pressure, 1, stringr::str_length(pressure)-6)) %>%
  dplyr::mutate(rgn_id = "global_3nm") %>%
  arrange(value)


write.csv(trend_data_df, "impacts/zonal_data_eez/global_3nm_impacts_trend.csv", row.names=FALSE)

```


## 3nm global regions: impacts

```{r}

rgns_3nm <- raster::raster(file.path(dir_M, "git-annex/globalprep/spatial/v2018/rgns_3nm_offshore_mol.tif"))
plot(rgns_3nm)

rgns_3nm_global <- rgns_3nm

rgns_3nm_global[!is.na(rgns_3nm_global)] <- 1

impact_files <- list.files(file.path(dir_M, "git-annex/impact_acceleration/impact/stressor_impact"), recursive = TRUE, 
                           pattern = ".tif", full="TRUE")

registerDoParallel(3)

foreach (year = 2003:2013) %dopar%{ # year = 2003
for(year in 2003:2013){
    
impact_year_files <- grep(year, impact_files, value=TRUE)

impact_stack <- raster::stack(impact_year_files)

impact_data <- raster::zonal(impact_stack, rgns_3nm_global, fun="mean", progress="text", na.rm=TRUE)

impact_data_df <- data.frame(impact_data) %>%
  tidyr::gather("pressure", "value", -1) %>%
  dplyr::rename("rgn_id" = zone) %>%
  dplyr::mutate(year = as.numeric(stringr::str_sub(pressure, -4,-1))) %>%
  dplyr::mutate(pressure = stringr::str_sub(pressure, 1, stringr::str_length(pressure)-5)) %>%
  dplyr::mutate(rgn_id = "global_3nm") %>%
  dplyr::arrange(value)

write.csv(impact_data_df, sprintf("impacts/zonal_data_eez/global_3nm_%s_impacts.csv", year), row.names=FALSE)

}

```

## Global: impact and chi trends

```{r}

global <- ocean
plot(global)

impact_trends <- list.files(file.path(dir_M, "git-annex/impact_acceleration/impact/trend/impacts"), full=TRUE)
chi_trend <- file.path(dir_M, "git-annex/impact_acceleration/impact/trend/chi_slope.tif")
trend <- stack(c(impact_trends, chi_trend))

trend_data <- zonal(trend, global, fun="mean", progress="text", na.rm=TRUE)

trend_data_df <- data.frame(trend_data) %>%
  rename("ocean" = zone) %>%
  gather("impact", "value", -1) %>%
  arrange(value)

write.csv(trend_data_df, "impacts/zonal_data_eez/global_impacts_trend.csv", row.names=FALSE)

```

## Global: impacts

```{r}

global <- ocean
plot(global)

impact_files <- list.files(file.path(dir_M, "git-annex/impact_acceleration/impact/stressor_impact"), recursive = TRUE, 
                           pattern = ".tif", full="TRUE")

impact_files <- impact_files[-(grep("uv", impact_files))]
impact_files <- impact_files[-(grep("benthic_str", impact_files))]

registerDoParallel(3)

foreach (year = 2003:2013) %dopar%{ # year = 2003
for(year in 2003:2013){
    
impact_year_files <- grep(year, impact_files, value=TRUE)

impact_stack <- raster::stack(impact_year_files)

impact_data <- raster::zonal(impact_stack, global, fun="mean", progress="text", na.rm=TRUE)

impact_data_df <- data.frame(impact_data) %>%
  tidyr::gather("pressure", "value", -1) %>%
  dplyr::rename("rgn_id" = zone) %>%
  dplyr::mutate(year = as.numeric(stringr::str_sub(pressure, -4,-1))) %>%
  dplyr::mutate(pressure = stringr::str_sub(pressure, 1, stringr::str_length(pressure)-5)) %>%
  dplyr::mutate(rgn_id = "global") %>%
  dplyr::arrange(value)

write.csv(impact_data_df, sprintf("impacts/zonal_data_eez/global_%s_impacts.csv", year), row.names=FALSE)

}

```

### Global: CHI

```{r}

global <- ocean
### chi
chi <- list.files(file.path(dir_M, "git-annex/impact_acceleration/impact/cumulative_impact"), full=TRUE,
                  pattern = "chi")

chi_stack <- stack(chi)

chi_data <- zonal(chi_stack, global, fun="mean", progress="text", na.rm=TRUE)

chi_data_df <- data.frame(chi_data) %>%
  gather("pressure", "value", -1) %>%
  rename("rgn_id" = zone) %>%
  mutate(year = as.numeric(substring(pressure, 5, 8))) %>%
  mutate(pressure = "chi") %>%
  inner_join(rgns_global, by="rgn_id")

write.csv(chi_data_df, "impacts/zonal_data_eez/global_chi.csv", row.names=FALSE)


```