---
title: "Reviewing raw-ish data"
output: html_document
---

```{r setup, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

library(rgdal)
library(raster)
library(RColorBrewer)

source("https://raw.githubusercontent.com/OHI-Science/ohiprep_v2018/master/src/R/spatial_common.R")

cols = rev(colorRampPalette(brewer.pal(11, 'Spectral'))(255)) # rainbow color scheme
cols <- c(NA, cols)
```

## Fisheries
These are the data layers prior to being rescaled from 0-1.


### Reference point compare
99.99th quantile across global rasters and years:

```{r, echo=FALSE, warning=FALSE, error=FALSE}
refs <- read.csv("stressors/comm_fish/comm_fish_ref_points.csv")
refs
```

### Demersal nondestructive high bycatch.

(0 and really low values not displayed)

```{r, echo=FALSE, warning=FALSE, error=FALSE}

dem_nd_hb <- raster(file.path(dir_M, "git-annex/impact_acceleration/data_check/new_dem_nondest_hb.tif"))

#calc(dem_nd_hb, fun = function(x)(log(x + 1)), filename = file.path(dir_M, "git-annex/impact_acceleration/data_check/new_dem_nondest_hb_log.tif"))

plot(dem_nd_hb, col=cols)

```

0 and low values displayed as blue.

```{r, echo=FALSE, warning=FALSE, error=FALSE}
plot(dem_nd_hb, col=cols[-1])

```

ln(x+1) to display low values

```{r, echo=FALSE, warning=FALSE, error=FALSE}
tmp <- raster(file.path(dir_M, "git-annex/impact_acceleration/data_check/new_dem_nondest_hb_log.tif"))
plot(tmp, col=cols)

```

```{r, echo=FALSE, warning=FALSE, error=FALSE}
#reclassify(dem_nd_hb, cbind(0, NA), filename = file.path(dir_M, "git-annex/impact_acceleration/data_check/new_dem_nondest_hb_no_zero.tif"), progress="text", overwrite=TRUE)

tmp <- raster(file.path(dir_M, "git-annex/impact_acceleration/data_check/new_dem_nondest_hb_no_zero.tif"))

#histogram(tmp)

```

### Demersal nondestructive low bycatch

```{r, echo=FALSE, warning=FALSE, error=FALSE}

dem_nd_lb <- raster(file.path(dir_M, "git-annex/impact_acceleration/data_check/new_dem_nondest_lb.tif"))

#calc(dem_nd_lb, fun = function(x)(log(x + 1)), filename = file.path(dir_M, "git-annex/impact_acceleration/data_check/new_dem_nondest_lb_log.tif"))

plot(dem_nd_lb, col=cols)

```

0 and low values displayed as blue.

```{r, echo=FALSE, warning=FALSE, error=FALSE}
plot(dem_nd_lb, col=cols[-1])

```

ln(x+1) to display low values

```{r, echo=FALSE, warning=FALSE, error=FALSE}
tmp <- raster(file.path(dir_M, "git-annex/impact_acceleration/data_check/new_dem_nondest_lb_log.tif"))
plot(tmp, col=cols)

```

```{r, echo=FALSE, warning=FALSE, error=FALSE}
#reclassify(dem_nd_lb, cbind(0, NA), filename = file.path(dir_M, "git-annex/impact_acceleration/data_check/new_dem_nondest_lb_no_zero.tif"), progress="text", overwrite=TRUE)

tmp <- raster(file.path(dir_M, "git-annex/impact_acceleration/data_check/new_dem_nondest_lb_no_zero.tif"))

#histogram(tmp)

```


### Demersal nondestructive high bycatch.

(0 and really low values not displayed)

```{r, echo=FALSE, warning=FALSE, error=FALSE}

dem_nd_hb <- raster(file.path(dir_M, "git-annex/impact_acceleration/data_check/new_dem_nondest_hb.tif"))

#calc(dem_nd_hb, fun = function(x)(log(x + 1)), filename = file.path(dir_M, "git-annex/impact_acceleration/data_check/new_dem_nondest_hb_log.tif"))

plot(dem_nd_hb, col=cols)

```

0 and low values displayed as blue.

```{r, echo=FALSE, warning=FALSE, error=FALSE}
plot(dem_nd_hb, col=cols[-1])

```

ln(x+1) to display low values

```{r, echo=FALSE, warning=FALSE, error=FALSE}
tmp <- raster(file.path(dir_M, "git-annex/impact_acceleration/data_check/new_dem_nondest_hb_log.tif"))
plot(tmp, col=cols)

```

```{r, echo=FALSE, warning=FALSE, error=FALSE}
#reclassify(dem_nd_hb, cbind(0, NA), filename = file.path(dir_M, "git-annex/impact_acceleration/data_check/new_dem_nondest_hb_no_zero.tif"), progress="text", overwrite=TRUE)

tmp <- raster(file.path(dir_M, "git-annex/impact_acceleration/data_check/new_dem_nondest_hb_no_zero.tif"))

#histogram(tmp)

```

### Demersal destructive

```{r, echo=FALSE, warning=FALSE, error=FALSE}

dem_dest <- raster(file.path(dir_M, "git-annex/impact_acceleration/data_check/new_dem_dest.tif"))

# calc(dem_dest, fun = function(x)(log(x + 1)), filename = file.path(dir_M, "git-annex/impact_acceleration/data_check/dem_dest_log.tif"))

plot(dem_dest, col=cols)

```

0 and low values displayed as blue.

```{r, echo=FALSE, warning=FALSE, error=FALSE}

plot(dem_dest, col=cols[-1])

```

ln(x+1) to display low values

```{r, echo=FALSE, warning=FALSE, error=FALSE}
tmp <- raster(file.path(dir_M, "git-annex/impact_acceleration/data_check/dem_dest_log.tif"))
plot(tmp, col=cols)

```


```{r, echo=FALSE, warning=FALSE, error=FALSE}
#reclassify(dem_dest, cbind(0, NA), filename = file.path(dir_M, "git-annex/impact_acceleration/data_check/new_dem_dest_no_zero.tif"), progress="text", overwrite=TRUE)

tmp <- raster(file.path(dir_M, "git-annex/impact_acceleration/data_check/new_dem_nondest_lb_no_zero.tif"))

#histogram(tmp)

```


### Pelagic high bycatch

```{r, echo=FALSE, warning=FALSE, error=FALSE}

pel_hb <- raster(file.path(dir_M, "git-annex/impact_acceleration/data_check/new_pel_hb.tif"))

# calc(pel_hb, fun = function(x)(log(x + 1)), filename = file.path(dir_M, "git-annex/impact_acceleration/data_check/pel_hb_log.tif"))

plot(pel_hb, col=cols)

```

0 and low values displayed as blue.

```{r, echo=FALSE, warning=FALSE, error=FALSE}

plot(pel_hb, col=cols[-1])

```

ln(x+1) to display low values

```{r, echo=FALSE, warning=FALSE, error=FALSE}
tmp <- raster(file.path(dir_M, "git-annex/impact_acceleration/data_check/pel_hb_log.tif"))
plot(tmp, col=cols)

```


```{r, echo=FALSE, warning=FALSE, error=FALSE}
#reclassify(pel_hb, cbind(0, NA), filename = file.path(dir_M, "git-annex/impact_acceleration/data_check/new_pel_hb_no_zero.tif"), progress="text", overwrite=TRUE)

tmp <- raster(file.path(dir_M, "git-annex/impact_acceleration/data_check/new_pel_hb_no_zero.tif"))

#histogram(tmp)

```


### Pelagic low bycatch

```{r, echo=FALSE, warning=FALSE, error=FALSE}

pel_lb <- raster(file.path(dir_M, "git-annex/impact_acceleration/data_check/new_pel_lb.tif"))

# calc(pel_lb, fun = function(x)(log(x + 1)), filename = file.path(dir_M, "git-annex/impact_acceleration/data_check/pel_lb_log.tif"))

plot(pel_lb, col=cols)

```

0 and low values displayed as blue.

```{r, echo=FALSE, warning=FALSE, error=FALSE}

plot(pel_lb, col=cols[-1])

```

ln(x+1) to display low values

```{r, echo=FALSE, warning=FALSE, error=FALSE}
tmp <- raster(file.path(dir_M, "git-annex/impact_acceleration/data_check/pel_lb_log.tif"))
plot(tmp, col=cols)

```


```{r, echo=FALSE, warning=FALSE, error=FALSE}
#reclassify(pel_lb, cbind(0, NA), filename = file.path(dir_M, "git-annex/impact_acceleration/data_check/new_pel_lb_no_zero.tif"), progress="text", overwrite=TRUE)

tmp <- raster(file.path(dir_M, "git-annex/impact_acceleration/data_check/new_pel_lb_no_zero.tif"))

#histogram(tmp)

```





