---
title: 'Figures'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '../../ohiprep_v2018/src/templates/ohi_hdr.html'
  pdf_document:
    toc: true
---

```{r}
library(tidyr)
library(dplyr)
library(raster)
library(RColorBrewer)
library(rgdal)
library(sf)
library(gstat)
library(stringr)
library(fields)
library(cowplot)

# load spatial files (ocean raster and regions shapefile)

source("https://raw.githubusercontent.com/OHI-Science/ohiprep_v2018/master/src/R/spatial_common.R")

regions_shape <- as(regions, "Spatial")
land <- regions_shape[regions_shape$rgn_type %in% c("land", "land-disputed", "land-noeez", "eez-inland"), ]


```

# Map of CHI

```{r chi_map}

## function to make plots

legend.shrink <- 0.7
legend.width <- 0.6


raster_breaks <- function(raster_data, title, title_legend=NULL, 
                          color_breaks=my_color_breaks, cols=my_cols,
                          legend_break_labels=my_legend_break_labels, 
                          legend_break_locations=my_legend_break_locations){
  par(mar=c(2,2,2,2)) # bottom, left, top, and right
  par(oma=c(0,0,0,0))
plot(raster_data, col=cols, axes=FALSE, box=FALSE, breaks=color_breaks, legend=FALSE)
  title(title, line=0)
# add axis with fields package function:
#def_breaks <- seq(0, length(myBreaks), length.out=length(myBreaks))
#loc_breaks_labels <- def_breaks[seq(1, length(def_breaks), 2)]
#break_labels <- myBreaks[seq(1, length(myBreaks), 2)]
break_locations <- seq(0, length(color_breaks), length.out=length(color_breaks))
fields::image.plot(raster_data, #zlim = c(min(myBreaks), max(myBreaks)), 
           legend.only = TRUE, 
           legend.shrink=legend.shrink,
           legend.width=legend.width,
           col = cols,
           #legend.lab=title_legend,
           breaks=break_locations,
           axis.args=list(cex.axis=0.6, at=legend_break_locations, labels=legend_break_labels))

plot(land, add=TRUE, border="gray80", col="gray90", lwd=0.5)
#plot(raster_data, col=cols, axes=FALSE, box=FALSE, breaks=myBreaks, legend=FALSE, add=TRUE)
}

my_color_breaks <- c(-1, 0, 0.2, 0.4, 0.6, 0.8, 1, 1.25, 1.5, 1.75, 2, 2.5, 3, 3.5, 4.0, 5, 100)
my_cols = rev(colorRampPalette(brewer.pal(11, 'Spectral'))(length(my_color_breaks)-1)) 

break_locations <- seq(0, length(my_color_breaks), length.out=length(my_color_breaks)) # calculated in fucntion, but needed for below
my_legend_break_labels <- c(0, 0.2, 0.6, 1, 1.5, 2, 3, 4.0, ">5")
my_legend_break_locations <- break_locations[seq(1, length(break_locations), 2)]


for(year in 2003:2013){ # year = 2013
png(sprintf("impacts/figures/chi_maps/chi_%s.png", year), res=600, width=6, height=3, units="in")  
p_rast <- raster(file.path(dir_M, sprintf("git-annex/impact_acceleration/impact/cumulative_impact/chi_%s.tif", year)))
raster_breaks(raster_data=p_rast,  title="", 
              cols=my_cols, color_breaks = my_color_breaks,
              legend_break_labels = my_legend_break_labels,
              legend_break_locations = my_legend_break_locations)
dev.off()
}


```

# Plot map of trend

```{r trend_map}

slope <- raster(file.path(dir_M, 'git-annex/impact_acceleration/impact/trend/chi_slope.tif'))
overlay_rast <- raster(file.path(dir_M,  'git-annex/impact_acceleration/impact/trend/sig_overlay.tif'))

# cols <- c("#9E0142",
#           "#9E0041", "#B92548", "#D43E4E", "#E45649", "#F36C43", "#F98E51", "#FDAD60",  
#           "#FDAE61", "#FEE08B", "#FFFFBF", "#E6F598", "#ABDDA4", "#66C2A5", "#3288BD", "#5E4FA2")

#colorRampPalette(brewer.pal(11, 'Spectral'))(18)
cols = rev(colorRampPalette(c("#9E0142", "#AE1245", "#BE2449", "#CE364D", "#DA464C", "#E35449", "#EC6145", "#F47044", "#F7834D", "#F99756", "#FCAA5F",             
               "#FEF0A7", "#F3FAAD",  
              "#CAE99D", "#A6DBA4", "#7ECBA4", "#59B4AA", "#3B92B8", "#4470B1", "#5E4FA2"))(254)) # 

png("impacts/figures/trend/chi_trend.png", res=600, width=6, height=3, units="in")  
par(mar=c(2, 2, 2, 2))
par(oma=c(0,0,0,0))
plot(slope, col=cols, axes=FALSE, box=FALSE, legend=FALSE)
plot(overlay_rast, col="#ffffff", add=TRUE, legend=FALSE, box=FALSE)
plot(land, add=TRUE, border="gray80", col="gray90", lwd=0.5)
par(mfrow=c(1, 1), mar=c(2, 0, 2, 0), new=FALSE)
plot(slope, legend.only=TRUE, legend.shrink=.7, legend.width=.6, col=cols, axis.args = list(cex.axis = 0.6))
dev.off()

```


# Map of CHI...no SST

```{r chi_map_noSST}

# create CHI map without SST
chi <- raster(file.path(dir_M, "git-annex/impact_acceleration/impact/cumulative_impact/chi_2013.tif"))
sst <- raster(file.path(dir_M, "git-annex/impact_acceleration/impact/stressor_impact/sst_2013.tif"))

chi_sst <- stack(chi, sst)
overlay(chi_sst, fun=function(x,y){x - y}, 
                    filename = file.path(dir_M, 'git-annex/impact_acceleration/impact/figures/chi_minus_sst.tif'), overwrite=TRUE, progress="text")

chi_no_sst <- raster(file.path(dir_M, 'git-annex/impact_acceleration/impact/figures/chi_minus_sst.tif'))

## function to make plots

legend.shrink <- 0.7
legend.width <- 0.6

raster_breaks <- function(raster_data, title, title_legend=NULL, myBreaks, cols){
  par(mar=c(2,2,2,2)) # bottom, left, top, and right
  par(oma=c(0,0,0,0))
plot(raster_data, col=cols, axes=FALSE, box=FALSE, breaks=myBreaks, legend=FALSE)
  title(title, line=0)
# add axis with fields package function:
def_breaks <- seq(0, length(myBreaks), length.out=length(myBreaks))
image.plot(raster_data, #zlim = c(min(myBreaks), max(myBreaks)), 
           legend.only = TRUE, 
           legend.shrink=legend.shrink,
           legend.width=legend.width,
           col = cols,
           legend.lab=title_legend,
            breaks=def_breaks,
           lab.breaks=c(0, NA, 0.2, NA, 0.6, NA, 1, NA, 1.5, NA, 2, NA, 3, NA, 4.0, NA, ">5"),
           axis.args = list(cex.axis = 0.6))

plot(land, add=TRUE, border="gray80", col="gray90", lwd=0.5)
#plot(raster_data, col=cols, axes=FALSE, box=FALSE, breaks=myBreaks, legend=FALSE, add=TRUE)
}

my_breaks <- c(-1, 0, 0.2, 0.4, 0.6, 0.8, 1, 1.25, 1.5, 1.75, 2, 2.5, 3, 3.5, 4.0, 5, 100) 
cols = rev(colorRampPalette(brewer.pal(11, 'Spectral'))(length(my_breaks)-1)) #(length(my_breaks)-2))


png("impacts/figures/chi_maps/chi_noSST.png", res=600, width=6, height=3, units="in")  
raster_breaks(raster_data=chi_no_sst,  
              myBreaks=my_breaks, cols=cols, title="")
dev.off()


```

# Plot map of trend..no SST

```{r trend_map_noSST}

chi_slope <- raster(file.path(dir_M, 'git-annex/impact_acceleration/impact/trend/chi_slope.tif'))
sst_slope <- raster(file.path(dir_M, 'git-annex/impact_acceleration/impact/trend/impacts/sst_trend.tif'))

chi_sst_slope <- stack(chi_slope, sst_slope)
overlay(chi_sst_slope, fun=function(x,y){x - y}, 
                    filename = file.path(dir_M, 'git-annex/impact_acceleration/impact/figures/chi_minus_sst_trend.tif'), overwrite=TRUE, progress="text")

slope_noSST <- raster(file.path(dir_M, 'git-annex/impact_acceleration/impact/figures/chi_minus_sst_trend.tif'))


cols = rev(colorRampPalette(c("#9E0142", "#9E0142", "#9E0142", "#9E0142", "#D53E4F", "#D53E4F", "#D53E4F", "#F46D43", "#F46D43", "#FDAE61", "#FDAE61",             
              "#FFFFBF", "#FFFFBF",  
             "#66C2A5", "#66C2A5", "#3288BD", "#3288BD", "#5E4FA2", "#5E4FA2",  "#5E4FA2"))(255)) # 

png("impacts/figures/trend/chi_noSST_trend.png", res=600, width=6, height=3, units="in")  
par(mar=c(2, 2, 2, 2))
par(oma=c(0,0,0,0))
plot(slope_noSST, col=cols, axes=FALSE, box=FALSE, legend=FALSE)
plot(land, add=TRUE, border="gray80", col="gray90", lwd=0.5)
par(mfrow=c(1, 1), mar=c(2, 0, 1, 0), new=FALSE)
plot(slope_noSST, legend.only=TRUE, legend.shrink=.7, legend.width=.5, col=cols, axis.args = list(cex.axis = 0.6))
dev.off()

```




# High/low trend and chi raster
```{r extremes_map}
#################################
# Raster identifying negative and positive slope trends 
slope <- raster(file.path(dir_M, 'git-annex/impact_acceleration/impact/trend/chi_slope.tif'))

# negative and zero slope is 1 and positive slope 0
# default is right is closed, which means the interval includes the endpoint
m <- c(-Inf, 0, 0,   
       0, Inf, 1)

rclmat <- matrix(m, ncol=3, byrow=TRUE)

reclassify(slope, rclmat, 
           filename = file.path(dir_M, 'git-annex/impact_acceleration/impact/figures/slope_pos_neg.tif'), overwrite=TRUE, progress="text")

slope_p_n <- raster(file.path(dir_M, 'git-annex/impact_acceleration/impact/figures/slope_pos_neg.tif'))
plot(slope_p_n)
freq(slope_p_n)
#    value     count
#     0  79187232
#     1 337003569
#    NA 329175249
79187232/(79187232+337003569)
# 19% decreasing trend

# ###################################
# ## slope 
# 
# identify quantiles
slope <- raster(file.path(dir_M, 'git-annex/impact_acceleration/impact/trend/chi_slope.tif'))
raster::quantile(slope, probs=c(0.1, 0.20, 0.25, 0.5, 0.75, 0.80, 0.90))
# #        10%          20%          25%          50%          75%          80%          90% 
# # -0.018059537  0.001269741  0.006122565  0.032733742  0.074715607  0.084770553  0.110365756 

chi <- raster(file.path(dir_M, "git-annex/impact_acceleration/impact/cumulative_impact/chi_2013.tif"))
raster::quantile(chi, probs=c(0.1, 0.20, 0.25, 0.5, 0.75, 0.80, 0.90))


## Identify rasters with values <20th and >80th quantiles

## Slope: Identify rasters with <=0 and >80th quantile
slope <- raster(file.path(dir_M, 'git-annex/impact_acceleration/impact/trend/chi_slope.tif'))

m <- c(-Inf, 0, 0,
       0, 0.084770553, NA,
       0.084770553, Inf ,1)
rclmat <- matrix(m, ncol=3, byrow=TRUE)

reclassify(slope, rclmat,
           filename = file.path(dir_M, 'git-annex/impact_acceleration/impact/figures/slope_upper_lower_quants.tif'), overwrite=TRUE, progress="text")

slope_extreme <-  raster(file.path(dir_M, 'git-annex/impact_acceleration/impact/figures/slope_upper_lower_quants.tif'))
plot(slope_extreme)
#freq(slope_extreme)


## CHI
chi <- raster(file.path(dir_M, "git-annex/impact_acceleration/impact/cumulative_impact/chi_2013.tif"))

m <- c(-Inf, 0.6165478, -1,   
       0.6165478, 1.5580050, NA,  
       1.5580050, Inf ,1)
rclmat <- matrix(m, ncol=3, byrow=TRUE)

reclassify(chi, rclmat, 
           filename = file.path(dir_M, 'git-annex/impact_acceleration/impact/figures/chi_upper_lower_quants.tif'), overwrite=TRUE, progress="text")

chi_extreme <-  raster(file.path(dir_M, 'git-annex/impact_acceleration/impact/figures/chi_upper_lower_quants.tif'))
plot(chi_extreme)
#freq(chi_extreme)

## combine the two

chi_slope <- stack(slope_extreme, chi_extreme)
overlay(chi_slope, fun=function(x,y){x + y}, 
                    filename = file.path(dir_M, 'git-annex/impact_acceleration/impact/figures/chi_slope_extremes.tif'), overwrite=TRUE, progress="text")

## check it out
chi_slope <- raster(file.path(dir_M, 'git-annex/impact_acceleration/impact/figures/chi_slope_extremes.tif'))
plot(chi_slope)
freq(chi_slope)
#    -1  20904567 3.2%, low chi, decreasing trend
#     0    173405 <1%, low chi, fast increasing trend
#     1  10391962 1.6%, high chi, decreasing trend
#     2  31476395 4.9%, high chi, fast increasing trend
#    NA 682419721
# total ocean: 648429346

my_breaks <- c(-2, -1, 0, 1, 3)
beyonce_palettes[[18]]
beyonce_palette(127)
# blue, light brown , black , red
cols = c( "#018AC4", "#EBC915", "#550133", "#CF2154")
labels=c("low/\ndecrease", "low/\nfast increase", "high/\ndecrease", "high/\nfast increase")

legend.shrink <- 0.6
legend.width <- 0.6
png("impacts/figures/trend/chi_trend_extremes_4_cats.png", res=600, width=6, height=3, units="in") 
par(mar=c(2, 2, 2, 2))
par(oma=c(0,0,0,0))
plot(chi_slope, col=cols, axes=FALSE, box=FALSE, breaks=my_breaks, legend=FALSE)
# add axis with fields package function:
def_breaks <- c(0, 1, 2, 3, 4)
fields::image.plot(chi_slope, #zlim = c(min(myBreaks), max(myBreaks)), 
           legend.only = TRUE, 
           legend.shrink=legend.shrink,
           legend.width=legend.width,
           col = cols,
           breaks=def_breaks,
           #lab.breaks=labels,
           axis.args = list(cex.axis = 0.8, at=c(0.5, 1.5, 2.5, 3.5), labels=labels))
plot(land, add=TRUE, border="gray80", col="gray90", lwd=0.5)
dev.off()

```



# Trend for each 3nm eez impact circleplot

```{r 3nm_eez_trend_circleplot}

# UN georegions
un_rgn <- UNgeorgn_nm %>%
  dplyr::select(rgn_id, georegion = r1_label)

# CHI for each region
chi <- read.csv("impacts/zonal_data_eez/eez_chi_3nm_trend.csv") %>%
  dplyr::select(rgn_id, rgn_name, chi=value) %>%
  left_join(un_rgn, by="rgn_id")


# get georegion ranks based on mean chi values, and use to rank chi data
rank_georgn <- chi %>%
  group_by(georegion) %>%
  summarize(mean_chi = mean(chi),
            count = length(chi)) %>%
  arrange(mean_chi)

chi$georegion <- factor(chi$georegion, levels=rank_georgn$georegion)


### Organize the chi data and add a few extra variables to help plotting
# Including empty spaces to add a y-axes

rank_rgn <- chi %>%
  arrange(georegion, chi)

# add empty space
empty_bar <- 6
to_add <- data.frame(matrix(NA, empty_bar, ncol(rank_rgn)) )
colnames(to_add) = colnames(rank_rgn)
to_add$rgn_name <- as.character(1:empty_bar)
rank_rgn  <- rbind(to_add, rank_rgn)

# modify region names to be shorter
rank_rgn <- rank_rgn %>%
  mutate(rgn_name_short = rgn_name, 
         rgn_name_short = gsub("Islands", "Isl", rgn_name_short),
         rgn_name_short = gsub("Island", "Isl", rgn_name_short),
         rgn_name_short = gsub("Democratic", "Dem", rgn_name_short),
         rgn_name_short = gsub("Republic", "Rep", rgn_name_short),
         rgn_name_short = gsub("South", "S", rgn_name_short),
         rgn_name_short = gsub("American", "Am", rgn_name_short),
         rgn_name_short = gsub("the United States", "US", rgn_name_short),
         rgn_name_short = gsub("Territory", "Terr", rgn_name_short),
         rgn_name_short = gsub("Saint", "St", rgn_name_short),
         rgn_name_short = gsub(" and ", " & ", rgn_name_short),
         rgn_name_short = gsub("Republique", "Rep", rgn_name_short),
         rgn_name_short = gsub("Dem Rep of the", "Dem Rep of", rgn_name_short),
         rgn_name_short = gsub("Georgia and the", "Georgia and", rgn_name_short),
         rgn_name_short = gsub("St Vincent and the", "St Vincent and", rgn_name_short),
         rgn_name_short = gsub("Howland Isl & Baker Isl", "Howland & Baker Isl", rgn_name_short),
         rgn_name_short = gsub("Northern", "N", rgn_name_short), 
         rgn_name_short = gsub("Reunion", "Reunion   ", rgn_name_short))


# some code to orient the country labels
sequence_length = length(unique(rank_rgn$rgn_name))
first_sequence = c(1:(sequence_length%/%2)) 
second_sequence = c((sequence_length%/%2+1):sequence_length) 
first_angles = c(90 - 180/length(first_sequence) * first_sequence)
second_angles = c(-90 - 180/length(second_sequence) * second_sequence)

rank_rgn$angle <- c(first_angles,second_angles)
rank_rgn$hjust <- c(rep(0, length(first_sequence)), 
                       rep(1, length(second_sequence)))

# color for region labels
rank_rgn <- rank_rgn %>%
  mutate(rgn_name = factor(rgn_name, unique(rgn_name))) %>%
  mutate(georegion = factor(georegion, unique(georegion))) %>%
  mutate(color = "black") %>%
  mutate(color = ifelse(is.na(chi), "white", color))


########### marking georegions

# add column to identify when georegion changes
rgn_shift <- rank_rgn %>%
  mutate(georegion = ifelse(is.na(georegion), "tmp", georegion)) %>%
  mutate(georegion = as.factor(georegion)) %>%
  mutate(region_shift = as.numeric(georegion) - lag(as.numeric(georegion)), default=first(as.numeric(georegion)))

rgn_shift <- which(rgn_shift$region_shift > 0)
rgn_shift <- c(1, rgn_shift) -0.5
rgn_shift <- data.frame(rgn_shift_x=rgn_shift,
                        georegion = rank_georgn$georegion,
                        name_x= c(5, 25, 55, 70, 110, 145, 200), # 140
                        name_y=c(0.12, 0.1, 0.1, 0.1, 0.11, 0.1, 0.11))
rgn_shift <- rgn_shift %>%
  mutate(georegion = as.character(georegion)) %>%
  mutate(georegion = ifelse(georegion == "Latin America and the Caribbean", "Latin America and\nthe Caribbean",
                            georegion))

# get impact data
impacts <- read.csv("impacts/zonal_data_eez/eez_3nm_impacts_trend.csv") %>%
  left_join(un_rgn)

# ## add some blanks to separate groups
to_add <-  data.frame( matrix(NA, empty_bar*nlevels(as.factor(impacts$pressure)), ncol(impacts)) )
colnames(to_add) <- colnames(impacts)
to_add$pressure <- rep(levels(as.factor(impacts$pressure)), each=empty_bar)
to_add$value <-  0
to_add$rgn_name <- as.character(rep(1:empty_bar, nlevels(as.factor(impacts$pressure)))) 

impacts <- rbind(to_add, impacts, to_add)


## rank of pressure categories
impacts %>%
  group_by(pressure) %>%
  summarize(mean = mean(value)) %>%
  arrange(mean)

pressure_name <- data.frame(pressure = c("sst", "slr", "oa", 
                                                  "shipping",
                                                  "light", "nutrient", "direct_human", "organic",
                                                  "pel_hb", "dem_nondest_hb","pel_lb", 
                                                  "dem_nondest_lb", "dem_dest", "art_fish"),
                            pressure_name = c("sst", "slr", "oa",
                                              "shipping", "light pollution", 
                                              "nutrient pollution", "direct human", "organic pollution",  
                                              "comm fish: pel hb", "comm fish: dem nondest hb", "comm fish: pel lb", 
                                              "comm fish: dem nondest lb", "comm fish: dem dest", "artisanal fishing"))

impacts <- impacts %>%
  left_join(pressure_name, by = "pressure")

impacts$pressure_name <- factor(impacts$pressure_name, levels=rev(pressure_name$pressure_name))
impacts$rgn_name <- factor(impacts$rgn_name, levels=unique(rank_rgn$rgn_name))


library(beyonce)

myPalette <- c(beyonce_palette(18, 15, type=c("continuous"))[1:6],
               beyonce_palette(18, 25, type=c("continuous"))[15:18],
               beyonce_palette(18, 15, type=c("continuous"))[8],
               beyonce_palette(18, 20, type=c("continuous"))[16:19])
               

## some theme stuff to make the circle plot look nice

circle_theme <- theme(axis.line=element_blank(),
      axis.text.y=element_blank(),
      axis.ticks=element_blank(),
      axis.title.x=element_blank(),
      axis.title.y=element_blank(),
      legend.position="none",
      panel.background=element_blank(),
      panel.border=element_blank(),
      panel.grid.major=element_blank(),
      panel.grid.minor=element_blank(),
      plot.background=element_blank(),
      axis.text.x = element_blank())

# circle plot
# https://www.r-graph-gallery.com/299-circular-stacked-barplot/
p <- ggplot(data=impacts, aes(x=rgn_name, y=value, fill=pressure_name)) + 
  geom_bar(stat="identity") +
  geom_errorbar(aes(x = 1, ymin = -0.15, ymax=0.25), alpha=0)  +
  geom_text(data=rank_rgn, aes(x=rgn_name, y=0.16, label=rgn_name_short, angle=angle, hjust=hjust, color=color), inherit.aes = FALSE) +
  geom_segment(x = 5, y = 0, xend = dim(rank_rgn)[1]+1, yend = 0, colour = "black", alpha=1, size=0.5) +
  geom_segment(x = 7, y = -0.05, xend = dim(rank_rgn)[1]-1, yend = -0.05, colour = "gray", alpha=1, size=0.5) +
  annotate("text", x = c(3,3,3,3), y = c(-0.05, 0, 0.05, 0.1), label = c(-0.05, 0, 0.05, 0.1), color="darkgrey", angle=-8, size=4) +
  scale_colour_identity() +
  scale_fill_manual(values=myPalette) +
  coord_polar() +
  geom_segment(data = rgn_shift, aes(x=rgn_shift_x, xend=rgn_shift_x, y=rep(-0.05, dim(rgn_shift)[1]), yend=rep(0.2, dim(rgn_shift)[1])), colour="gray", size=0.5, inherit.aes=FALSE) +
geom_text(data=rgn_shift, aes(x=name_x, y=name_y, label=georegion), inherit.aes=FALSE, size=5) +
  #geom_point(data=chi, aes(y=chi, x=rgn_name), fill="black", shape="|", size=2.5) +
  circle_theme

p
ggsave('impacts/figures/trend/eez_3nm_trend_circleplot.jpg', height=18, width=18, units=c("in"))



```




### Habitat/impact heat map

```{r hab_impact_heat_map}

## get habitat names that are readable
habitat_nms <- read.csv("habitats/habitat.csv") %>%
  dplyr::select(habitat=tif_name, name = plot_name3)

### get impact data for each habitat
hab_imp <- read.csv("impacts/zonal_data_habitat/habitat_pressure_2013.csv") %>%
  left_join(habitat_nms, by="habitat") 

### rank of habitats for ordering
total_hab_data <- hab_imp %>%
  group_by(name) %>%
  summarize(impact_on_hab=sum(mean)) %>%
  arrange(impact_on_hab)
  

## rank of pressure categories for ordering 
## (although the categories weren't super coordinated)
hab_imp %>%
  group_by(impact) %>%
  summarize(mean = mean(mean)) %>%
  arrange(mean)

impact_name <- data.frame(impact = c("sst", "oa", "slr", 
                                                  "shipping",
                                                  "nutrient", "organic", "direct_human", "light",
                                                  "pel_hb", "pel_lb", "dem_dest", "dem_nondest_hb",
                                                  "dem_nondest_lb", "art_fish"),
                            impact_name = c("sst", "oa", "slr", 
                                              "shipping", 
                                              "nutrient pollution", "organic pollution", "direct human", "light pollution",
                                              "comm fish: pel hb", "comm fish: pel lb", "comm fish: dem dest", 
                                              "comm fish: dem nondest hb", "comm fish: dem nondest lb",
                                              "artisanal fishing"))

hab_imp <- hab_imp %>%
  left_join(impact_name, by = "impact")

# order factors for plotting
hab_imp$name <- factor(hab_imp$name, 
                             levels = total_hab_data$name)

hab_imp$impact_name <- factor(hab_imp$impact_name, levels=rev(impact_name$impact_name))


# assign color scale
cols <- c("#9E0142",
          "#9E0041", "#B92548", "#D43E4E", "#E45649", "#F36C43", "#F98E51", "#FDAD60",  
          "#FDAE61", "#FEE08B", "#FFFFBF", "#E6F598", "#ABDDA4", "#66C2A5", "#3288BD", "#5E4FA2")


# plot the data
tmp <- ggplot(hab_imp, aes(y=impact_name, x=name)) +
  geom_tile(aes(fill=mean, 
                text=sprintf("habitat: %s \nimpact: %s \ntotal impact: %s" , name, impact_name, round(mean, 2))), color="white") +
  scale_fill_gradientn(colors=rev(cols), name="Impact") +
  ylab("") +
  xlab("") +
  theme_bw() + 
  theme(panel.border = element_blank(), 
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black"),
        axis.text.x=element_text(angle=45, hjust=1, vjust=1, size=11),
        axis.text.y=element_text(size=11),
        plot.margin=unit(c(1,1,1,1.5),"cm")) 
tmp
ggsave("impacts/figures/habitat/carpet_plot_habitat_impact.png", width=10, height=5, units=c("in"), dpi=300)

```

## Habitat Trend and CHI scatterplot
```{r}

filenames <- list.files("impacts/zonal_data_habitat", pattern="impacts.csv", full=TRUE)
datalist = lapply(filenames, function(x){read.csv(file=x,header=T)})

hab_chi <- rbind_all(datalist) %>%
  filter(year==2013)

hab_trends <- list.files("impacts/zonal_data_habitat", pattern="trend.csv", full=TRUE)
datalist = lapply(hab_trends, function(x){read.csv(file=x,header=T)})
  
hab_data <- bind_rows(datalist) %>%
  group_by(habitat) %>%
  summarize(trend_total=sum(trend)) %>%
  left_join(hab_chi)

hab_names <- read.csv("habitats/habitat.csv") %>%
  dplyr::select(habitat=tif_name, name=plot_name3)

# combine habitats with same raster locations
hab_data <- hab_data %>%
  left_join(hab_names) %>%
  rename(trend = trend_total,
            chi = impact) %>%
  filter(habitat != "ice")

  
library(ggrepel)
library(cowplot)
ggplot(hab_data, aes(x=chi, y=trend)) +
  geom_point(size=5, alpha=0.5) +
  geom_text_repel(aes(label = name),
                 # box.padding   = 0.35, 
                #  point.padding = 0.5,
                  segment.color = 'grey50',
                segment.size = 0.2,
                point.padding = 0.2,
                size = 4) +
  labs(x = "Cumulative human impact, 2013", y = "Trend, cumulative human impact")

ggsave("impacts/figures/avg_chi_vs_trend_habitat.png", width=8, height=5, units=c("in"), dpi=300)

```



# Trend for each habitat impact circleplot

```{r}

hab_trends <- list.files("impacts/zonal_data_habitat", pattern="trend.csv", full=TRUE)
datalist = lapply(hab_trends, function(x){read.csv(file=x,header=T)})
  
hab_trend_df <- bind_rows(datalist)

trend_data <- read.csv("habitats/habitat.csv") %>%
  dplyr::select(habitat=tif_name, name=plot_name3) %>%
  left_join(hab_trend_df) %>%
  filter(name != "ice edge")
  

# determine which habitats have the largest/smallest trends:
total_trend <- trend_data %>%
  group_by(name) %>%
  summarize(trend_total = sum(trend)) %>%
  arrange(trend_total) 


# some code to create a naming file that 
# orients the country labels
# and adds a gap in the circle plot for a axis label

empty_bar <- 1
to_add <- data.frame(matrix(NA, empty_bar, ncol(total_trend)) )
colnames(to_add) = colnames(total_trend)
to_add$name <- as.character(1:empty_bar)
total_trend  <- rbind(to_add, total_trend)

sequence_length = dim(total_trend)[1]
first_sequence = c(1:(sequence_length%/%2)) 
second_sequence = c((sequence_length%/%2+1):sequence_length) 
first_angles = c(90 - 160/length(first_sequence) * first_sequence)
second_angles = c(-90 - 185/length(second_sequence) * second_sequence)

total_trend$angle <- c(first_angles,second_angles)
total_trend$hjust <- c(rep(0, length(first_sequence)), 
                    rep(1, length(second_sequence)))

total_trend <- total_trend %>%
  mutate(name = factor(name, unique(name))) %>%
  mutate(color = "black") %>%
  mutate(color = ifelse(is.na(trend_total), "white", color))

## add empty data rows to full trend data here
to_add <-  data.frame( matrix(NA, empty_bar*nlevels(as.factor(trend_data$impact)), ncol(trend_data)))
colnames(to_add) <- colnames(trend_data)
to_add$trend <-  0
to_add$impact <- rep(unique(trend_data$impact), empty_bar)
to_add$name <- as.character(rep(1:empty_bar, nlevels(as.factor(trend_data$impact)))) 

trend_data <- bind_rows(to_add, data.frame(trend_data))

## rank of pressure categories
trend_data %>%
  group_by(impact) %>%
  summarize(mean = mean(trend)) %>%
  arrange(mean)

impact_name <- data.frame(impact = c("sst", "oa", "slr", 
                                                  "shipping",
                                                  "nutrient", "organic", "direct_human", "light",
                                                  "pel_hb", "pel_lb", "dem_dest", "dem_nondest_hb",
                                                  "dem_nondest_lb", "art_fish"),
                            impact_name = c("sst", "oa", "slr", 
                                              "shipping", 
                                              "nutrient pollution", "organic pollution", "direct human", "light pollution",
                                              "comm fish: pel hb", "comm fish: pel lb", "comm fish: dem dest", 
                                              "comm fish: dem nondest hb", "comm fish: dem nondest lb",
                                              "artisanal fishing"))

trend_data <- trend_data %>%
  mutate(impact = gsub("_trend", "", impact)) %>%
  left_join(impact_name, by = "impact")


trend_data$impact_name <- factor(trend_data$impact_name, levels=rev(impact_name$impact_name))
trend_data$name <- factor(trend_data$name, levels=total_trend$name)

library(beyonce)

myPalette <- c(beyonce_palette(18, 15, type=c("continuous"))[1:6],
               beyonce_palette(18, 25, type=c("continuous"))[15:18],
               beyonce_palette(18, 15, type=c("continuous"))[8],
               beyonce_palette(18, 20, type=c("continuous"))[16:19])
               

## some theme stuff to make the plot look nice
circle_theme <- theme(axis.line=element_blank(),
      axis.text.y=element_blank(),
      axis.ticks=element_blank(),
      axis.title.x=element_blank(),
      axis.title.y=element_blank(),
      legend.position="none",
      panel.background=element_blank(),
      panel.border=element_blank(),
      panel.grid.major=element_blank(),
      panel.grid.minor=element_blank(),
      plot.background=element_blank(),
      axis.text.x = element_blank())

# circle plot
# https://www.r-graph-gallery.com/299-circular-stacked-barplot/
p <- ggplot(data=trend_data, aes(x=name, y=trend, fill=impact_name)) +
  geom_bar(stat="identity") +
  geom_errorbar(aes(x = 1, ymin = -0.04, ymax=0.24), alpha=0)  +
 geom_text(data=total_trend, aes(x=name, y=trend_total+0.01, label=name, angle=angle, hjust=hjust, color=color), size=3, inherit.aes = FALSE) +
    scale_colour_identity() +
geom_segment(x = 1.5, y = 0, xend = dim(total_trend)[1]+0.5, yend = 0, colour = "black", alpha=1, size=0.3) +
  annotate("text", x = c(1,1,1), y = c(0, 0.05, 0.1), label = c(0, 0.05, 0.1), color="darkgrey", angle=-10, size=3) +
  scale_fill_manual(values=myPalette) + 
  coord_polar() +
  circle_theme 
 

p

ggsave('impacts/figures/trend/habitat_trend_circleplot.jpg', height=6, width=6, units=c("in"))


# getting the legend:
p <- ggplot(data=trend_data, aes(x=name, y=trend, fill=impact_name)) + 
  geom_bar(stat="identity") +
  geom_errorbar(aes(x = 1, ymin = -0.04, ymax=0.20), alpha=0)  +
  scale_fill_manual(values=myPalette) + 
  coord_polar() +
  geom_hline(yintercept=0, color="black") +
  geom_text(data=total_trend, aes(x=name, y=0.09, label=name, angle=angle+10, hjust=hjust), inherit.aes = FALSE) +
 theme(legend.title=element_blank())

p
leg <- cowplot::get_legend(p)
plot(leg)


ggsave('impacts/figures/trend/legend_circleplot.jpg', height=6, width=6, units=c("in"))


```


# region zooms

Project to lat long

```{r zoom_data}


## project raster to lat/long

# make a lat/long raster to copy
sim_rast <- raster(ncol=40000, nrow=20000)

chi <- raster(file.path(dir_M, "git-annex/impact_acceleration/impact/cumulative_impact/chi_2013.tif"))
projectRaster(chi, sim_rast, method="ngb", filename=file.path(dir_M, "git-annex/impact_acceleration/impact/figures/chi_lat_long_2013.tif"), progress="text", overwrite=TRUE)
chi <- raster(file.path(dir_M, "git-annex/impact_acceleration/impact/figures/chi_lat_long_2013.tif"))
plot(chi)

extremes <- raster(file.path(dir_M, 'git-annex/impact_acceleration/impact/figures/chi_slope_extremes.tif'))
projectRaster(extremes, sim_rast, method="ngb", filename=file.path(dir_M, "git-annex/impact_acceleration/impact/figures/extremes_lat_long.tif"), progress="text", overwrite=TRUE)

slope <- raster(file.path(dir_M, 'git-annex/impact_acceleration/impact/trend/chi_slope.tif'))
projectRaster(slope, sim_rast, method="ngb", filename=file.path(dir_M, "git-annex/impact_acceleration/impact/figures/slope_lat_long.tif"), progress="text", overwrite=TRUE)

slope_overlay <- raster(file.path(dir_M, 'git-annex/impact_acceleration/impact/trend/sig_overlay.tif'))
projectRaster(slope_overlay, sim_rast, method="ngb", filename=file.path(dir_M, "git-annex/impact_acceleration/impact/figures/sig_overlay_lat_long.tif"), progress="text", overwrite=TRUE)

## impacts
## (starting with 2013 files)
library(doParallel)
library(foreach)
library(parallel)



impact_files <- list.files(file.path(dir_M, "git-annex/impact_acceleration/impact/stressor_impact"), recursive = TRUE, 
                           pattern = ".tif", full="TRUE")

impacts_2013 <- grep("2013", impact_files, value=TRUE) 
impacts_2013 <- grep("uv|benthic_str", impacts_2013, value=TRUE, invert=TRUE)

registerDoParallel(8)

foreach(impact = impacts_2013,.packages="dplyr") %dopar%{
#for(impact in impacts_2013){ # impact = impacts_2013[2]
impact_name <- paste("latlong", basename(impact), sep="_")
  
imp_rast_mol <- raster(impact)
  
raster::projectRaster(imp_rast_mol, sim_rast, method="ngb", over=TRUE, filename=file.path(dir_M, sprintf("git-annex/impact_acceleration/impact/figures/impacts_lat_long/%s", impact_name)), overwrite=TRUE)
}

check <- raster("/home/shares/ohi/git-annex/impact_acceleration/impact/figures/impacts_lat_long/latlong_dem_dest_2013.tif")

### Trend data

trend_files <- list.files(file.path(dir_M, "git-annex/impact_acceleration/impact/trend/impacts"), 
                           pattern = ".tif", full="TRUE")

registerDoParallel(8)

foreach(trend = trend_files,.packages="dplyr") %dopar%{
#for(impact in impacts_2013){ # trend = trend_files[2]
trend_name <- paste("latlong", basename(trend), sep="_")
  
imp_rast_mol <- raster(trend)
  
raster::projectRaster(imp_rast_mol, sim_rast, method="ngb", over=TRUE, filename=file.path(dir_M, sprintf("git-annex/impact_acceleration/impact/figures/trend_lat_long/%s", trend_name)), overwrite=TRUE)
}

check <- raster("/home/shares/ohi/git-annex/impact_acceleration/impact/figures/trend_lat_long/latlong_dem_dest_2013.tif")



#regions_wgs <- st_transform(regions, 4326)
land_wgs <- spTransform(land, CRS("+init=epsg:4326"))

```

Figure out regions.
```{r zoom_regions}


chi_slope <- raster(file.path(dir_M, "git-annex/impact_acceleration/impact/figures/extremes_lat_long.tif"))

plot(chi_slope)
tmp <- zoom(chi_slope)


# high high
#africa <- extent(2700000, 2800000, 3770000, 3870000)

#high decrease
#uswest <- extent(-10000000, -9400000, 5400000, 6000000)
#kor <- extent(11200000, 11350000, 4100000, 4250000)

# low decrease
#ak <- extent(-159, -156.5, 56.5, 59)
#aus <- extent(126, 130, -33.0, -31.5)

# low fast
mad <- extent(45, 49, -17, -13)

###
# high decrease
ns <- extent(5.5, 11, 54, 59)

# high increase
gomaine <- extent(-72, -65, 41, 46)

# New zealand
nz <- extent(164, 174.5, -48, -38)

```

### region: chi plots

```{r zoom chi}
raster_breaks <- function(raster_data, crop_rgn, myBreaks, cols, filename){
png(sprintf("impacts/figures/rgn_maps/%s.png", filename), res=600, width=6, height=5, units="in")  
  
  par(mar=c(2,2,2,2)) # bottom, left, top, and right
  par(oma=c(0,0,0,0))

master <- raster_data
master_crop <- crop(master, crop_rgn)

plot(c(crop_rgn[1], crop_rgn[2]), c(crop_rgn[3], crop_rgn[4]), 
     type="n", xaxs="i", yaxs="i", axes=FALSE)
plot(master_crop, col=cols,  breaks=myBreaks, legend=FALSE, add=TRUE)
plot(land_wgs, add=TRUE, border="gray80", col="gray90", lwd=0.5)
box("plot", col="gray")

dev.off()
}


my_breaks <- c(-1, 0, 0.2, 0.4, 0.6, 0.8, 1, 1.25, 1.5, 1.75, 2, 2.5, 3, 3.5, 4.0, 5, 100) 
cols = rev(colorRampPalette(brewer.pal(11, 'Spectral'))(length(my_breaks)-1)) 
chi <- raster(file.path(dir_M, "git-annex/impact_acceleration/impact/figures/chi_lat_long_2013.tif"))


raster_breaks(raster_data=chi,  crop_rgn=nz,
              myBreaks=my_breaks, cols=cols, filename="chi_nz")
raster_breaks(raster_data=chi,  crop_rgn=gomaine,
              myBreaks=my_breaks, cols=cols, filename="chi_gomaine")
raster_breaks(raster_data=chi,  crop_rgn=ns,
              myBreaks=my_breaks, cols=cols, filename="chi_ns")
raster_breaks(raster_data=chi,  crop_rgn=mad,
              myBreaks=my_breaks, cols=cols, filename="chi_mad")


```

### region trend plots

```{r zoom trend}


region_plot <- function(raster_data, crop_rgn, saveFile, cols){
 #raster_data = slope
# region_extent = africa
# saveFile = "slope_africa"  
# cols = cols  
  
## crop region
master_crop <- crop(raster_data, crop_rgn)

png(sprintf("impacts/figures/rgn_maps/%s.png",  saveFile), 
    res=500, width=6, height=5, units="in")  
  par(mar=c(2,2,2,2)) # bottom, left, top, and right
  par(oma=c(0,0,0,0))

plot(c(crop_rgn[1], crop_rgn[2]), c(crop_rgn[3], crop_rgn[4]), 
     type="n", xaxs="i", yaxs="i", axes=FALSE)
plot(master_crop, col=cols,  
     breaks=seq(minValue(raster_data),
                  maxValue(raster_data), length.out=255), 
     legend=FALSE, add=TRUE)
plot(land_wgs, add=TRUE, border="gray80", col="gray90", lwd=0.5)
box("plot", col="gray")
  dev.off()
}

slope <- raster(file.path(dir_M, "git-annex/impact_acceleration/impact/figures/slope_lat_long.tif"))

cols = rev(colorRampPalette(c("#9E0142", "#AE1245", "#BE2449", "#CE364D", "#DA464C", "#E35449", "#EC6145", "#F47044", "#F7834D", "#F99756", "#FCAA5F",             
               "#FEF0A7", "#F3FAAD",  
              "#CAE99D", "#A6DBA4", "#7ECBA4", "#59B4AA", "#3B92B8", "#4470B1", "#5E4FA2"))(254)) # 

region_plot(raster_data=slope, crop_rgn=gomaine, saveFile="slope_gomaine", cols=cols)
region_plot(raster_data=slope, crop_rgn=ns, saveFile="slope_ns", cols=cols)
region_plot(raster_data=slope, crop_rgn=nz, saveFile="slope_nz", cols=cols)
region_plot(raster_data=slope, crop_rgn=mad, saveFile="slope_mad", cols=cols)

```

#### Slope/CHI combo plots
```{r zoom_extremes}


region_plot_extreme=function(raster_data=chi_slope, crop_rgn=ns, saveFile="x_ns", cols=cols,
                    breaks=my_breaks){
# raster_data = chi_slope
# region_extent = africa
# saveFile = "ext_africa"  
# cols = cols  
# breaks = my_breaks  
  
## crop region
master_crop <- crop(raster_data, crop_rgn)

png(sprintf("impacts/figures/rgn_maps/%s.png",  saveFile), 
    res=500, width=6, height=5, units="in")  
  par(mar=c(2,2,2,2)) # bottom, left, top, and right
  par(oma=c(0,0,0,0))

plot(c(crop_rgn[1], crop_rgn[2]), c(crop_rgn[3], crop_rgn[4]), 
     type="n", xaxs="i", yaxs="i", axes=FALSE)
plot(master_crop, col=cols,  
     breaks=my_breaks, 
     legend=FALSE, add=TRUE)
plot(land_wgs, add=TRUE, border="gray80", col="gray90", lwd=0.5)
box("plot", col="gray")
  dev.off()
}



chi_slope <- raster(file.path(dir_M, "git-annex/impact_acceleration/impact/figures/extremes_lat_long.tif"))

my_breaks <- c(-2, -1, 0, 1, 3)
cols = c("#018AC4", "#EBC915", "#550133", "#CF2154")


region_plot_extreme(raster_data=chi_slope, crop_rgn=gomaine, saveFile="x_gomaine", cols=cols,
                    breaks=my_breaks)
region_plot_extreme(raster_data=chi_slope, crop_rgn=ns, saveFile="x_ns", cols=cols,
                    breaks=my_breaks)
region_plot_extreme(raster_data=chi_slope, crop_rgn=nz, saveFile="x_nz", cols=cols,
                    breaks=my_breaks)
region_plot_extreme(raster_data=chi_slope, crop_rgn=mad, saveFile="x_mad", cols=cols,
                    breaks=my_breaks)

```


### combine images

```{r zoom_combine}
library(cowplot)
library(magick)

chi1 <- ggdraw() + draw_image("impacts/figures/rgn_maps/chi_gomaine.png")
chi2 <- ggdraw() + draw_image("impacts/figures/rgn_maps/chi_ns.png")
chi3 <- ggdraw() + draw_image("impacts/figures/rgn_maps/chi_mad.png")
chi4 <- ggdraw() + draw_image("impacts/figures/rgn_maps/chi_nz.png")

chi_plots <- plot_grid(chi1, chi2, chi3, chi4, nrow=1)
save_plot("impacts/figures/rgn_maps/chi.png", chi_plots, ncol=1, base_height=1, base_width=4)


trend1 <- ggdraw() + draw_image("impacts/figures/rgn_maps/slope_gomaine.png")
trend2 <- ggdraw() + draw_image("impacts/figures/rgn_maps/slope_ns.png")
trend3 <- ggdraw() + draw_image("impacts/figures/rgn_maps/slope_mad.png")
trend4 <- ggdraw() + draw_image("impacts/figures/rgn_maps/slope_nz.png")

trend_plots <- plot_grid(trend1, trend2, trend3, trend4, nrow=1)
save_plot("impacts/figures/rgn_maps/trend.png", trend_plots, ncol=1, base_height=1, base_width=4)

x1 <- ggdraw() + draw_image("impacts/figures/rgn_maps/x_gomaine.png")
x2 <- ggdraw() + draw_image("impacts/figures/rgn_maps/x_ns.png")
x3 <- ggdraw() + draw_image("impacts/figures/rgn_maps/x_mad.png")
x4 <- ggdraw() + draw_image("impacts/figures/rgn_maps/x_nz.png")

extreme_plots <- plot_grid(x1, x2, x3, x4, nrow=1)
save_plot("impacts/figures/rgn_maps/extremes.png", extreme_plots, ncol=1, base_height=1, base_width=4)

```



#### Make map showing regions

```{r zoom_global_loc}

legend.shrink <- 0.7
legend.width <- 0.6

# function to convert lat/long extents to moll points and add to plot
plotwgs <- function(x){ # x = aus
  long <- mean(c(x[1], x[2]))
  lat <- mean(c(x[3], x[4]))
  x_sp <- data.frame(lat=lat, long=long)
  coordinates(x_sp) <- c("long", "lat")
  proj4string(x_sp) <- CRS("+proj=longlat +ellps=WGS84 +datum=WGS84")
  x_mol <- spTransform(x_sp, CRS(proj4string(land)))
  plot(x_mol, col="red", add=TRUE, pch=16, cex=.7)
}


my_breaks <- c(-1, 0, 0.2, 0.4, 0.6, 0.8, 1, 1.25, 1.5, 1.75, 2, 2.5, 3, 3.5, 4.0, 5, 100) 
cols = rev(colorRampPalette(brewer.pal(11, 'Spectral'))(length(my_breaks)-1)) 
cols_trans = paste0(cols, "80")

p_rast <- raster(file.path(dir_M,  "git-annex/impact_acceleration/impact/cumulative_impact/chi_2013.tif"))

png("impacts/figures/rgn_maps/rgn_locations.png", 
    res=500, width=6, height=5, units="in")  
  par(mar=c(2,2,2,2)) # bottom, left, top, and right
  par(oma=c(0,0,0,0))
plot(p_rast, col=cols_trans, axes=FALSE, box=FALSE, breaks=my_breaks, legend=FALSE)
  
# add axis with fields package function:
def_breaks <- seq(0, length(my_breaks), length.out=length(my_breaks))
fields::image.plot(raster_data,  
           legend.only = TRUE, 
           legend.shrink=legend.shrink,
           legend.width=legend.width,
           col = cols,
           legend.lab="impact",
            breaks=def_breaks,
           lab.breaks=c(0, NA, 0.2, NA, 0.6, NA, 1, NA, 1.5, NA, 2, NA, 3, NA, 4.0, NA, ">5"),
           axis.args = list(cex.axis = 0.6))

plot(land, add=TRUE, border="gray80", col="gray90", lwd=0.5)

plotwgs(ns)
plotwgs(nz)
plotwgs(gomaine)
plotwgs(mad)

dev.off()

```


## Extremes map density plots

Extract the data for each response from the raster (high/low slope and chi) and save (5 parts):"impacts/figures/map_histogram_data/lat_extreme_N_%s.csv"
```{r}

extremes <- raster(file.path(dir_M, 'git-annex/impact_acceleration/impact/figures/chi_slope_extremes.tif'))

x_count <- function(x, raster=extremes){
  tmp <- getValues(raster, row=x)
  table(tmp)
}

start_time <- Sys.time()
tmp <- lapply(1:100, x_count)
end_time <- Sys.time()

end_time - start_time

extreme_N <- purrr::map_df(tmp, dplyr::bind_rows)



# ocean

ocean_count <- function(x, raster=ocean){
  tmp <- getValues(raster, row=x)
  table(tmp)
}
start_time <- Sys.time()
tmp2 <- lapply(1:100, ocean_count)
end_time <- Sys.time()

end_time - start_time

ocean_N <- purrr::map_df(tmp2, dplyr::bind_rows) 
names(ocean_N) <- "N_ocean_cells"

##
prop_cells <- cbind(extremes, ocean_N)


## Do in parallel
library(doParallel)
library(foreach)
library(parallel)

pullfrom <- data.frame(min = c(1, 3862, 7723, 11584, 15445), 
                               max = c(3861, 7722, 11583, 15444, 19305),
                       lab=c("part1", "part2", "part3", "part4", "part5"))

registerDoParallel(5)

foreach(run = 1:5) %dopar%{ # run=4

extremes <- raster::raster(file.path(dir_M, 'git-annex/impact_acceleration/impact/figures/chi_slope_extremes.tif'))

x_count <- function(x, raster=extremes){
  tmp <- raster::getValues(raster, row=x)
  table(tmp)
}

tmp <- lapply(pullfrom$min[run]:pullfrom$max[run], x_count)

extreme_N <- purrr::map_df(tmp, dplyr::bind_rows) %>%
  data.frame()

write.csv(extreme_N, sprintf("impacts/figures/map_histogram_data/lat_extreme_N_%s.csv", 
          pullfrom$lab[run]), row.names=FALSE)

}
```

### get ocean data
Compile the data extracted from the rasters.

```{r}
ocean_N <- rowSums(ocean, na.rm=TRUE)


lat_data <- bind_rows(
  read.csv("impacts/figures/map_histogram_data/lat_extreme_N_part1.csv"),
  read.csv("impacts/figures/map_histogram_data/lat_extreme_N_part2.csv"),
  read.csv("impacts/figures/map_histogram_data/lat_extreme_N_part3.csv"),
  read.csv("impacts/figures/map_histogram_data/lat_extreme_N_part4.csv"),
  read.csv("impacts/figures/map_histogram_data/lat_extreme_N_part5.csv")
)

lat_data <- cbind(lat_data, ocean_N)

names(lat_data) <- c("low_decrease", "high_fast", "high_decrease", "low_fast", "ocean_N")

lat_data[is.na(lat_data)] <- 0
lat_data$row_id = 1:dim(lat_data)[1]

# A dataset with proportion of cells:
# (currently not using this one)
lat_data_plot <-lat_data %>%
  mutate(prop_low_decrease = low_decrease/ocean_N) %>%
  mutate(prop_high_fast = high_fast/ocean_N) %>%
  mutate(prop_high_decrease = high_decrease/ocean_N) %>%
  mutate(prop_low_fast = low_fast/ocean_N)

library(zoo)
lat_data_plot2 <- gather(lat_data, "category", "cells", -ocean_N, -row_id) 

# grouping varible for combining lats
# this is my smoothing function
grouping_id <- lat_data %>%
  arrange(row_id) %>%
  mutate(group_id = gl(ceiling(nrow(.)/100), 100, nrow(.))) %>%
  dplyr::select(row_id, group_id)


lat_data_plot2_smooth <- lat_data_plot2 %>%
  left_join(grouping_id) %>%
  group_by(category, group_id) %>%
  summarize(row_id=unique(group_id),
            cells = sum(cells)) %>%
  ungroup() %>%
  mutate(category = as.factor(category)) %>%
  mutate(row_id = as.numeric(as.character(row_id)))
  


ggplot(lat_data_plot2_smooth) +
  geom_ribbon(aes(ymin = 0, ymax = cells, x=row_id, y=cells, group=category, fill=category)) +
  geom_line(aes(x=row_id, y=cells, group=category, color=category), size=0.3) +
  scale_x_reverse() +
  coord_flip() + 
#  facet_grid(. ~ category) +
scale_color_manual(values = c("#EBC915", "#018AC4", "#550133", "#CF2154"),
                    limits = c("low_fast", "low_decrease", "high_decrease", "high_fast")) +
  scale_fill_manual(values = c("#EBC915", "#018AC44D", "#33271BA6", "#CF21544D"),
                    limits = c("low_fast", "low_decrease", "high_decrease", "high_fast")) +
  theme(
  strip.background = element_blank(),
  strip.text = element_blank(),
  axis.title = element_blank(),
  axis.text  = element_blank(),
  axis.ticks =element_blank(),
  axis.line = element_blank(),
  panel.border = element_blank(),
  panel.grid = element_blank(),
  panel.spacing = unit(-3, "lines"),
  legend.position = "none")

ggsave("impacts/figures/map_histogram_data/latitudinal_extremes.png", width=3, height=5, units=c("in"), dpi=300)

```


