---
title: 'Trends in human impact'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '../../ohiprep_v2018/src/templates/ohi_hdr.html'
  pdf_document:
    toc: true
---


#Summary

This script calculates trend data for cumulative human impacts as well as individual pressures.

***  


## Setup
```{r setup, message=FALSE, warning=FALSE, error=FALSE}

knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',message = FALSE, warning = FALSE)

# setwd("trend")

library(tidyr)
library(dplyr)
library(raster)
library(RColorBrewer)
library(tidyverse)
library(rgdal)
library(doParallel)
library(foreach)
library(sf)
library(gstat)
library(stringr)

# load spatial files (ocean raster and regions shapefile)
source("https://raw.githubusercontent.com/OHI-Science/ohiprep_v2018/master/src/R/spatial_common.R")

cols = rev(colorRampPalette(brewer.pal(11, 'Spectral'))(255)) # rainbow color scheme


# paralell processing
cl<-makeCluster(5)
registerDoParallel(cl)


```

# Creating plots of individual random raster cells to explore change over time.
```{r ref point, eval=FALSE}

chi   <- list.files(file.path(dir_M,'git-annex/impact_acceleration/impact/cumulative_impact'), pattern='chi', full.names=TRUE)

chi_stack <- stack(chi)

samp_n <- 100

rand_samp <- sampleRandom(chi_stack, size=samp_n) %>%
  data.frame()

rand_samp$sample_id <- 1:samp_n

rand_samp_data <- rand_samp %>%
  gather("year", "chi", starts_with("chi")) %>%
  mutate(year = substr(year, 5, 8)) %>%
  mutate(year = as.numeric(year))
   
# check that everything went well
summary(rand_samp_data)        
table(rand_samp_data$sample_id)  

write.csv(rand_samp_data, "int/rand_samp.csv", row.names=FALSE)
````


```{r plot, fig.width=4, fig.height=4}


samp_n <- 100

rand_samp_data <- read.csv("int/rand_samp.csv")


for(i in 1:samp_n){ # i = 2
 
plot_chi <- ggplot(dplyr::filter(rand_samp_data, sample_id==i), aes(y = chi, x = year)) +
    geom_point(size = 2) +
   geom_line() + 
   stat_smooth(method=lm, se=FALSE, color="red")
  
plot(plot_chi)  
}


```

# Calculate slope, intercept

A regression model will be applied to each cell across all years of data, and the slope and intercept will be used to estimate 2013. 

```{r estimate_2013, eval=FALSE}

chi   <- list.files(file.path(dir_M,'git-annex/impact_acceleration/impact/cumulative_impact'), pattern='chi', full.names=TRUE)

chi_stack <- stack(chi)

years <-   as.numeric(substr(basename(chi), 5, 8))


# function to split raster
SplitRas <- function(raster,ppside,save,plot){
  h        <- ceiling(ncol(raster)/ppside)
  v        <- ceiling(nrow(raster)/ppside)
  agg      <- aggregate(raster[[1]],fact=c(h,v), fun=max)
  agg[]    <- 1:ncell(agg)
  agg_poly <- rasterToPolygons(agg)
  names(agg_poly) <- "polis"
  
  r_list <- list()
  for(i in 1:ncell(agg)){ # i=1
    e1          <- extent(agg_poly[agg_poly$polis==i,])
    crop(raster, e1, filename = file.path(dir_M, "git-annex/impact_acceleration/impact/cumulative_impact/tmp", paste("SplitRas",i,".R", sep="")), 
         overwrite=TRUE)
  }
  return(r_list)
}

# this will split raster into 9 blocks
SplitRas(raster=chi_stack, ppside=3, saveLoc=file.path(dir_M, "git-annex/impact_acceleration/impact/cumulative_impact/tmp"))

# check one to make sure all is well:
tmp <- stack(file.path(dir_M, 'git-annex/impact_acceleration/impact/cumulative_impact/tmp/SplitRas1'))

plot(tmp[[1]], col=cols)
          
# check to make sure I can actually combine them:
list2 <- list()
for(i in 1:9){ #i=1
  rx <- stack(file.path(dir_M, "git-annex/impact_acceleration/impact/cumulative_impact/tmp", 
                        sprintf("SplitRas%s", i)))
  list2[[i]] <- rx
}

list2$fun   <- max
list2$na.rm <- TRUE
rast.mosaic <- do.call(mosaic, list2)

plot(rast.mosaic, col=cols)
## YES!!!


# need to run and get error
# https://stat.ethz.ch/pipermail/r-sig-geo/2013-November/019856.html

funb <- function(x) {
if (all(is.na(x))) {
return(cbind(NA,NA))
}
m <- lm(x ~ years)
s  <- summary(m)
slope <- s$coefficients[2,1]
se <- s$coefficients[2,2]
return(cbind(slope, se))
}

# potential alternative
# ## much reduced regression model; [2] is to get the slope
# quickfun <- function(y) (invXtX %*% y)[2]
# x4 <- calc(s, quickfun) 


registerDoParallel(3)

foreach(i = 1:9) %dopar%{ #i=1

rx <- raster::stack(file.path(dir_M, "git-annex/impact_acceleration/impact/cumulative_impact/tmp", 
                        sprintf("SplitRas%s", i)))  
  
raster::calc(rx, fun=funb, filename=
               file.path(dir_M, "git-annex/impact_acceleration/impact/cumulative_impact/tmp",                      sprintf("SplitRas%s_lm", i)), overwrite=TRUE)
}


# combine tiles
list2 <- list()
for(i in 1:9){ #i=1
  rx <- stack(file.path(dir_M, "git-annex/impact_acceleration/impact/cumulative_impact/tmp", 
                        sprintf("SplitRas%s_lm", i)))
  list2[[i]] <- rx
}

list2$fun   <- max
list2$na.rm <- TRUE
rast.mosaic <- do.call(mosaic, list2)

writeRaster(rast.mosaic[[1]], file.path(dir_M, 'git-annex/impact_acceleration/trend/chi_slope.tif'))
writeRaster(rast.mosaic[[2]], file.path(dir_M, 'git-annex/impact_acceleration/trend/chi_se.tif'))

```


# View rasters
```{r}

se <- raster(file.path(dir_M, 'git-annex/impact_acceleration/trend/chi_se.tif'))
plot(se, col=cols, axes=FALSE)
slope <- raster(file.path(dir_M, 'git-annex/impact_acceleration/trend/chi_slope.tif'))
plot(slope, col=cols, axes=FALSE)

```

# Calculate upper and lower 95% CI rasters

```{r}
t_stat <- qt(0.975, 11-2)

se <- raster(file.path(dir_M, 'git-annex/impact_acceleration/trend/chi_se.tif'))
slope <- raster(file.path(dir_M, 'git-annex/impact_acceleration/trend/chi_slope.tif'))

# upper limit
s <- stack(slope, se)
overlay(s, fun=function(x,y){x + y*t_stat}, 
                    filename = file.path(dir_M, 'git-annex/impact_acceleration/trend/chi_upper_ci.tif'), overwrite=TRUE, progress="text")
upper_ci <- raster(file.path(dir_M, 'git-annex/impact_acceleration/trend/chi_upper_ci.tif'))

plot(upper_ci, col=cols, axes=FALSE)

# lower limit
overlay(s, fun=function(x,y) x - y*t_stat, 
                    filename = file.path(dir_M, 'git-annex/impact_acceleration/trend/chi_lower_ci.tif'), overwrite=TRUE, progress="text")
lower_ci <- raster(file.path(dir_M, 'git-annex/impact_acceleration/trend/chi_lower_ci.tif'))
plot(lower_ci, col=cols, axes=FALSE)

```

# Raster manipulation
```{r}

### replace negative values with -1 and positive values with +1
# upper
upper_ci <- raster(file.path(dir_M, 'git-annex/impact_acceleration/trend/chi_upper_ci.tif'))
reclassify(upper_ci, c(-Inf,0,-1, 0,Inf,1), 
           filename = file.path(dir_M, 'git-annex/impact_acceleration/trend/chi_upper_ci_convert.tif'), overwrite=TRUE, progress="text")

upper_ci_class <- raster(file.path(dir_M, 'git-annex/impact_acceleration/trend/chi_upper_ci_convert.tif'))
plot(upper_ci_class, col=cols, axes=FALSE)

# lower
lower_ci <- raster(file.path(dir_M, 'git-annex/impact_acceleration/trend/chi_lower_ci.tif'))
reclassify(lower_ci, c(-Inf,0,-1, 0,Inf,1), 
           filename = file.path(dir_M, 'git-annex/impact_acceleration/trend/chi_lower_ci_convert.tif'), overwrite=TRUE, progress="text")

lower_ci_class <- raster(file.path(dir_M, 'git-annex/impact_acceleration/trend/chi_lower_ci_convert.tif'))
plot(lower_ci_class, col=cols, axes=FALSE)

## combine
s <- stack(raster(file.path(dir_M, 'git-annex/impact_acceleration/trend/chi_upper_ci_convert.tif')),
           raster(file.path(dir_M, 'git-annex/impact_acceleration/trend/chi_lower_ci_convert.tif')))

overlay(s, fun=function(x,y){x + y}, 
                    filename = file.path(dir_M, 'git-annex/impact_acceleration/trend/trend_pattern.tif'), overwrite=TRUE, progress="text")
trend_pattern <- raster(file.path(dir_M, 'git-annex/impact_acceleration/trend/trend_pattern.tif'))
plot(trend_pattern, col=cols, axes=FALSE)

### trend overlay

df <- data.frame(id=c(2, 0, -2), v=c(NA, 1, NA))
subs(trend_pattern, df, filename = file.path(dir_M,  'git-annex/impact_acceleration/trend/sig_overlay.tif'), overwrite=TRUE, progress="text")

```


## Alternative trend calc
```{r}

chi   <- list.files(file.path(dir_M,'git-annex/impact_acceleration/impact/cumulative_impact'), pattern='chi', full.names=TRUE)

chi_stack <- stack(chi)

years <-   as.numeric(substr(basename(chi), 5, 8))

# add 1 for a model with an intercept
X <- cbind(1, years)
invXtX <- solve(t(X) %*% X) %*% t(X)

quickfun <- function(y) (invXtX %*% y)[2]

trend <- calc(chi_stack, quickfun, progress="text")
trend

# ## compare to other trend calc
# slope <- raster(file.path(dir_M, 'git-annex/impact_acceleration/trend/chi_slope.tif'))
# slope
# plot(slope, col=cols, axes=FALSE)
# plot(trend, col=cols, axes=FALSE)
# s <- stack(slope, trend)
# tmp <- overlay(s, fun=function(x,y){x - y}, progress="text")
# plot(tmp, col=cols, axes=FALSE)
# tmp
# # awesome!!!

```


# Trend for each impact

```{r}

impact <- c("sst", "oa", "slr", 
            "shipping",
            "nutrient", "organic", "direct_human", "light",
            "pel_hb", "pel_lb", "dem_dest", "dem_nondest_hb", "dem_nondest_lb", "art_fish")

impact_files   <- list.files(file.path(dir_M, "git-annex/impact_acceleration/impact/stressor_impact"), full=TRUE)

years <-   as.numeric(substr(basename(chi), 5, 8))

# add 1 for a model with an intercept
X <- cbind(1, years)
invXtX <- solve(t(X) %*% X) %*% t(X)
quickfun <- function(y) (invXtX %*% y)[2]

registerDoParallel(3)

foreach (file = impact) %dopar%{ # file = impact[1]
  
  impact_from_files <- grep(file, impact_files, value=TRUE)
  print(length(impact_from_files))
  impact_stack <- raster::stack(impact_from_files)
  raster::calc(impact_stack, quickfun, filename = file.path(dir_M, sprintf("git-annex/impact_acceleration/trend/impacts/%s_trend.tif", file)),
       overwrite=TRUE)
  
}



```