---
title: "Annual Artisanal Fisheries Catch"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ohi-northeast/src/templates/ohi_hdr.html'
pdf_document:
  toc: true
---

This script combines spatial catch data from [Watson (2017)](https://www.nature.com/articles/sdata201739) with species & gear specific categories defined in [`watson_gear_matching`](https://github.com/OHI-Science/impact_acceleration/blob/master/stressors/comm_fish/watson_gear_matching.Rmd) to create annual catch rasters for artisanal fishing.

## Overview
- Raw catch data is combined with `taxa_gear_types.csv` as defined in `watson_gear_matching.Rmd`
- Data is filtered for every year and catch type, rasterized to half degree cells, and saved as a GeoTIFF

```{r setup, message=F,warning=F}

knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',message = FALSE, warning = FALSE)

source('~/github/ohiprep/src/R/common.R')

library(seaaroundus)
library(raster)
library(readr)
library(dplyr)
library(data.table)
library(foreach)
library(doParallel)
library(RColorBrewer)

cols = rev(colorRampPalette(brewer.pal(11, 'Spectral'))(255)) # rainbow color scheme
```

#Load Data

Load annual catch data

```{r}

annual_files <- list.files(file.path(dir_M,'git-annex/impact_acceleration/stressors/comm_fish/int/catch_annual_data'),full.names=T)

```


#Map annual catch

## Get base raster

Using the [searoundus R package](https://github.com/ropensci/seaaroundus), we create a raster with cell values set equal to cell IDs in the data, which are defined in the column `Seq`.

```{r}

  #get template raster for SAUP data with cell values equal to CellID
  
   saup_cells <- getcells("POLYGON ((-180 90, 180 90, 180 -90, -180 -90, -180 90))")

   saup_rast <- raster(ncol=720, nrow=360)
   saup_rast[] <- saup_cells
   
   plot(saup_rast,col=cols,main = "SAUP Cell IDs")
```

## Map annual artisanal catch

Each annual stressor layer is created by aggregate all artisanal fishing catch per cell per year. The catch is identified by the column `SSF_CR` (small scale fishing rate)

```{r map,eval=F}

registerDoParallel(10) #register 10 cores

foreach(file = annual_files) %dopar% {
  
  yr <- substr(file,101,104)
  
  data <- readRDS(file)%>%
      rowwise()%>%
      group_by(Seq)%>%  #group by cell number
      summarise(cell_catch = sum(SSF_CR))%>%
    data.frame()#summing up all types of catch rates (remember these are in tonnes/km2 so we don't need to multiply by Ocean Area since we have 1km2 resolution)

  
  #rasterize catch by swapping cell ids with 
      raster::subs(saup_rast, data, by = 1, which = 2, subsWithNA=TRUE, filename = paste0(file.path(dir_M),'/git-annex/impact_acceleration/stressors/art_fish/int/catch_annual_rasters/art_fishing_',yr,'.tif'),overwrite=T)
}

```




