---
title: 'Stressor 2018: Ocean Acidification Data Prep '
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '../../../ohiprep_v2018/src/templates/ohi_hdr.html'
  pdf_document:
    toc: true
---

#Summary

NOTES: This follows the same methods as those used to generate the OHI pressure files (https://rawgit.com/OHI-Science/ohiprep/master/globalprep/prs_oa/v2017/create_oa_layer.html).  The only difference is that a modified version of the ocean mask is used to mask the data in the final step.    

***

#Data Source
**Reference**: [Feely et al.(2009)](https://darchive.mblwhoilibrary.org/bitstream/handle/1912/3180/22-4_feely.pdf?sequence=1&isAllowed=y)

**Downloaded**: July 19, 2017

**Description**:  Aragonite Saturation State  $\Omega_{arg}$

**Native data resolution**: 1 degree cells

**Time range**: 1880-1889 and 1958-2100, monthly data provided for each year. Future years are based on model projections for RCP 8.5. Previous years are hindcast/historical data.

**Format**:  NetCDF

  
**Notes about the data**:  

This data was shared with us by Ivan Lima from Woods Hole Institue for Oceanography in July 2017. The data came as NetCDFs with a resolution of about 1 degree. The data values are monthly average surface &#937; aragonite saturation state.


***

# Methods


##Setup  

```{r setup, message=FALSE, warning=FALSE, verbose=FALSE}

#set options for all chunks in code
knitr::opts_chunk$set(warning=FALSE, message=FALSE,fig.width=6, fig.height=6)

#libraries
library(raster)
library(ncdf4)
library(maps)
library(parallel)
library(foreach)
library(doParallel)
library(RColorBrewer)
library(stringr)

source("https://raw.githubusercontent.com/OHI-Science/ohiprep_v2018/master/src/R/spatial_common.R")

cols = rev(colorRampPalette(brewer.pal(11, 'Spectral'))(255)) # rainbow color scheme

# setwd("stressors/oa")

#define paths for the raw data and OA folder held on git-annex on our NCEAS server, Mazu

raw_dir    = file.path(dir_M, 'git-annex/globalprep/_raw_data')
oagit_dir  = file.path(dir_M, 'git-annex/globalprep/prs_oa')


# paralell processing
cl<-makeCluster(10)
registerDoParallel(cl)


```
  


## Resample & Land Mask

All pressure layers need to be resampled to 1km^2^ cell resolution. We have a template ocean raster with cells at this resolution that we use to resample all pressure layers. You won't see any difference between the plot above and this one since we are using the nearest neighbor method when resampling which maintains the original cell value for each of resampled cell.

```{r resample, eval=FALSE}

files <- list.files(file.path(oagit_dir,'v2017/int/annual_avg_moll_rescaled_int'),
                    full.names = TRUE)

# not going to bother with the full data range
files <- files[grep(paste0(paste0("_", 1980:2017), collapse="|"), files)]

foreach(file = files)%dopar% {
#for(file in files){ # file=files[1] # for some reason cores weren't all working for me.  Did it the long way!
  yr <- substr(file, nchar(file)-7, nchar(file)-4)
  cat(yr)
  
  tmp <- raster(file)%>%
  resample(ocean, method = 'ngb') %>%
  mask(ocean, filename = file.path(dir_M, sprintf("git-annex/impact_acceleration/stressors/oa/final/oa_%s_rescaled_mol.tif", yr)), overwrite=TRUE)
  
  # check = raster(file.path(dir_M, "git-annex/impact_acceleration/stressors/oa/final/oa_1980_rescaled_mol.tif"))
  # plot(check)
}

```


# Final Pressure Layer

```{r plot_final}
plot(raster(file.path(dir_M, "git-annex/impact_acceleration/stressors/oa/final/oa_2016_rescaled_mol.tif"), box=FALSE, axes=FALSE, col=cols, main='Final Ocean Acidification \nPressure Layer 2016'))

```


# Gap-filled cells

We want to create a raster layer that shows all cells that were gap-filled. Since they were the same cells interpolated across all years, we only need to create one raster.

```{r, eval=FALSE}

#Rescaled data before interpolation 
pre_int = raster(file.path(oagit_dir, 'v2017/int/annual_avg_moll_rescaled/oa_rescaled_2016.tif'))


pre_int[is.na(pre_int)] <- 999

resample(pre_int, ocean, method = "ngb") %>%
  mask(ocean, filename=file.path(dir_M, "git-annex/impact_acceleration/stressors/oa/int/oa_gf.tif"),
       overwrite = TRUE)

```

```{r plot_interp_cells}

plot(raster(file.path(dir_M, "git-annex/impact_acceleration/stressors/oa/int/oa_gf.tif")))

```


#Citation information  

Woods Hole Oceanographic Institution. 2014 update to data originally published in: Feely, R.A., S.C. Doney, and
S.R. Cooley. 2009. Ocean acidification: Present conditions and future changes in a high-CO2 world.
Oceanography 22(4):36–47

