---
title: "wastewater: calculate stressor and impact layers"
output: html_document
---

```{r}

library(RColorBrewer)

library(raster)
library(rgdal)

library(dplyr)
library(stringr)

library(parallel)
library(foreach)
library(doParallel)

library(here)

source("https://raw.githubusercontent.com/OHI-Science/ohiprep_v2019/gh-pages/workflow/R/common.R")
ohi_rasters()
region_data()

cols <- rev(colorRampPalette(brewer.pal(11, 'Spectral'))(255)) # rainbow color scheme
cols <- cols[150:255]

```


Rescale data to values 0-1 based on 99.99th quantile of data.

```{r}

ww_mol <- raster::raster(file.path(dir_M, "git-annex/land-based/wastewater/output/global_effluent_2015_all_raw_mol_nonegs_clean.tif"))

## find and save 99.99th quantile across raster
  quant9999 <-  quantile(ww_mol, .9999)
  
quant_df <- data.frame(stressor = "wastewater", quant = quant9999)

write.csv(quant_df, 
          here("projects/wastewater/quant.csv"),
                               row.names = FALSE)

## fertilizer/nutrient
quantiles <- read.csv(here("projects/wastewater/quant.csv"))

ref_point <- quantiles$quant[quantiles$stressor=="wastewater"]

raster::calc(ww_log, fun=function(x){ifelse(x>ref_point, 1, x/ref_point)},
       filename = file.path(dir_M, "git-annex/land-based/wastewater/output/no_log_data/stressor/global_effluent_2015_mol_rescaled.tif"), 
       overwrite=TRUE, progress="text")


ww <- raster::raster(file.path(dir_M, "git-annex/land-based/wastewater/output/no_log_data/stressor/global_effluent_2015_mol_rescaled.tif"))
plot(ww)
zoom(ww)

```

Create vulnerability x habitat x stressor combos of data.  I will use the nutrient pollution vulnerabilities.

```{r}

## preparing the data

vulnerability <- read.csv(here("vulnerability_weighting_matrix.csv")) %>%
  filter(pressure != "pressure")


# Habitats: 21 habitats
# make sure habitat rasters from vulnerability matrix are available
# Habitats in vulnerability matrix:
names(vulnerability)
# List of habitat rasters
habs <- list.files(file.path(dir_M, "git-annex/impact_acceleration/habitats"))
habs <- habs[-(grep(".vat.dbf|.xml|.ovr", habs))]

# The following should be zero
# The habitat rasters should all be in the vulnerability matrix
setdiff(habs, paste0(names(vulnerability), '.tif'))

# there are a couple habitats in the vulnerability table that we do not have raster habitat data for:
# "vent"        "Soft.Canyon" "Hard.Canyon"
xtra_vul_habs <- setdiff(paste0(names(vulnerability), '.tif'), habs) 
xtra_vul_habs <- xtra_vul_habs[-which(xtra_vul_habs=="pressure.tif")]
xtra_vul_habs <- gsub(".tif", "", xtra_vul_habs)
xtra_vul_habs

vulnerability_clean <- vulnerability %>%
  dplyr::select(-one_of(xtra_vul_habs)) %>%  # cut habitats we do not have raster data for
  dplyr::filter(pressure == "nutrient") %>%  # cut stressors we do not have raster data for
  tidyr::gather("habitat", "vulnerability", -1) %>%
  dplyr::mutate(stress_loc = NA) %>%
  dplyr::mutate(output = NA)

```


```{r}

registerDoParallel(5)
foreach(row = 1:dim(vulnerability_clean)[1], .packages="dplyr") %dopar%{ # row=1

  combo_data <- vulnerability_clean[row, ]
  
  #obtain stressor raster location
  stress_rast <- file.path(dir_M, "git-annex/land-based/wastewater/output/no_log_data/stressor/global_effluent_2015_mol_rescaled.tif")
  
  #obtain habitat raster location
  hab_rast <- sprintf("/home/shares/ohi/git-annex/impact_acceleration/habitats/%s.tif", combo_data$habitat)
  
  #vulnerability
  vuln <- as.numeric(combo_data$vulnerability)
  
  # multiply stressor * habitat * vulnerability:
  combo_stack <- raster::stack(stress_rast, hab_rast)
  raster::overlay(combo_stack, fun=function(x,y){(x*y*vuln)}, 
          filename = file.path(dir_M, sprintf("git-annex/land-based/wastewater/output/no_log_data/hab_stressor_combo/global-effluent__%s__2015.tif", 
                               combo_data$habitat)), overwrite=TRUE)
  }


```

Average the combos to get stressor impact

```{r}

hab_num <- raster(file.path(dir_M, "git-annex/impact_acceleration/habitat_number/habitat_num.tif"))

# should be 21...and there is!
combos <- list.files(file.path(dir_M, "git-annex/land-based/wastewater/output/no_log_data/hab_stressor_combo"), full=TRUE)


stress_stack <- raster::stack(combos)
  
raster::calc(stress_stack, fun=sum, na.rm=TRUE,
      filename=file.path(dir_M, "git-annex/land-based/wastewater/output/no_log_data/tmp/summed_raster_global-effluent_2015.tif"), 
      overwrite=TRUE, progress="text")

summed_rast <- raster::raster(file.path(dir_M, "git-annex/land-based/wastewater/output/no_log_data/tmp/summed_raster_global-effluent_2015.tif"))
  
  raster::overlay(summed_rast, hab_num, fun=function(x,y){x/y}, 
                  filename = file.path(dir_M, "git-annex/land-based/wastewater/output/no_log_data/tmp/summed_raster_global-effluent_2015_rescaled.tif"), overwrite=TRUE, progress="text")


```

clean raster created above. 
```{r}

## Ocean mask for each layer created above.
```{r}

 rescaled <- raster::raster(file.path(dir_M, "git-annex/land-based/wastewater/output/no_log_data/tmp/summed_raster_global-effluent_2015_rescaled.tif"))

 # mask data and save
 raster::mask(rescaled, ocean,
                 filename=file.path(dir_M, "git-annex/land-based/wastewater/output/no_log_data/stressor_impact/global-effluent_2015.tif"),
                 overwrite=TRUE, progress="text")
 

 final_ww <- raster::raster(file.path(dir_M, "git-annex/land-based/wastewater/output/no_log_data/stressor_impact/global-effluent_2015.tif"))
 plot(final_ww, col=c("white", cols))
 zoom(final_ww, col=c("white", cols))
 
```

Extract data

```{r}
# r <- raster(ncol=360, nrow=180, xmx=180, xmn=-180, ymn=-90, ymx=90)
# plot(area(r))

ww <- list.files(file.path(dir_M, "git-annex/land-based/wastewater/output/no_log_data/stressor_impact"), full=TRUE)
ww <- raster(ww)
#chi_stack <- stack(chi)

## full eez
ww_eez <- zonal(ww, zones, fun="mean", progress="text", na.rm=TRUE)


## 3nm
rgns_3nm <- raster::raster(file.path(dir_M, "git-annex/globalprep/spatial/v2018/rgns_3nm_offshore_mol.tif"))
plot(rgns_3nm)
ww_3nm <- zonal(ww, rgns_3nm, fun="mean", progress="text", na.rm=TRUE)

regions <- rgns_all %>%
  select(type_w_ant, rgn_id = rgn_ant_id, rgn_name) %>%
  filter(type_w_ant %in% c("eez", "eez-ccamlr", "fao"))

##
ww_data_df <- data.frame(ww_eez) %>%
  dplyr::select(zone, eez_avg=mean) %>%
  dplyr::left_join(data.frame(ww_3nm)) %>%
  dplyr::select(rgn_id=zone, eez_avg, nm3_avg=mean) %>%
  dplyr::left_join(regions, by="rgn_id")

#should have 254 regions, but only going to look at eezs, so this will end up as 220 regions


ww_data_df <- ww_data_df %>%
  dplyr::filter(type_w_ant == "eez") %>%
  dplyr::select(rgn_id, rgn_name, eez_avg, nm3_avg) %>%
  dplyr::arrange(nm3_avg)


```


Compare data to other pressures.
(realized I never extracted these particular data)
```{r}

## all impacts 2013 data
impact2013 <- list.files(file.path(dir_M, "git-annex/impact_acceleration/impact/stressor_impact"), pattern="2013", full.names=TRUE)

impact2013 <- grep("uv_2013|str_2013", impact2013, invert=TRUE, value=TRUE)

impact_stack <- stack(impact2013)

impacts_3nm <- zonal(impact_stack, rgns_3nm, fun="mean", progress="text", na.rm=TRUE)
write.csv(impacts_3nm, here("paper/zonal_data_eez/eez_3nm_2013_impacts.csv"), row.names=FALSE)


impacts_3nm <- read.csv(here("paper/zonal_data_eez/eez_3nm_2013_impacts.csv"))
impacts_all <- data.frame(impacts_3nm) %>%
  dplyr::rename(rgn_id=zone) %>%
  dplyr::left_join(ww_data_df)

## add in chi
chi <- read.csv(here("paper/zonal_data_eez/eez_3nm_chi.csv")) %>%
  dplyr::filter(year == 2013) %>%
  dplyr::select(rgn_id, chi_2103=value)

impacts_all <- impacts_all %>%
  left_join(chi, by="rgn_id")

write.csv(impacts_all, here("projects/wastewater/eez_3nm.csv"), row.names=FALSE)

```


Compare the data.
```{r}

impacts <- read.csv(here("projects/wastewater/eez_3nm.csv"))

library(ggplot2)


ggplot(impacts, aes(y=nm3_avg, x=nutrient_2013)) +
  geom_point(alpha=0.3, size=2.5) +
  theme_bw() + 
  labs(y="wastewater: nutrient impact", x="fertilizer: nutrient impact") +
  geom_abline(slope=1, intercept=0, color="red")

```