---
title: "Calc_stressor_habitat"
output: html_document
---

This script creates stressor x habitat x vulnerability rasters for each pressure/habitat/year.
I'm adding in the additional years of SST.

## Loading packages

```{r}
#libraries
library(raster)
library(RColorBrewer)
library(sf)
library(dplyr)
library(doParallel)
library(foreach)
library(parallel)
library(here)

source("https://raw.githubusercontent.com/OHI-Science/ohiprep_v2018/gh-pages/src/R/spatial_common.R")


```

## Reading in data files

This code also checks that the vulnerability matrix is harmonious with the habitat and stressor files.
```{r}
years <- 2014:2017

years_subset <- paste(years, collapse="|")

# Vulnerability matrix
vulnerability <- read.csv(here("vulnerability_weighting_matrix.csv")) %>%
  filter(pressure != "pressure")


# Habitats: 21 habitats
# make sure habitat rasters from vulnerability matrix are available
# Habitats in vulnerability matrix:
names(vulnerability)
# List of habitat rasters
habs <- list.files(file.path(dir_M, "git-annex/impact_acceleration/habitats"))
habs <- habs[-(grep(".vat.dbf|.xml|.ovr", habs))]

# The following should be zero
# The habitat rasters should all be in the vulnerability matrix
setdiff(habs, paste0(names(vulnerability), '.tif'))

# there are a couple habitats in the vulnerability table that we do not have raster habitat data for:
# "vent"        "Soft.Canyon" "Hard.Canyon"
xtra_vul_habs <- setdiff(paste0(names(vulnerability), '.tif'), habs) 
xtra_vul_habs <- xtra_vul_habs[-which(xtra_vul_habs=="pressure.tif")]
xtra_vul_habs <- gsub(".tif", "", xtra_vul_habs)
xtra_vul_habs


####### Stressors
# List of stressor files we have for 16 stressors across all years
stress_files <- list.files(file.path(dir_M, "git-annex/impact_acceleration/stressors"), recursive = TRUE, 
           pattern = ".tif", full="TRUE")
stress_files <- grep("/final/", stress_files, value=TRUE)
stress_files <- grep("/archive/", stress_files, value=TRUE, invert=TRUE)
stress_files <- grep("sst", stress_files, value=TRUE)

#filter to relevant years 
# (this includes uv and benthic structures...which are not included due to not having enough years data)
stress_files <- grep(years_subset, stress_files, value=TRUE)

# make sure all stressor files are available
files <- basename(stress_files)
files <- gsub("_rescaled_mol.tif", "", files)
files <- gsub('_[[:digit:]]+', '', files)
files <- unique(files)

# list of stressors
files

#should be zero: all stressors rasters should be in the vulnerability matrix
setdiff(files, vulnerability$pressure)

# only sst pressure
xtra_vul_stressor <- setdiff(vulnerability$pressure, files)
xtra_vul_stressor

######### Clean vulnerability table
vulnerability_clean <- vulnerability %>%
  dplyr::select(-one_of(xtra_vul_habs)) %>%  # cut habitats we do not have raster data for
  dplyr::filter(!(pressure %in% xtra_vul_stressor)) %>%  # cut stressors we do not have raster data for
  tidyr::gather("habitat", "vulnerability", -1) %>%
  merge(data.frame(years), by=NULL) %>%
  dplyr::mutate(stress_loc = NA) %>%
  mutate(output = NA)
```


## Create raster of each habitat x stressor x vulnerability combo

```{r}
stress_files <- list.files(file.path(dir_M, "git-annex/impact_acceleration/stressors"), recursive = TRUE, 
                           pattern = ".tif", full="TRUE")
stress_files <- grep("/final/", stress_files, value=TRUE)
stress_files <- grep("/archive/", stress_files, value=TRUE, invert=TRUE)

#filter to relevant years
stress_files <- grep(years_subset, stress_files, value=TRUE)
stress_files <- grep("sst", stress_files, value=TRUE)

registerDoParallel(5)
foreach(row = 1:dim(vulnerability_clean)[1], .packages="dplyr") %dopar%{ # row=1

  combo_data <- vulnerability_clean[row, ]
  
  #obtain stressor raster location
  pxy <- paste(combo_data$pressure, combo_data$years, sep="_")
  stress_rast <- grep(pxy, stress_files, value=TRUE)
  
  #obtain habitat raster location
  hab_rast <- sprintf("/home/shares/ohi/git-annex/impact_acceleration/habitats/%s.tif", combo_data$habitat)
  
  #vulnerability
  vuln <- as.numeric(combo_data$vulnerability)
  
  # multiply stressor * habitat * vulnerability:
  combo_stack <- raster::stack(stress_rast, hab_rast)
  raster::overlay(combo_stack, fun=function(x,y){(x*y*vuln)}, 
          filename = file.path(dir_M, sprintf("git-annex/impact_acceleration/no_sst/hab_stressor_combo/%s__%s__%s.tif", 
                               combo_data$pressure, combo_data$habitat, combo_data$years)), overwrite=TRUE)
  }

```

# check to see if all combos were created

```{r}
combos_obs <- list.files(file.path(dir_M, "git-annex/impact_acceleration/no_sst/hab_stressor_combo"))
length(combos_obs)

dim(vulnerability_clean)

```
