---
title: "Calc_impacts"
output: html_document
---

Caculate impact for each stressor by summing the stressor x habitat x vulnerability rasters for each stressor, and then dividing by the number of habitats.  This is only for the additional years of SST data.

## Loading packages

```{r}
#libraries
library(raster)
library(RColorBrewer)
library(sf)
library(dplyr)
library(doParallel)
library(foreach)
library(parallel)

source("https://raw.githubusercontent.com/OHI-Science/ohiprep_v2018/gh-pages/src/R/spatial_common.R")


```


## Check stressor data
All stressor/years should have 21 rasters, one for each habitat.

```{r}

hab_num <- raster(file.path(dir_M, "git-annex/impact_acceleration/habitat_number/habitat_num.tif"))

combos <- list.files(file.path(dir_M, "git-annex/impact_acceleration/no_sst/hab_stressor_combo"), full=TRUE)

length(combos)/length(2014:2017) # should equal 21

```

## Create raster that sums all habitat x impacts for a stressor/year
For each stressor and year, identify the stressor x habitat x vulnerability rasters.  Sum these and then divide by the number of habitats.  

```{r}
hab_num <- raster(file.path(dir_M, "git-annex/impact_acceleration/habitat_number/habitat_num.tif"))

combos <- list.files(file.path(dir_M, "git-annex/impact_acceleration/no_sst/hab_stressor_combo"), full=TRUE)

registerDoParallel(4)
foreach(years = 2014:2017) %dopar%{ # years=2014

  year <- years
  tmp <- grep(year, combos, value=TRUE)
  
  stress_stack <- raster::stack(tmp)
  
  raster::calc(stress_stack, fun=sum, na.rm=TRUE,
      filename=file.path(dir_M, sprintf("git-annex/impact_acceleration/no_sst/tmp/summed_raster_sst_%s.tif", year)), 
      overwrite=TRUE)
  
  summed_rast <- raster::raster(file.path(dir_M, sprintf("git-annex/impact_acceleration/no_sst/tmp/summed_raster_sst_%s.tif", year)))
  
  raster::overlay(summed_rast, hab_num, fun=function(x,y){x/y}, 
                  filename = file.path(dir_M, sprintf("git-annex/impact_acceleration/no_sst/tmp/summed_raster_sst_%s_rescaled.tif", year)), overwrite=TRUE)

}

```


## Ocean mask for each layer created above.
```{r}
rescaled_data <- list.files(file.path(dir_M, sprintf("git-annex/impact_acceleration/no_sst/tmp")), full=TRUE)
rescaled_data <- rescaled_data[grep("rescaled", rescaled_data)]

for(rescale in rescaled_data) { # rescale=rescaled_data[1]
  
 rescaled <- raster::raster(rescale)

 # get stressor and year from file name
stressor <- basename(rescale)
 stressor <- gsub("_rescaled.tif", "", stressor)
 stressor <- gsub("summed_raster_", "", stressor)
 stressor <- data.frame(combo=stressor)
 
 combo_data <- stressor %>% 
   tidyr::separate(combo, c("stressor", "year"), sep=-4) %>%
   dplyr::mutate(stressor = substr(stressor, 1, nchar(stressor)-1))
 
 # mask data and save
 raster::mask(rescaled, ocean,
                 filename=file.path(dir_M, sprintf("git-annex/impact_acceleration/no_sst/stressor_impact/%s_%s.tif", 
                                                   combo_data$stressor, combo_data$year)),
                 overwrite=TRUE, progress="text")
 
}

```
