---
title: "Data statistics"
output: html_document
---

The following script extracts data from maps and other sources to provide information used in the paper.


```{r}
library(raster)
library(sf)
library(dplyr)
library(stringr)
library(doParallel)
library(foreach)
library(parallel)
library(tidyr)

source("https://raw.githubusercontent.com/OHI-Science/ohiprep_v2018/gh-pages/src/R/spatial_common.R")

regions_shape <- as(regions, "Spatial")
land <- regions_shape[regions_shape$rgn_type %in% c("land", "land-disputed", "land-noeez", "eez-inland"), ]


```


## Trend patterns
Nearly three-quarters (XX%) of the global ocean experienced significantly increasing CHI over the 11 year period from 2003-2013, with only XX% significantly decreasing (Fig. 1A). 

The trend_pattern.tif is obtained by getting the upper and lower 95% confidence interval for the slope estimate.  For the upper and lower ci layers, negative values are converted to -1 and postitive values are converted to 1.  These are then summed.  When both are negative (sum to -2) this indicates a significant negative trend.  When both are positive (sum to 2) this indicates a significant positive trend.  When the trend is not statistically significant, the values sum to 0.  
```{r}

trend <- raster::raster(file.path(dir_M, "git-annex/impact_acceleration/impact/trend/trend_pattern.tif"))

plot(trend)
click(trend) # make sure land and border are NA!

freq(trend)
# [1,]    -2  19341841
# [2,]     0 149756597
# [3,]     2 247092363
# [4,]    NA 329175249

# also need the number of ocean cells
plot(ocean)

freq(ocean)
# [1,]     1 416190801
# [2,]    NA 329175249


## final values
# increasing
247092363/416190801

# decreasing
19341841/416190801

```

Nearly a third of the ocean (XX%) experienced relatively high rates of increase (linear regression slope >0.1 yr-1).

Reclassify: intervals closed (i.e., a closed interval includes the value) on the right (and open on the left)

```{r}

trend <- raster::raster(file.path(dir_M, "git-annex/impact_acceleration/impact/trend/chi_slope.tif"))

# slope = 0.0
reclassify(trend, c(-Inf,0, -1, 0,Inf,1), 
           filename = file.path(dir_M, 'git-annex/impact_acceleration/paper/tmp/slope_gt_0.tif'), overwrite=TRUE, progress="text")

trend_gt_zero <- raster(file.path(dir_M, 'git-annex/impact_acceleration/paper/tmp/slope_gt_0.tif'))
freq(trend_gt_zero)

# [1,]    -1  77581546
# [2,]     1 338609255
# [3,]    NA 329175249
338609255/416190801


# slope = 0.05
reclassify(trend, c(-Inf,0.05, NA, 0.05,Inf,1), 
           filename = file.path(dir_M, 'git-annex/impact_acceleration/paper/tmp/slope_gt_point05.tif'), overwrite=TRUE, progress="text")

trend_gt_point05 <- raster(file.path(dir_M, 'git-annex/impact_acceleration/paper/tmp/slope_gt_point05.tif'))
freq(trend_gt_point05)

# [1,]     1 166531994
# [2,]    NA 578834056
166531994/416190801

# slope = 0.1
reclassify(trend, c(-Inf,0.1, NA, 0.1,Inf,1), 
           filename = file.path(dir_M, 'git-annex/impact_acceleration/paper/tmp/slope_gt_point1.tif'), overwrite=TRUE, progress="text")

trend_gt_point1 <- raster(file.path(dir_M, 'git-annex/impact_acceleration/paper/tmp/slope_gt_point1.tif'))
freq(trend_gt_point1)

#      value     count
# [1,]     1  63228284
# [2,]    NA 682137766

63228284/416190801

# slope = 0.15
reclassify(trend, c(-Inf,0.15, NA, 0.15,Inf,1), 
           filename = file.path(dir_M, 'git-annex/impact_acceleration/paper/tmp/slope_gt_point15.tif'), overwrite=TRUE, progress="text")

trend_gt_point15 <- raster(file.path(dir_M, 'git-annex/impact_acceleration/paper/tmp/slope_gt_point15.tif'))
freq(trend_gt_point15)

# [1,]     1  15149943
# [2,]    NA 730216107

15149943/416190801

par(mar=c(2,2,2,2)) # bottom, left, top, and right
par(oma=c(0,0,0,0))

plot(trend_gt_point15, col="red", main="trend >0.15", legend=FALSE, box=FALSE, axes=FALSE)
#zoom(trend_gt_point15, col="red")
plot(land, add=TRUE, border="gray80", col="gray90", lwd=0.2)



```


Are coastal areas more likely to have fastest pace of change?  No. 

```{r}

rgns_3nm <- raster::raster(file.path(dir_M, "git-annex/globalprep/spatial/v2018/rgns_3nm_offshore_mol.tif"))

plot(rgns_3nm)

# convert regions to 1
rgns_3nm_global <- rgns_3nm
rgns_3nm_global[!is.na(rgns_3nm_global)] <- 1
freq(rgns_3nm_global)
#      value     count
# [1,]     1   5396062
# [2,]    NA 739969988

s <- stack(rgns_3nm_global,
           raster(file.path(dir_M, 'git-annex/impact_acceleration/paper/tmp/slope_gt_point15.tif')))

# sum_narm = function(x,...){sum(x, na.rm=TRUE)}
# overlay(s, fun=sum_narm, 
#                     filename = file.path(dir_M, 'git-annex/impact_acceleration/impact/trend/slope_gt_point2_plus_3nm.tif'), overwrite=TRUE, progress="text")

overlay(s, fun=sum, filename = file.path(dir_M, 'git-annex/impact_acceleration/impact/trend/slope_gt_point15_plus_3nm.tif'), overwrite=TRUE, progress="text")

combined <- raster(file.path(dir_M, 'git-annex/impact_acceleration/impact/trend/slope_gt_point15_plus_3nm.tif'))

plot(combined)

# the 1s are extreme trend within 3nm of shoreline
freq(combined)
# N = 95955


extreme_point2_count <- 15149943
extreme_3nm <- 95955
extreme_3nm/extreme_point2_count

nm3_area <- 5396062
ocean_area <- 416190801

nm3_area/ocean_area

```

## CHI quantiles
```{r}
files <- list.files("paper/quantiles", full=TRUE, pattern="chi")

datalist = lapply(files, function(x){read.csv(file=x, header=TRUE)})
quants <- bind_rows(datalist)

quants$year <- as.numeric(str_sub(quants$impact,-8,-5))
quants <- quants %>%
  select(quants, quant, year) %>%
  spread(quant, quants)

write.csv(quants, "paper/tables/CHI_quantiles.csv", row.names=FALSE)

```  


## country summaries
```{r}

chi <- read.csv("paper/zonal_data_eez/eez_chi_3nm_trend.csv")
head(chi)
dim(chi)
dim(chi[chi$value>0,])
190/220
dim(chi[chi$value<=0,])
30/220
dim(chi[chi$value>0.1,])
19/220

chi <- chi %>%
  arrange(value) %>%
  select(region = rgn_name, chi_trend=value)
write.csv(chi, "paper/tables/trend_3nm_rgns.csv", row.names=FALSE)

```

Closer look at individual impacts

## get overview of each pressure

```{r}
trends <- read.csv("paper/zonal_data_eez/eez_3nm_impacts_trend.csv")
chi <- read.csv("paper/zonal_data_eez/eez_chi_3nm_trend.csv")
  
trends_table <- bind_rows(trends, chi) %>%  
  mutate(trend = ifelse(value >0, 1, -1)) %>%
  group_by(pressure) %>%
  summarize(positive_trend = sum(trend==1),
            negative_trend = sum(trend==-1)) %>%
  rowwise() %>%
  mutate(countries_positive_trend = positive_trend/220)

# merge with other values
nm3global <- read.csv("paper/zonal_data_eez/global_3nm_impacts_trend.csv") %>%
  dplyr::select(pressure, global_3nm_trend = value) %>%
  mutate(global_3nm_trend = round(global_3nm_trend, 5))

nm3chi <- data.frame(pressure="chi_trend", global_3nm_trend = sum(nm3global$global_3nm_trend))
nm3global <- bind_rows(nm3global, nm3chi)

global <- read.csv("paper/zonal_data_eez/global_impacts_trend.csv") %>%
  dplyr::select(pressure=impact, global_trend = value) %>%
  mutate(global_trend = round(global_trend, 5)) %>%
  mutate(pressure = gsub("_trend", "", pressure)) %>%
  mutate(pressure= gsub("_slope", "_trend", pressure))

## join together
trends_table <- left_join(trends_table, nm3global) %>%
  left_join(global) %>%
  arrange(global_3nm_trend)

write.csv(trends_table, "paper/tables/trend_summary.csv", row.names=FALSE)

.03245/0.04234 # global: sst
0.00675/0.04234 # global: oa
0.0146/0.03724 #coastal:sst
0.01531/0.03724 # coastal: slr
(0.0146 + 0.01531)/0.03724

```

### Fishing
```{r}

all <- read.csv("paper/zonal_data_eez/eez_3nm_impacts_trend.csv") 
table(all$pressure)

fish <- read.csv("paper/zonal_data_eez/eez_3nm_impacts_trend.csv") %>%
  filter(pressure %in% c("dem_dest", "dem_nondest_hb", "dem_nondest_lb", "pel_hb", "pel_lb")) %>%
  mutate(decreasing = ifelse(value <=0, 1, 0)) %>%
  group_by(rgn_id, rgn_name) %>%
  summarize(total_decreasing = sum(decreasing)) %>%
    arrange(-total_decreasing) %>% 
  data.frame()

table(fish$total_decreasing)
(43+51+23)/220

fish <- read.csv("paper/zonal_data_eez/eez_3nm_impacts_trend.csv") %>%
  filter(pressure %in% c("dem_dest", "dem_nondest_hb", "dem_nondest_lb", "pel_hb", "pel_lb")) %>%
    arrange(value) %>% 
  data.frame()

fish %>%
  group_by(rgn_name) %>%
  summarize(sum_trend = sum(value)) %>%
  arrange(sum_trend)
```



#Trend with climate change pressures removed

```{r}
all <- read.csv("paper/zonal_data_eez/eez_3nm_impacts_trend.csv") 

no_cc <- all %>%
  filter(!(pressure %in% c("sst", "slr"))) %>%
  group_by(rgn_id, rgn_name) %>%
  summarize(chi_trend_no_cc = sum(value))

dim(no_cc[no_cc$chi_trend_no_cc >=0, ])
#N=184/220

chi <- read.csv("paper/zonal_data_eez/eez_chi_3nm_trend.csv")
compare <- left_join(chi, no_cc)

plot(chi_trend_no_cc ~ value, data=compare, xlab="chi")
abline(0,1, col="red")

# No SST
no_sst <- all %>%
  filter(!(pressure %in% c("sst"))) %>%
  group_by(rgn_id, rgn_name) %>%
  summarize(chi_trend_no_cc = sum(value))

dim(no_sst[no_sst$chi_trend_no_cc >=0, ])
#N=202/220

# All pressures
all_stressors <- all %>%
  group_by(rgn_id, rgn_name) %>%
  summarize(chi_trend_no_cc = sum(value))

dim(all_stressors[all_stressors$chi_trend_no_cc >= 0, ])
#N=190/220


```


## Global trend patterns with no SST

```{r}
trend <- raster::raster(file.path(dir_M, "git-annex/impact_acceleration/no_sst/trend/trend_pattern.tif"))

plot(trend)
click(trend) # make sure land and border are NA!

freq(trend)
# [1,]    -2   2723242
# [2,]     0  93295225
# [3,]     2 320172334
# [4,]    NA 329175249

# also need the number of ocean cells
freq(ocean)
# [1,]     1 416190801
# [2,]    NA 329175249


## final values
# increasing
320172334/416190801

# decreasing
3268273/416190801

```

