---
title: "KNB_file_paths"
output: html_document
---
This script organizes the files for KNB archiving (https://knb.ecoinformatics.org/).

```{r setup, include=FALSE}
library(dplyr)
library(stringr)
library(here)
source("https://raw.githubusercontent.com/OHI-Science/ohiprep_v2018/gh-pages/src/R/spatial_common.R")

```

The main KNB page will include the overall CHI and trend results, as well as supplementary files.  Each stressor will get its own KNB page.  This will include the raster files for the raw/unscaled data, rescaled data (0-1), and impact data.

I will start by identifying the individual stressor datasets.

## SST
```{r}

# raw
sst_raw_files <- list.files(file.path(dir_M, 'git-annex/impact_acceleration/stressors/sst/int'), pattern = 'sst_diff', full.names=TRUE)

sst_raw_data <- data.frame(year= NA, data_stage="raw", filepath=sst_raw_files) %>%
  mutate(year = str_sub(filepath, -8, -5)) %>%
  filter(year %in% 2003:2013)

# rescaled
sst_rescaled_files <- list.files(file.path(dir_M, sprintf('git-annex/impact_acceleration/stressors/sst/final')), full.names = TRUE)
sst_rescaled_data <- data.frame(year= NA, data_stage="rescaled", filepath=sst_rescaled_files) %>%
  mutate(year = str_sub(filepath, -21, -18)) %>%
  filter(year %in% 2003:2013)

# impact
sst_impact_files <- list.files(file.path(dir_M, sprintf('git-annex/impact_acceleration/impact/stressor_impact')), full.names = TRUE, pattern="sst")

sst_impact_data <- data.frame(year= NA, data_stage="impact", filepath=sst_impact_files) %>%
  mutate(year = str_sub(filepath, -8, -5)) %>%
  filter(year %in% 2003:2013)


# trend
sst_trend_data <- data.frame(year = "2003-2013",
                         data_stage = "trend",
                         filepath = file.path(dir_M, "git-annex/impact_acceleration/impact/trend/impacts/sst_trend.tif"))

sst_data <- rbind(sst_raw_data, sst_rescaled_data, sst_impact_data, sst_trend_data) %>%
  mutate(KNBpackage = "SST")

```


## OA
```{r}

# raw
oa_raw_files <- list.files(file.path(dir_M, 'git-annex/globalprep/prs_oa/v2017/int/annual_avg_moll'), full.names=TRUE)

oa_raw_data <- data.frame(year= NA, data_stage="raw", filepath=oa_raw_files) %>%
  mutate(year = str_sub(filepath, -8, -5)) %>%
  filter(year %in% 2003:2013)

# rescaled
oa_rescaled_files <- list.files(file.path(dir_M, 'git-annex/globalprep/prs_oa/v2017/int/annual_avg_moll_rescaled'), full.names=TRUE)
oa_rescaled_data <- data.frame(year= NA, data_stage="rescaled", filepath=oa_rescaled_files) %>%
  mutate(year = str_sub(filepath, -8, -5)) %>%
  filter(year %in% 2003:2013)

# impact
oa_impact_files <- list.files(file.path(dir_M, sprintf('git-annex/impact_acceleration/impact/stressor_impact')), full.names = TRUE, pattern="oa")

oa_impact_data <- data.frame(year= NA, data_stage="impact", filepath=oa_impact_files) %>%
  mutate(year = str_sub(filepath, -8, -5)) %>%
  filter(year %in% 2003:2013)

# trend
oa_trend_data <- data.frame(year = "2003-2013",
                         data_stage = "trend",
                         filepath = file.path(dir_M, "git-annex/impact_acceleration/impact/trend/impacts/oa_trend.tif"))


oa_data <- rbind(oa_raw_data, oa_rescaled_data, oa_impact_data, oa_trend_data) %>%
  mutate(KNBpackage="OA")

```


## SLR
```{r}

# raw
slr_raw_files <- list.files(file.path(dir_M, "git-annex/impact_acceleration/stressors/slr/int"),
                           pattern = 'gf_mol_mask', full.names = TRUE)

slr_raw_data <- data.frame(year= NA, data_stage="raw", filepath=slr_raw_files) %>%
  mutate(year = str_sub(filepath, -8, -5)) %>%
  filter(year %in% 2003:2013)

# rescaled
slr_rescaled_files <- list.files(file.path(dir_M, "git-annex/impact_acceleration/stressors/slr/final"), full.names=TRUE)

slr_rescaled_data <- data.frame(year= NA, data_stage="rescaled", filepath=slr_rescaled_files) %>%
  mutate(year = str_sub(filepath, -21, -18)) %>%
  filter(year %in% 2003:2013)

# impact
slr_impact_files <- list.files(file.path(dir_M, sprintf('git-annex/impact_acceleration/impact/stressor_impact')), full.names = TRUE, pattern="slr")

slr_impact_data <- data.frame(year= NA, data_stage="impact", filepath=slr_impact_files) %>%
  mutate(year = str_sub(filepath, -8, -5)) %>%
  filter(year %in% 2003:2013)

# trend
slr_trend_data <- data.frame(year = "2003-2013",
                         data_stage = "trend",
                         filepath = file.path(dir_M, "git-annex/impact_acceleration/impact/trend/impacts/slr_trend.tif"))


slr_data <- rbind(slr_raw_data, slr_rescaled_data, slr_impact_data, slr_trend_data) %>%
  mutate(KNBpackage="SLR")

```


## Fishing: artisanal
```{r}

# raw
art_fish_raw_files <- list.files(file.path(dir_M, 'git-annex/impact_acceleration/stressors/art_fish/int/catch_annual_npp_rasters'), full.names=TRUE)


art_fish_raw_data <- data.frame(year= NA, data_stage="raw", filepath=art_fish_raw_files) %>%
  mutate(year = str_sub(filepath, -8, -5)) %>%
  filter(year %in% 2003:2013)

# rescaled
art_fish_rescaled_files <- list.files(file.path(dir_M, "git-annex/impact_acceleration/stressors/art_fish/final"), full.names=TRUE)

art_fish_rescaled_data <- data.frame(year= NA, data_stage="rescaled", filepath=art_fish_rescaled_files) %>%
  mutate(year = str_sub(filepath, -21, -18)) %>%
  filter(year %in% 2003:2013)

# impact
art_fish_impact_files <- list.files(file.path(dir_M, sprintf('git-annex/impact_acceleration/impact/stressor_impact')), full.names = TRUE, pattern="art_fish")

art_fish_impact_data <- data.frame(year= NA, data_stage="impact", filepath=art_fish_impact_files) %>%
  mutate(year = str_sub(filepath, -8, -5)) %>%
  filter(year %in% 2003:2013)

# trend
art_fish_trend_data <- data.frame(year = "2003-2013",
                         data_stage = "trend",
                         filepath = file.path(dir_M, "git-annex/impact_acceleration/impact/trend/impacts/art_fish_trend.tif"))


art_fish_data <- rbind(art_fish_raw_data, art_fish_rescaled_data, art_fish_impact_data, art_fish_trend_data) %>%
  mutate(KNBpackage="Artisanal fishing")

```


## Commercial Fishing: demersal destructive
```{r}

# raw
dem_dest_raw_files <- list.files(file.path(dir_M, 'git-annex/impact_acceleration/stressors/comm_fish/int/catch_annual_npp_rasters/dem_dest'), full.names=TRUE)


dem_dest_raw_data <- data.frame(year= NA, data_stage="raw", filepath=dem_dest_raw_files) %>%
  mutate(year = str_sub(filepath, -8, -5)) %>%
  filter(year %in% 2003:2013)

# rescaled
dem_dest_rescaled_files <- list.files(file.path(dir_M, "git-annex/impact_acceleration/stressors/comm_fish/final/dem_dest"), full.names=TRUE)

dem_dest_rescaled_data <- data.frame(year= NA, data_stage="rescaled", filepath=dem_dest_rescaled_files) %>%
  mutate(year = str_sub(filepath, -21, -18)) %>%
  filter(year %in% 2003:2013)

# impact
dem_dest_impact_files <- list.files(file.path(dir_M, sprintf('git-annex/impact_acceleration/impact/stressor_impact')), full.names = TRUE, pattern="dem_dest")

dem_dest_impact_data <- data.frame(year= NA, data_stage="impact", filepath=dem_dest_impact_files) %>%
  mutate(year = str_sub(filepath, -8, -5)) %>%
  filter(year %in% 2003:2013)

# trend
dem_dest_trend_data <- data.frame(year = "2003-2013",
                         data_stage = "trend",
                         filepath = file.path(dir_M, "git-annex/impact_acceleration/impact/trend/impacts/dem_dest_trend.tif"))


dem_dest_data <- rbind(dem_dest_raw_data, dem_dest_rescaled_data, dem_dest_impact_data, dem_dest_trend_data) %>%
  mutate(KNBpackage="Commercial fishing: demersal destructive")

```


## Commercial Fishing: demersal nondestructive high bycatch
```{r}

# raw
dem_nondest_hb_raw_files <- list.files(file.path(dir_M, 'git-annex/impact_acceleration/stressors/comm_fish/int/catch_annual_npp_rasters/dem_nondest_hb'), full.names=TRUE)


dem_nondest_hb_raw_data <- data.frame(year= NA, data_stage="raw", filepath=dem_nondest_hb_raw_files) %>%
  mutate(year = str_sub(filepath, -8, -5)) %>%
  filter(year %in% 2003:2013)

# rescaled
dem_nondest_hb_rescaled_files <- list.files(file.path(dir_M, "git-annex/impact_acceleration/stressors/comm_fish/final/dem_nondest_hb"), full.names=TRUE)

dem_nondest_hb_rescaled_data <- data.frame(year= NA, data_stage="rescaled", filepath=dem_nondest_hb_rescaled_files) %>%
  mutate(year = str_sub(filepath, -21, -18)) %>%
  filter(year %in% 2003:2013)

# impact
dem_nondest_hb_impact_files <- list.files(file.path(dir_M, sprintf('git-annex/impact_acceleration/impact/stressor_impact')), full.names = TRUE, pattern="dem_nondest_hb")

dem_nondest_hb_impact_data <- data.frame(year= NA, data_stage="impact", filepath=dem_nondest_hb_impact_files) %>%
  mutate(year = str_sub(filepath, -8, -5)) %>%
  filter(year %in% 2003:2013)

# trend
dem_nondest_hb_trend_data <- data.frame(year = "2003-2013",
                         data_stage = "trend",
                         filepath = file.path(dir_M, "git-annex/impact_acceleration/impact/trend/impacts/dem_nondest_hb_trend.tif"))


dem_nondest_hb_data <- rbind(dem_nondest_hb_raw_data, dem_nondest_hb_rescaled_data, dem_nondest_hb_impact_data, dem_nondest_hb_trend_data) %>%
  mutate(KNBpackage="Commercial fishing: demersal nondestructive high bycatch")

```


## Commercial Fishing: demersal nondestructive high bycatch
```{r}

# raw
dem_nondest_lb_raw_files <- list.files(file.path(dir_M, 'git-annex/impact_acceleration/stressors/comm_fish/int/catch_annual_npp_rasters/dem_nondest_lb'), full.names=TRUE)


dem_nondest_lb_raw_data <- data.frame(year= NA, data_stage="raw", filepath=dem_nondest_lb_raw_files) %>%
  mutate(year = str_sub(filepath, -8, -5)) %>%
  filter(year %in% 2003:2013)

# rescaled
dem_nondest_lb_rescaled_files <- list.files(file.path(dir_M, "git-annex/impact_acceleration/stressors/comm_fish/final/dem_nondest_lb"), full.names=TRUE)

dem_nondest_lb_rescaled_data <- data.frame(year= NA, data_stage="rescaled", filepath=dem_nondest_lb_rescaled_files) %>%
  mutate(year = str_sub(filepath, -21, -18)) %>%
  filter(year %in% 2003:2013)

# impact
dem_nondest_lb_impact_files <- list.files(file.path(dir_M, sprintf('git-annex/impact_acceleration/impact/stressor_impact')), full.names = TRUE, pattern="dem_nondest_lb")

dem_nondest_lb_impact_data <- data.frame(year= NA, data_stage="impact", filepath=dem_nondest_lb_impact_files) %>%
  mutate(year = str_sub(filepath, -8, -5)) %>%
  filter(year %in% 2003:2013)

# trend
dem_nondest_lb_trend_data <- data.frame(year = "2003-2013",
                         data_stage = "trend",
                         filepath = file.path(dir_M, "git-annex/impact_acceleration/impact/trend/impacts/dem_nondest_lb_trend.tif"))


dem_nondest_lb_data <- rbind(dem_nondest_lb_raw_data, dem_nondest_lb_rescaled_data, dem_nondest_lb_impact_data, dem_nondest_lb_trend_data) %>%
  mutate(KNBpackage="Commercial fishing: demersal nondestructive low bycatch")

```


## Commercial Fishing: pelagic high bycatch
```{r}

# raw
pel_hb_raw_files <- list.files(file.path(dir_M, 'git-annex/impact_acceleration/stressors/comm_fish/int/catch_annual_npp_rasters/pel_hb'), full.names=TRUE)


pel_hb_raw_data <- data.frame(year= NA, data_stage="raw", filepath=pel_hb_raw_files) %>%
  mutate(year = str_sub(filepath, -8, -5)) %>%
  filter(year %in% 2003:2013)

# rescaled
pel_hb_rescaled_files <- list.files(file.path(dir_M, "git-annex/impact_acceleration/stressors/comm_fish/final/pel_hb"), full.names=TRUE)

pel_hb_rescaled_data <- data.frame(year= NA, data_stage="rescaled", filepath=pel_hb_rescaled_files) %>%
  mutate(year = str_sub(filepath, -21, -18)) %>%
  filter(year %in% 2003:2013)

# impact
pel_hb_impact_files <- list.files(file.path(dir_M, sprintf('git-annex/impact_acceleration/impact/stressor_impact')), full.names = TRUE, pattern="pel_hb")

pel_hb_impact_data <- data.frame(year= NA, data_stage="impact", filepath=pel_hb_impact_files) %>%
  mutate(year = str_sub(filepath, -8, -5)) %>%
  filter(year %in% 2003:2013)

# trend
pel_hb_trend_data <- data.frame(year = "2003-2013",
                         data_stage = "trend",
                         filepath = file.path(dir_M, "git-annex/impact_acceleration/impact/trend/impacts/pel_hb_trend.tif"))


pel_hb_data <- rbind(pel_hb_raw_data, pel_hb_rescaled_data, pel_hb_impact_data, pel_hb_trend_data) %>%
  mutate(KNBpackage="Commercial fishing: pelagic high bycatch")

```


## Commercial Fishing: pelagic low bycatch
```{r}

# raw
pel_lb_raw_files <- list.files(file.path(dir_M, 'git-annex/impact_acceleration/stressors/comm_fish/int/catch_annual_npp_rasters/pel_lb'), full.names=TRUE)


pel_lb_raw_data <- data.frame(year= NA, data_stage="raw", filepath=pel_lb_raw_files) %>%
  mutate(year = str_sub(filepath, -8, -5)) %>%
  filter(year %in% 2003:2013)

# rescaled
pel_lb_rescaled_files <- list.files(file.path(dir_M, "git-annex/impact_acceleration/stressors/comm_fish/final/pel_lb"), full.names=TRUE)

pel_lb_rescaled_data <- data.frame(year= NA, data_stage="rescaled", filepath=pel_lb_rescaled_files) %>%
  mutate(year = str_sub(filepath, -21, -18)) %>%
  filter(year %in% 2003:2013)

# impact
pel_lb_impact_files <- list.files(file.path(dir_M, sprintf('git-annex/impact_acceleration/impact/stressor_impact')), full.names = TRUE, pattern="pel_lb")

pel_lb_impact_data <- data.frame(year= NA, data_stage="impact", filepath=pel_lb_impact_files) %>%
  mutate(year = str_sub(filepath, -8, -5)) %>%
  filter(year %in% 2003:2013)

# trend
pel_lb_trend_data <- data.frame(year = "2003-2013",
                         data_stage = "trend",
                         filepath = file.path(dir_M, "git-annex/impact_acceleration/impact/trend/impacts/pel_lb_trend.tif"))


pel_lb_data <- rbind(pel_lb_raw_data, pel_lb_rescaled_data, pel_lb_impact_data, pel_lb_trend_data) %>%
  mutate(KNBpackage="Commercial fishing: pelagic low bycatch")

```


## Nutrient pollution
```{r}

# raw
nutrient_raw_files <- list.files(file.path(dir_M, 
        "git-annex/impact_acceleration/stressors/land_based/int"), full.names = TRUE, pattern="log")
nutrient_raw_files <- grep("fert", nutrient_raw_files, value=TRUE)

nutrient_raw_data <- data.frame(year= NA, data_stage="raw", filepath=nutrient_raw_files) %>%
  mutate(year = str_sub(filepath, -12, -9)) %>%
  filter(year %in% 2003:2013)

# rescaled
nutrient_rescaled_files <- list.files(file.path(dir_M, "git-annex/impact_acceleration/stressors/land_based/final/nutrient"), full.names=TRUE)

nutrient_rescaled_data <- data.frame(year= NA, data_stage="rescaled", filepath=nutrient_rescaled_files) %>%
  mutate(year = str_sub(filepath, -21, -18)) %>%
  filter(year %in% 2003:2013)

# impact
nutrient_impact_files <- list.files(file.path(dir_M, sprintf('git-annex/impact_acceleration/impact/stressor_impact')), full.names = TRUE, pattern="nutrient")

nutrient_impact_data <- data.frame(year= NA, data_stage="impact", filepath=nutrient_impact_files) %>%
  mutate(year = str_sub(filepath, -8, -5)) %>%
  filter(year %in% 2003:2013)

# trend
nutrient_trend_data <- data.frame(year = "2003-2013",
                         data_stage = "trend",
                         filepath = file.path(dir_M, "git-annex/impact_acceleration/impact/trend/impacts/nutrient_trend.tif"))


nutrient_data <- rbind(nutrient_raw_data, nutrient_rescaled_data, nutrient_impact_data, nutrient_trend_data) %>%
  mutate(KNBpackage="Nutrient pollution")

```

## Organic chemical pollution
```{r}

# raw
organic_raw_files <- list.files(file.path(dir_M, 
        "git-annex/impact_acceleration/stressors/land_based/int"), full.names = TRUE, pattern="log")
organic_raw_files <- grep("pest", organic_raw_files, value=TRUE)

organic_raw_data <- data.frame(year= NA, data_stage="raw", filepath=organic_raw_files) %>%
  mutate(year = str_sub(filepath, -12, -9)) %>%
  filter(year %in% 2003:2013)

# rescaled
organic_rescaled_files <- list.files(file.path(dir_M, "git-annex/impact_acceleration/stressors/land_based/final/organic"), full.names=TRUE)

organic_rescaled_data <- data.frame(year= NA, data_stage="rescaled", filepath=organic_rescaled_files) %>%
  mutate(year = str_sub(filepath, -21, -18)) %>%
  filter(year %in% 2003:2013)

# impact
organic_impact_files <- list.files(file.path(dir_M, sprintf('git-annex/impact_acceleration/impact/stressor_impact')), full.names = TRUE, pattern="organic")

organic_impact_data <- data.frame(year= NA, data_stage="impact", filepath=organic_impact_files) %>%
  mutate(year = str_sub(filepath, -8, -5)) %>%
  filter(year %in% 2003:2013)


# trend
organic_trend_data <- data.frame(year = "2003-2013",
                         data_stage = "trend",
                         filepath = file.path(dir_M, "git-annex/impact_acceleration/impact/trend/impacts/organic_trend.tif"))

organic_data <- rbind(organic_raw_data, organic_rescaled_data, organic_impact_data, organic_trend_data) %>%
  mutate(KNBpackage="Organic chemical pollution")

```

## Direct human
```{r}

# raw
direct_human_raw_files <- list.files(file.path(dir_M, "git-annex/impact_acceleration/stressors/direct_human/int"), pattern="log_mask", full.name=TRUE)

direct_human_raw_data <- data.frame(year= NA, data_stage="raw", filepath=direct_human_raw_files) %>%
  mutate(year = str_sub(filepath, -17, -14)) %>%
  filter(year %in% 2003:2013)

# rescaled
direct_human_rescaled_files <- list.files(file.path(dir_M, "git-annex/impact_acceleration/stressors/direct_human/final"), full.names=TRUE)

direct_human_rescaled_data <- data.frame(year= NA, data_stage="rescaled", filepath=direct_human_rescaled_files) %>%
  mutate(year = str_sub(filepath, -21, -18)) %>%
  filter(year %in% 2003:2013)

# impact
direct_human_impact_files <- list.files(file.path(dir_M, sprintf('git-annex/impact_acceleration/impact/stressor_impact')), full.names = TRUE, pattern="direct_human")

direct_human_impact_data <- data.frame(year= NA, data_stage="impact", filepath=direct_human_impact_files) %>%
  mutate(year = str_sub(filepath, -8, -5)) %>%
  filter(year %in% 2003:2013)

# trend
direct_human_trend_data <- data.frame(year = "2003-2013",
                         data_stage = "trend",
                         filepath = file.path(dir_M, "git-annex/impact_acceleration/impact/trend/impacts/direct_human_trend.tif"))


direct_human_data <- rbind(direct_human_raw_data, direct_human_rescaled_data, direct_human_impact_data, direct_human_trend_data) %>%
  mutate(KNBpackage="Direct human")

```


## Light
```{r}

# raw
light_raw_files <- list.files(file.path(dir_M, "git-annex/impact_acceleration/stressors/light_pollution/int/calibrated_rasts"), full=TRUE)


light_raw_data <- data.frame(year= NA, data_stage="raw", filepath=light_raw_files) %>%
  mutate(year = str_sub(filepath, -21, -18)) %>%
  filter(year %in% 2003:2013)

# rescaled
light_rescaled_files <- list.files(file.path(dir_M, "git-annex/impact_acceleration/stressors/light_pollution/final"), full.names=TRUE)

light_rescaled_data <- data.frame(year= NA, data_stage="rescaled", filepath=light_rescaled_files) %>%
  mutate(year = str_sub(filepath, -21, -18)) %>%
  filter(year %in% 2003:2013)

# impact
light_impact_files <- list.files(file.path(dir_M, sprintf('git-annex/impact_acceleration/impact/stressor_impact')), full.names = TRUE, pattern="light")

light_impact_data <- data.frame(year= NA, data_stage="impact", filepath=light_impact_files) %>%
  mutate(year = str_sub(filepath, -8, -5)) %>%
  filter(year %in% 2003:2013)

# trend
light_trend_data <- data.frame(year = "2003-2013",
                         data_stage = "trend",
                         filepath = file.path(dir_M, "git-annex/impact_acceleration/impact/trend/impacts/light_trend.tif"))


light_data <- rbind(light_raw_data, light_rescaled_data, light_impact_data, light_trend_data) %>%
  mutate(KNBpackage="Light")

```

## Shipping
```{r}

# raw
shipping_raw_files <- list.files(file.path(dir_M, 
      "git-annex/impact_acceleration/stressors/shipping/int"), full.names = TRUE, pattern="mol_log")


shipping_raw_data <- data.frame(year= NA, data_stage="raw", filepath=shipping_raw_files) %>%
  mutate(year = str_sub(filepath, -16, -13)) %>%
  filter(year %in% 2003:2013)

# rescaled
shipping_rescaled_files <- list.files(file.path(dir_M, "git-annex/impact_acceleration/stressors/shipping/final"), full.names=TRUE)

shipping_rescaled_data <- data.frame(year= NA, data_stage="rescaled", filepath=shipping_rescaled_files) %>%
  mutate(year = str_sub(filepath, -21, -18)) %>%
  filter(year %in% 2003:2013)

# impact
shipping_impact_files <- list.files(file.path(dir_M, sprintf('git-annex/impact_acceleration/impact/stressor_impact')), full.names = TRUE, pattern="shipping")

shipping_impact_data <- data.frame(year= NA, data_stage="impact", filepath=shipping_impact_files) %>%
  mutate(year = str_sub(filepath, -8, -5)) %>%
  filter(year %in% 2003:2013)

# trend
shipping_trend_data <- data.frame(year = "2003-2013",
                         data_stage = "trend",
                         filepath = file.path(dir_M, "git-annex/impact_acceleration/impact/trend/impacts/shipping_trend.tif"))


shipping_data <- rbind(shipping_raw_data, shipping_rescaled_data, shipping_impact_data, shipping_trend_data) %>%
  mutate(KNBpackage="Shipping")

```

## Combine stressor data

Combine, check and final formatting of the data.

```{r}

all_stressors <- rbind(sst_data, oa_data, slr_data, 
                       art_fish_data, dem_dest_data, dem_nondest_hb_data, dem_nondest_lb_data, pel_hb_data, pel_lb_data,
                       nutrient_data, organic_data, direct_human_data, light_data,
                       shipping_data)

table(all_stressors$KNBpackage)
length(table(all_stressors$KNBpackage))

all_stressors_formatted <- all_stressors %>%
  mutate(new_file_name = sprintf("%s_%s_%s.tif", gsub(" ", "_", tolower(KNBpackage)), year, data_stage)) %>%
    mutate(new_file_name = gsub(":", "", new_file_name))%>%
  mutate(data_name = sprintf("%s data for year %s (%s)", KNBpackage,  year, data_stage)) %>%
  select(KNBpackage, year, data_stage, filepath, new_file_name, data_name)                              

sum(duplicated(all_stressors_formatted$new_file_name))
```


## Cumulative impacts
```{r}

## cumulative impact
# impact
cumulative_impact_files <- list.files(file.path(dir_M, sprintf('git-annex/impact_acceleration/impact/cumulative_impact')), full.names = TRUE, pattern="chi")

cumulative_impact_data <- data.frame(year= NA, data_stage="cumulative_impact", filepath=cumulative_impact_files) %>%
  mutate(year = str_sub(filepath, -8, -5)) %>%
  filter(year %in% 2003:2013)

# trend
cumulative_impact_trend_data <- data.frame(year = "2003-2013",
                         data_stage = "cumulative_impact_trend",
                         filepath = file.path(dir_M, "git-annex/impact_acceleration/impact/trend/chi_slope.tif"))


cumulative_impacts <- rbind(cumulative_impact_data, cumulative_impact_trend_data) %>%
  mutate(KNBpackage="Cumulative Impact") %>%
  mutate(new_file_name = sprintf("%s_%s.tif", data_stage, year)) %>%
  mutate(data_name = sprintf("%s data for year %s (%s)", KNBpackage,  year, gsub("_", " ", data_stage))) %>%
  select(KNBpackage, year, data_stage, filepath, new_file_name, data_name)                  



```

# Habitat data

```{r}

# habitat data
habitat_files <- list.files(file.path(dir_M, sprintf('git-annex/impact_acceleration/habitats')), full.names = TRUE)
habitat_files <- grep(".tif.", habitat_files, value=TRUE, invert=TRUE)

habitat_names <- read.csv(here("habitats/habitat.csv")) %>%
  select(tif_name, habitat_name_full = plot_name, habitat_name_abb = plot_name3) %>%
   mutate(habitat_name_abb = gsub("\\.", "", habitat_name_abb)) %>%
     mutate(habitat_name_abb = gsub(" ", "_", habitat_name_abb)) %>%
   mutate(habitat_name_abb = paste0(habitat_name_abb, ".tif")) %>%
  mutate(tif_name = paste0(tif_name, ".tif")) 

sum(duplicated(habitat_names$habitat_name_abb))

habitat_data <- data.frame(KNBpackage = "Cumulative Impact", year = NA, data_stage = "habitat", filepath=habitat_files, tif_name = basename(habitat_files)) %>%
  left_join(habitat_names, by = "tif_name") %>%
mutate(new_file_name = sprintf("%s_%s", data_stage, habitat_name_abb)) %>% 
  mutate(data_name = sprintf("habitat data describing location of: %s", habitat_name_full)) %>%
select(KNBpackage, year, data_stage, filepath, new_file_name, data_name) 


```

## location of GitHub file
```{r}

github_files <- file.path(dir_M, 'git-annex/impact_acceleration/impact_acceleration-1.0.zip')

github_data <- data.frame(KNBpackage = "Cumulative Impact", year= NA, data_stage="cumulative_impact", filepath=github_files, new_file_name = "impact_acceleration-1.0.zip",
    data_name = "Github repository with data prep scripts and other files")


```


## location of vulnerability matrix
```{r}

vulnerability_file <- file.path(dir_M, 'git-annex/impact_acceleration/vulnerability_weighting_matrix.csv')

vulnerability_data <- data.frame(KNBpackage = "Cumulative Impact", year= NA, data_stage="cumulative_impact", 
                          filepath = vulnerability_file, new_file_name = "vulnerability_weighting_matrix.csv",
    data_name = "Vulnerability table describing impact of stressors on marine ecosystems")


```



## combine all data

```{r}

final_list <- rbind(all_stressors_formatted, cumulative_impacts, habitat_data, github_data, vulnerability_data) %>%
  mutate(filepath=as.character(filepath))


table(final_list$KNBpackage)
length(table(final_list$KNBpackage))
filter(final_list, KNBpackage=="Cumulative Impact")
sum(duplicated(final_list$new_file_name))
sum(file.exists(final_list$filepath))
dim(final_list)


write.csv(final_list, here::here("paper/KNB_datalist.csv"), row.names=FALSE)