---
title: 'OHI 2018: Sea Surface Temperature Data Explore'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '../../../src/templates/ohi_hdr.html'
  pdf_document:
    toc: true
---

## Setup

```{r setup, message=F,warning=F, eval = FALSE}

knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/', message = FALSE, warning = FALSE)

source("https://raw.githubusercontent.com/OHI-Science/ohiprep_v2018/gh-pages/src/R/spatial_common.R")

library(raster)
library(RColorBrewer)
library(tidyverse)
library(rgdal)
library(doParallel)
library(foreach)
library(sf)
library(ncdf4)
library(httr)
library(lubridate)
library(animation)
library(ggplot2)
library(plotly)
library(here)
library(stringr)

cols <- rev(colorRampPalette(brewer.pal(11, 'Spectral'))(255)) # rainbow color scheme
land <- regions %>% subset(rgn_type %in% c("land", "land-disputed", "land-noeez"))

yrs_data=1982:2017


global_plot_function <- function(plotRast, cols=cols, title = "", include_legend=TRUE){

  par(mar=c(0.1,1,1,1)) # bottom, left, top, and right
  par(oma=c(0,0,0,0))

plot(plotRast, col=cols,  
     breaks=c(-.1, seq(0,
                  200, length.out=254)), 
     legend=FALSE, axes=FALSE, box=FALSE)

  title(main=list(title, cex=1.2), line=0)


  if(include_legend){
par(mfrow=c(1, 1), mar=c(2, 0, 1, 0), new=FALSE)
plot(plotRast, legend.only=TRUE, legend.shrink=.8, legend.width=.5, col=cols, breaks=c(-.1, seq(0,
                  200, length.out=254)), axis.args = list(cex.axis = 0.6, at=c(0, 50, 100, 150, 200), labels=c(0, 50, 100, 150, 200)))
}
}


```


## Baseline raster

Cumulative number of weekly extreme temperature events during a 5 year period from 1985-1989.  Extreme events are the number of weeks in a year in which the weekly temperature exceeds the mean + 1 SD.    

```{r}

# make mask
org_data         <- stack(list.files(file.path(dir_M, "git-annex/globalprep/_raw_data/CoRTAD_sst/d2018/"),
                                 pattern = "SSTA.nc",
                                 full.names = TRUE), varname = "SSTA")


mask_data <- org_data[[1]] %>%
  projectRaster(crs=raster::crs(ocean), method="ngb", over=TRUE)

# data
l   <- list.files(file.path(dir_M,'git-annex/impact_acceleration/stressors/sst/int'), pattern='sst_anom', full.names=TRUE)
  

# Get baseline 5 year aggregates
baseline_years <- grep(c('1985|1986|1987|1988|1989'), l, value=TRUE)

baseline <- stack(ref_years) %>% 
    sum(.) %>%
    mask(mask_data)


global_plot_function(plotRast=baseline, col=cols)

  
  
```

### Cumulative extreme temperature events
The cumulative number of extreme temperature events is calculated for the year of interest and the 4 years previous.  We use a 5 year window to help smooth variation not due to longer term climate patterns.  

```{r}

sst <- raster(file.path(dir_M, 'git-annex/impact_acceleration/stressors/sst/int/sst_5_yr_2009-2013.tif')) %>%
  mask(mask_data)

global_plot_function(plotRast=sst, col=cols)

```

### Increase in cumulative extreme temperature events relative to baseline
We determine the increase in cumulative extreme temperature events relative to the baseline period.  This is calculated by subtracting the current number of extreme events from the baseline period.  
```{r}

sst <- raster(file.path(dir_M, 'git-annex/impact_acceleration/stressors/sst/int/sst_diff_2009-2013.tif')) %>%
  mask(mask_data)

plot(sst, col=cols)
global_plot_function(plotRast=sst, col=cols)

sst <- raster(file.path(dir_M, 'git-annex/impact_acceleration/stressors/sst/int/sst_diff_2013-2017.tif')) %>%
  mask(mask_data)

plot(sst, col=cols)
global_plot_function(plotRast=sst, col=cols)

```
