---
title: 'OHI 2018: Sea Surface Temperature Pressure Layer'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '../../../src/templates/ohi_hdr.html'
  pdf_document:
    toc: true
---


# Summary

This script creates the Sea Surface Temperature (SST) data for the Cumulative Human Impacts project.



# Data Source

Data comes from [CoRTAD version 6](https://data.nodc.noaa.gov/cgi-bin/iso?id=gov.noaa.nodc:NCEI-CoRTADv6)


**Native Data Resolution**: ~4km   
**Description**: 
Cortadv6_SSTA.nc = SST anomalies (weekly SST minus weekly climatological SST), weekly data for all years, degrees Kelvin
**Time Range**: 1982 - 2017 (weekly averages across all years)  
**Format**: NetCDF
**Downloaded**: August 21, 2018

***  

# Methods

1. Extreme events per year based calculated as number of times SST anomaly exceeds SST Standard Deviation based on weekly values.
2. Sum extreme events for five year periods to control for yearly variation.
3. Change in extreme events: Subtract number of extreme events for each five year period from control period (1985-1989).
4. Rescale "Change in extreme events" data to values between 0 and 1 by dividing by the 99.99th quantile among all years of data.

## Setup

```{r setup, message=F,warning=F, eval = FALSE}

knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/', message = FALSE, warning = FALSE)

source("https://raw.githubusercontent.com/OHI-Science/ohiprep_v2018/gh-pages/src/R/spatial_common.R")

library(raster)
library(RColorBrewer)
library(tidyverse)
library(rgdal)
library(doParallel)
library(foreach)
library(sf)
library(ncdf4)
library(httr)
library(lubridate)
library(animation)
library(ggplot2)
library(plotly)
library(here)
library(stringr)

cols <- rev(colorRampPalette(brewer.pal(11, 'Spectral'))(255)) # rainbow color scheme
land <- regions %>% subset(rgn_type %in% c("land", "land-disputed", "land-noeez"))

yrs_data=1982:2017

```

***


***

## Generate annual positive anomalies

Calculate the standard deviation of the anomolies for each week across all years.
```{r sd, eval = FALSE}

## load netcdf uv radiation data
ssta         <- stack(list.files(file.path(dir_M, "git-annex/globalprep/_raw_data/CoRTAD_sst/d2018"),
                                 pattern = "SSTA.nc",
                                 full.names = TRUE), varname = "SSTA")
# weekly_sst   <- stack(list.files(file.path(dir_M, "git-annex/globalprep/_raw_data/CoRTAD_sst/d2018"), pattern = "WeeklySST.nc",
#                                  full.names = TRUE), varname = "WeeklySST")

names_ssta   <- names(ssta)
#names_weekly <- names(weekly_sst)

ssta_df <- names_ssta %>% # View(ssta_df)
  data.frame() %>% 
  rename(name = ".") %>% 
  mutate(year = substr(name, 2, 5), 
         month = substr(name, 7, 8), 
         day = substr(name, 10, 11)) %>% 
  mutate(week = week(as.Date(sprintf("%s-%s-%s", year, month, day))))

## the next for-loop takes a long time, ~22min for each of 53 layers
## create weekly standard deviations across all years
for(i in 1:53){ # i = 1
  t0 = Sys.time()
  print(paste("calculating sd for week", i, "-- started at", t0))
  s = stack()
  
  for (j in yrs_data){ # FOR APPROACH OF USING REF PERIOD TO CALC EXTREME EVENTS: CHANGE 'YRS' HERE TO INCLUDE JUST REFERENCE YEARS #yrs=
    w = which(substr(names_weekly, 2, 5) == j)[i]
    if(is.na(w)) next() # most yrs don't have 53 weeks; 'next' works in for loop but not foreach+dopar
    w_week = weekly_sst[[w]]
    s = stack(s, w_week)
  }
  
  sd_sst = calc(s, fun = function(x){sd(x, na.rm = TRUE)},
                progress ="text",
                filename = file.path(dir_M, "git-annex/globalprep/prs_sst/v2018/int"), sprintf("sd_sst_week_%s.tif", i)))
}

```


Count the number of times in each year that the weekly anomoly surpasses the sd of the anomolies. Note: this is equivalent to calculating whether the weekly SST surpasses weekly mean SST + 1SD.
```{r}

registerDoParallel(5)
combine_fun = function(x){sum(x, na.rm = TRUE)} # takes raster stack object as x arg

## calculate annual positive anomalies; ~17 minutes per year with 5 cores
for(j in yrs){ #j=1982
  t0 = Sys.time()
  print(paste("calculating anomaly for", j, "-- started at", t0))
  s = stack()
  
  # wks = ssta_df %>% 
  #   filter(year == j) %>% 
  #   select(week)
  
  wks <- data.frame(week = 1:52) # only include 1:52 to keep even across years (a few years have 53 weeks)
  
  s <- foreach(i = wks$week, .packages = c("raster", "ncdf4", "rgdal"), .combine = "stack") %dopar%{ # i = 1
    sd_sst = raster::raster(file.path(dir_M, "git-annex/globalprep/prs_sst/v2018/int", sprintf("sd_sst_week_%s.tif", i))) 
    w = which(substr(names_ssta, 2, 5) == j)[i]
    w_ssta = ssta[[w]]
    raster::overlay(w_ssta, sd_sst, 
                    fun = function(x, y){ifelse(is.na(x) | is.na(y), 0, ifelse(x > y, 1, 0))})
  }
  
  yr = combine_fun(s)
  raster::writeRaster(yr, filename = file.path(dir_M, "git-annex/impact_acceleration/stressors/sst/int", 
          sprintf("annual_pos_anomalies_sd_%s.tif", j)))
}
```



## Identify gapfilled cells
```{r, eval=FALSE}

# number of positive anomalies
l   <- list.files(file.path(dir_M, 'git-annex/impact_acceleration/stressors/sst/int'), pattern='annual_pos_anomalies', full.names=TRUE)

# use the original data as mask at this spatial scale
org_data         <- stack(list.files(file.path(dir_M, "git-annex/globalprep/_raw_data/CoRTAD_sst/d2018/"),
                                 pattern = "SSTA.nc",
                                 full.names = TRUE), varname = "SSTA")

mask_data <- org_data[[1]]

# identify cells that need to be gapfilled
file_gf <- raster(l[1]) %>%
  mask(mask_data)

file_gf[is.na(file_gf)] <- 999

projectRaster(file_gf, ocean, method = "ngb", over=TRUE) %>%
  mask(ocean, filename = file.path(dir_M, "git-annex/impact_acceleration/stressors/sst/int/sst_gf.tif"),
       overwrite = TRUE)


file_gf <- raster(file.path(dir_M, "git-annex/impact_acceleration/stressors/sst/int/sst_gf.tif"))

plot(file_gf)

# get count of gapfilled cells
file_gf_count <- getValues(file_gf)
gf_count <- sum(file_gf_count==999, na.rm=TRUE)
gf_count # 904099
ocean_vals <- getValues(ocean)
ocean_size <- sum(ocean_vals == 1, na.rm=TRUE)
ocean_size
904099/416190801 # 0.2% gapfilled


```


## interpolate each year of sst raster data, crs to mollweide, mask, and save
```{r interpolate, eval=FALSE}

l   <- list.files(file.path(dir_M, 'git-annex/impact_acceleration/stressors/sst/int'), pattern='annual_pos_anomalies', full.names=TRUE)
  
plot(raster(l[1]))
#click(raster(l[1])) # unfortunately, land areas are 0 and not NA, need to mask prior to gapfilling

# use the original data as mask at this spatial scale
org_data         <- stack(list.files(file.path(dir_M, "git-annex/globalprep/_raw_data/CoRTAD_sst/d2018/"),
                                 pattern = "SSTA.nc",
                                 full.names = TRUE), varname = "SSTA")

mask_data <- org_data[[1]]

## Gapfill using mean of surrounding cells that are NA
gf_raster <- function(x){raster::focal(x, w = matrix(1,3,3), fun = mean, na.rm=TRUE, pad = TRUE, NAonly=TRUE)}

registerDoParallel(2)
foreach(file_name = l,.packages="dplyr") %dopar%{ # file_name = l[1]
  
  yr <- substr(file_name, nchar(file_name)-7, nchar(file_name)-4)
  
  r = raster::raster(file_name) %>%
    raster::mask(mask_data) %>%
    raster::projectRaster(crs=raster::crs(ocean), method="ngb", over=TRUE)
  

## Repeat 400 times (I found this was enough iterations to gapfill all missing values)
i <- 0
while (i <= 400){
r <- gf_raster(r)
i <- i + 1
print(i)
}

raster::writeRaster(r, file.path(dir_M, sprintf("git-annex/impact_acceleration/stressors/sst/int/sst_anom_interpolated_%s.tif", yr)), overwrite=TRUE)

}


```


## Create 5 year cumulative sum of extreme events and calculate difference from historical
```{r, eval=FALSE}

l   <- list.files(file.path(dir_M,'git-annex/impact_acceleration/stressors/sst/int'), pattern='sst_anom', full.names=TRUE)
  
plot(raster(l[1]))

# Get 5 year aggregates
ref_years <- grep(c('1985|1986|1987|1988|1989'), l, value=TRUE)

  ref <- stack(ref_years) %>% 
    sum(.) # This is the time period we are using for historical comparison (1985 - 1989)

  
for (i in 1994:(max(yrs_data))){ #i=2017
  print(i)
  yrs <- c(i:(i-4))
  yrs_comb <- paste(yrs, collapse="|")
  s   <- stack(grep(yrs_comb, l, value=TRUE)) %>% 
    sum(.)

    writeRaster(s,
              filename = file.path(dir_M, sprintf('git-annex/impact_acceleration/stressors/sst/int/sst_5_yr_%s-%s.tif', yrs[5], yrs[1])), overwrite=TRUE)
    
  diff = overlay(s, ref, fun=function(x,y){x-y}) #calculate difference between recent 5 year cumulative sum and historical (1985-1989)

  writeRaster(diff,
              filename = file.path(dir_M, sprintf('git-annex/impact_acceleration/stressors/sst/int/sst_diff_%s-%s.tif', yrs[5], yrs[1])), overwrite=TRUE)
}

```


### Calculate reference value

Reference values are the 99.99th quantile of the total positive anomolies across five years subtracted from the total positive anomolies for a five year reference period (1985-1989).  This is performed on non gapfilled data.  
```{r ref point, eval=FALSE}

l   <- list.files(file.path(dir_M, 'git-annex/impact_acceleration/stressors/sst/int'), pattern='annual_pos_anomalies', full.names=TRUE)

# use the original data as mask at this spatial scale
# the files have land as zero (but there are also legitimate zeros, 
# so these need to be masked)
org_data         <- stack(list.files(file.path(dir_M, "git-annex/globalprep/_raw_data/CoRTAD_sst/d2018/"),
                                 pattern = "SSTA.nc",
                                 full.names = TRUE), varname = "SSTA")


mask_data <- org_data[[1]]

ref <- stack(grep(c('1985|1986|1987|1988|1989'), l, value=TRUE)) %>%  # Time period we are using for historical comparison (1985 - 1989)
  mask(mask_data) %>%
  sum(.)


vals <- c()

for (i in 1994:max(yrs_data)){ #i=1994
  
  yrs <- c(i:(i-4))
  
  s   <- stack(l[str_sub(l, -8, -5) %in% yrs]) %>% 
    sum(.)

  diff <-  overlay(s, ref, fun=function(x,y){x-y}) #calculate difference between recent 5 year cumulative sum and historical (1985-1989)

  diff_vals <- getValues(diff)
    
  vals <- c(vals, diff_vals)
  print(i)
}

    
# 99.99th quantile

resc_num  <- quantile(vals, prob=0.9999, na.rm=TRUE) ### 130, 26 weeks of extreme events
resc_num

hist(vals, main="SST extreme events across all years/cells")
abline(v = resc_num, col="red")

ref_value <- data.frame(pressure = "sst", quant_9999 = resc_num)

write.csv(ref_value, 
          here("stressors/sst/rescale_val.csv"),
          row.names=FALSE)

```



## Rescaling

```{r rescale, eval=FALSE}

diffs <- list.files(file.path(dir_M, 'git-annex/impact_acceleration/stressors/sst/int'), pattern = 'sst_diff', full.names=TRUE)


resc_num <- read.csv( 
          here("stressors/sst/rescale_val.csv")) %>%
  .$quant_9999


for (diff_rast in diffs){ # diff_rast <- diffs[23]

  print(diff_rast)

  r <- raster(diff_rast)

  yrs <- str_sub(diff_rast, -8, -5)
  
  out = raster::calc(r, fun=function(x){ifelse(x>0, 
                                       ifelse(x>resc_num, 1, x/resc_num), 
                                       0)}) %>%
  raster::resample(ocean, method = "ngb") %>%
  raster::mask(ocean)
  
    
  writeRaster(out, file.path(dir_M, sprintf('git-annex/impact_acceleration/stressors/sst/final/sst_%s_rescaled_mol.tif', yrs)), overwrite=TRUE)

}

```


## Quick compare with previous version

```{r rescale, eval=FALSE}

new <- raster(file.path(dir_M, "git-annex/impact_acceleration/stressors/sst/final/sst_2012_rescaled_mol.tif"))
plot(new)

old <- raster(file.path(dir_M, "git-annex/impact_acceleration/stressors/sst/final/sst_2012_rescaled_mol.tif"))
plot(old)

plot(new, old)
abline(0,1, col="red")  
```


***

# Citation information  

Selig, E.R., K.S. Casey, and J.F. Bruno (2010), New insights into global patterns of ocean temperature anomalies: implications for coral reef health and management, Global Ecology and Biogeography, DOI: 10.1111/j.1466-8238.2009.00522.x.
