---
title: "Annual Commercial Fisheries Catch"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ohi-northeast/src/templates/ohi_hdr.html'
pdf_document:
  toc: true
---

This script combines spatial catch data from [Watson (2017)](https://www.nature.com/articles/sdata201739) with species & gear specific categories defined in [`watson_gear_matching`](https://github.com/OHI-Science/impact_acceleration/blob/master/stressors/comm_fish/watson_gear_matching.Rmd) to create annual catch rasters for each of the five fishing stressor categories:

- demersal destructive  
- demersal non-destructive high bycatch
- demersal non-destructive low bycatch
- pelagic high bycatch
- pelagic low bycatch

## Overview
- Raw catch data is combined with `taxa_gear_types.csv` as defined in `watson_gear_matching.Rmd`
- Data is filtered for every year and catch type, rasterized to half degree cells, and saved as a GeoTIFF

```{r setup, message=F,warning=F}

knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',message = FALSE, warning = FALSE)

library(seaaroundus)
library(raster)
library(readr)
library(dplyr)
library(data.table)
library(parallel)
library(foreach)
library(doParallel)
library(RColorBrewer)
library(stringr)

source("https://raw.githubusercontent.com/OHI-Science/ohiprep_v2018/master/src/R/spatial_common.R")

cols = rev(colorRampPalette(brewer.pal(11, 'Spectral'))(255)) # rainbow color scheme

# setwd("stressors/comm_fish")

```

#Load Data

Load raw catch data and the dataset matching gear and taxon to the five stressor categories.

```{r raw_data_files, eval=F}

watson_data <- list.files(file.path(dir_M, 'git-annex/impact_acceleration/stressors/comm_fish/data/catch'), full.names=TRUE, pattern = '.rds')

#dataset created in watson_gear_matching that assigns each species/gear pair a gear type as defined in Halpern et al. (2008)
gear_taxa <- read_csv('taxa_gear_types.csv')%>%
            dplyr::select(-X1)%>%
            rename(Taxonkey = TaxonKey)
```

#Save annual catch data

The raw data comes with 5 years of data in each file. Here we disaggregate the data by year, join it with the `taxa_gear` data, and save the filtered data as a standalone dataset.

```{r annual_catch_data,eval=F}

#annual_csv is a function that reads in a raw RDS file, finds all years in the data, filters the data for each year, joins the gear_taxa data and saves its as an .rds on the server

registerDoParallel(10) #register cores

foreach(file = watson_data) %dopar% { # file = watson_data[1]
  
data <- readRDS(file)

yrs <- unique(data$Year)

for(i in 1:length(yrs)){
  print(i)
  
yr = yrs[i]
  
yr_data <- data%>%
            filter(Year == yr)%>%
            left_join(gear_taxa, by = c('Taxonkey','TaxonName','CommonName', 'Gear', 'GearName'))

write_rds(yr_data,path = paste0(file.path(dir_M),'/git-annex/impact_acceleration/stressors/comm_fish/int/catch_annual_data/catch_data_',yr,'.rds'))

}
}

```

#Map annual catch

## List annual catch files

To use the datasets just created we define a list of all annual catch datasets.

```{r}
annual_files <- list.files(file.path(dir_M, 'git-annex/impact_acceleration/stressors/comm_fish/int/catch_annual_data'), full.names=TRUE)

```


## Get base raster

Using the [searoundus R package](https://github.com/ropensci/seaaroundus), we create a raster with cell values set equal to cell IDs in the data, which are defined in the column `Seq`.

```{r}

  #get template raster for SAUP data with cell values equal to CellID
  
   saup_cells <- getcells("POLYGON ((-180 90, 180 90, 180 -90, -180 -90, -180 90))")

   saup_rast <- raster(ncol=720, nrow=360)
   saup_rast[] <- saup_cells
   
   plot(saup_rast, col=cols, main = "SAUP Cell IDs")
   
```


## Create a buffer to gapfill shoreline NA values

The ocean mask is used to create a buffer to identify shoreline values, but first it needs to have the same resolution and CRS as the fisheries data.  

This buffer is one cell layer inland and two cell layers offshore (in this case the buffer region is NA values and the other regions are 1 values).  The NA's in the fishing layer will be converted to 0 values.  The buffer and fishing layers will be summed (na.rm=FALSE) resulting in . 
```{r ocean}

## Make a boundary mask
 saup_rast_mol <- projectRaster(saup_rast, crs=crs(ocean), over=TRUE)
 saup_rast_mol[saup_rast_mol>=0] <- 1  
 plot(saup_rast_mol)
 boundary_mask <- boundaries(saup_rast_mol, type="inner", asNA=TRUE) 
 boundary_mask <- boundaries(boundary_mask, type="outer", asNA=FALSE) 
 boundary_mask[boundary_mask==0] <- 1 
 boundary_mask <- boundaries(boundary_mask, type="outer", asNA=FALSE) 
 boundary_mask[boundary_mask==0] <- 1 

plot(boundary_mask)

ocean_low_res = resample(ocean, saup_rast_mol, method="ngb")
plot(ocean_low_res)

# this creates a buffer that extends one raster cell inland and 2 raster cells offshore
buffer1 <- boundaries(ocean_low_res, type="inner", asNA=TRUE, progress="text") 
buffer2 <- boundaries(buffer1, type="outer", asNA=FALSE, progress="text") 
buffer2[buffer2==0] <- 1 
buffer2 <- subs(buffer2, data.frame(id=c(NA,1), v=c(0,NA)))
plot(buffer2)
#zoom(buffer2)

# get rid of outer boundary on buffer
coastal_buffer <- overlay(boundary_mask, buffer2, fun = function(x,y){ifelse(x==1 & is.na(y), 0, y)})
plot(coastal_buffer)

```

#### Create a function to gapfill missing coastal data

Function to identify the cells that need to be gapfilled. Then average of nearest neighbors is used to gapfill near-coastal raster cells that are NA. 
```{r}

fish_gf <- function(catch_path, file_name, yr){

rast <-  raster(file.path(dir_M, catch_path, file_name, paste0(file_name, "_", yr, ".tif"))) 
#plot(rast)

# temporarily convert NA values that are close to land to 999 (these will be converted to NA ultimately)
fis_gf <- overlay(rast, coastal_buffer, fun = function(x,y){ifelse(is.na(x) & is.na(y), 999, x)})
#plot(fis_gf)

# convert remaining NA values to zero (assuming these are zero pressure)
fis_gf[is.na(fis_gf)] <- 0
#plot(fis_gf)

# now make land NA (don't want to gapfill using land 0 values)
fis_gf <- overlay(fis_gf, ocean_low_res, fun = function(x,y){ifelse(x==0 & is.na(y), NA, x)})
# plot(fis_gf)
# fis_gf

# convert 999 values created above into NA values that will be gapfilled
fis_gf[fis_gf==999] <- NA
# plot(fis_gf)
# fis_gf

## Function to gapfill NA cells using mean of surrounding cells
gf_raster <- function(x){focal(x, w = matrix(1,3,3), fun = mean, na.rm=TRUE, pad = TRUE, NAonly=TRUE)}

## Iterative gapfilling for each year of NPP data
## Repeat until critical NA cells are filled
i <- 0

while (i <= 50){
fis_gf <- gf_raster(fis_gf)
i <- i + 1
#print(i)
}

#plot(fis_gf)
writeRaster(fis_gf, file.path(dir_M, catch_path, file_name, paste0(file_name, "_", yr, "_gf", ".tif")), overwrite=TRUE)

}

```


## Map individual stressor layers

Each annual stressor layer is created by filtering the data for catch records that are consistent with what the stressor is measuring (each of the 5 fishing categories).

### Demersal destructive
```{r map,eval=F}

foreach(file = annual_files) %dopar%{ # file = annual_files[65]
  
  yr <- str_sub(file, -8, -5)
  
  data <- readRDS(file) %>%
      filter(destruction == 'destructive',
             fish_type == 'demersal') %>%
      rowwise() %>%
      mutate(catch = sum(LSF_CR, IUU_CR, Discards_CR, na.rm=TRUE)) %>%
                  group_by(Seq) %>%  #group by cell number
                  summarise(cell_catch = sum(catch)) %>%
    data.frame()#summing up all types of catch rates (remember these are in tonnes/km2 so we don't need to multiply by Ocean Area since we have 1km2 resolution)

  
  #rasterize catch by swapping cell ids with 

      raster::subs(saup_rast, data, by = 1, which = 2) %>%
        projectRaster(crs=crs(ocean), method="ngb", over = TRUE,
                      filename = file.path(dir_M, sprintf('git-annex/impact_acceleration/stressors/comm_fish/int/catch_annual_rasters/dem_dest/dem_dest_%s.tif', yr)), overwrite=TRUE)

}

## apply gapfill function to each of the above rasters
for(year_catch in 1950:2014){ # year = 1950
fish_gf(catch_path="git-annex/impact_acceleration/stressors/comm_fish/int/catch_annual_rasters",
        file_name = "dem_dest",
        yr = year_catch)
}

```


### Demersal non destructive high bycatch
```{r dem_nd_hb, eval=F}

foreach(file = annual_files) %dopar%{
  
  yr <- str_sub(file, -8, -5)
  
  data <- readRDS(file)%>%
      filter(destruction == 'non-destructive',
             fish_type == 'demersal',
             bycatch   == 'high')%>%
      rowwise() %>%
      mutate(catch = sum(LSF_CR, IUU_CR, Discards_CR, na.rm=TRUE)) %>%
                  group_by(Seq) %>%  #group by cell number
                  summarise(cell_catch = sum(catch))
  
     raster::subs(saup_rast, data, by = 1, which = 2) %>%
       projectRaster(crs=crs(ocean), method = "ngb", over=TRUE, 
                     filename = file.path(dir_M,  sprintf('git-annex/impact_acceleration/stressors/comm_fish/int/catch_annual_rasters/dem_nondest_hb/dem_nondest_hb_%s.tif', yr)), overwrite=TRUE)
}
     
## apply gapfill function to each of the above rasters
for(year_catch in 1950:2014){ # year = 1950
fish_gf(catch_path="git-annex/impact_acceleration/stressors/comm_fish/int/catch_annual_rasters",
        file_name = "dem_nondest_hb",
        yr = year_catch)
} 
     
```


### Demersal non destructive low bycatch
```{r dem_nd_lb, eval=F}

foreach(file = annual_files) %dopar%{
  
    yr <- str_sub(file, -8, -5)
  
  data <- readRDS(file) %>%
      filter(destruction == 'non-destructive',
             fish_type == 'demersal',
             bycatch   == 'low') %>%
      rowwise() %>%
      mutate(catch = sum(LSF_CR, IUU_CR, Discards_CR, na.rm=TRUE)) %>%
                  group_by(Seq) %>%  #group by cell number
                  summarise(cell_catch = sum(catch)) %>%
    data.frame()

     raster::subs(saup_rast, data, by = 1, which = 2) %>%
       projectRaster(crs=crs(ocean), method= "ngb", over = TRUE,
                    filename = file.path(dir_M, sprintf('git-annex/impact_acceleration/stressors/comm_fish/int/catch_annual_rasters/dem_nondest_lb/dem_nondest_lb_%s.tif', yr)), overwrite=TRUE)
}

## apply gapfill function to each of the above rasters
for(year_catch in 1950:2014){ # year = 1950
fish_gf(catch_path="git-annex/impact_acceleration/stressors/comm_fish/int/catch_annual_rasters",
        file_name = "dem_nondest_lb",
        yr = year_catch)
} 

```


### Pelagic low bycatch

```{r pel_lb,eval=F}

foreach(file = annual_files) %dopar%{
  
    yr <- str_sub(file, -8, -5)
  
  data <- readRDS(file) %>%
      filter(fish_type == 'pelagic',
             bycatch   == 'low') %>%
      rowwise() %>%
      mutate(catch = sum(LSF_CR, IUU_CR, Discards_CR, na.rm=TRUE)) %>%
                  group_by(Seq)%>%  #group by cell number
                  summarise(cell_catch = sum(catch)) %>%
    data.frame()
  
     raster::subs(saup_rast, data, by = 1, which = 2) %>%
       projectRaster(crs=crs(ocean), method = "ngb", over = TRUE,
                     filename =  file.path(dir_M, sprintf('git-annex/impact_acceleration/stressors/comm_fish/int/catch_annual_rasters/pel_lb/pel_lb_%s.tif', yr)), overwrite=TRUE)

}

## apply gapfill function to each of the above rasters
for(year_catch in 1950:2014){ # year = 1950
fish_gf(catch_path="git-annex/impact_acceleration/stressors/comm_fish/int/catch_annual_rasters",
        file_name = "pel_lb",
        yr = year_catch)
} 

```

### Pelagic high bycatch

```{r pel_hb, eval=F}

foreach(file = annual_files) %dopar%{
  
    yr <- str_sub(file, -8, -5)
  
  data <- readRDS(file) %>%
      filter(fish_type == 'pelagic',
             bycatch   == 'high') %>%
      rowwise() %>%
      mutate(catch = sum(LSF_CR, IUU_CR, Discards_CR, na.rm=TRUE)) %>%
                  group_by(Seq) %>%  #group by cell number
                  summarise(cell_catch = sum(catch)) %>%
    data.frame()
  
     raster::subs(saup_rast, data, by = 1, which = 2) %>%
       projectRaster(crs=crs(ocean), method = "ngb", over=TRUE, 
                     filename =  file.path(dir_M, sprintf('git-annex/impact_acceleration/stressors/comm_fish/int/catch_annual_rasters/pel_hb/pel_hb_%s.tif', yr)), overwrite=TRUE)

}

## apply gapfill function to each of the above rasters
for(year_catch in 1950:2014){ # year = 1950
fish_gf(catch_path="git-annex/impact_acceleration/stressors/comm_fish/int/catch_annual_rasters",
        file_name = "pel_hb",
        yr = year_catch)
} 

```

# Check results

If we add up the 5 layers it should equal the total catch for that year.

```{r}

#for 2014

#stack the 5 layers and add

r <- list.files(file.path(dir_M, 'git-annex/impact_acceleration/stressors/comm_fish/int/catch_annual_rasters'), pattern = '2014.tif', full.names = TRUE, recursive = TRUE) %>% 
  stack() %>% 
  sum(., na.rm = TRUE)


#all 0 cells set to NA mainly to remove land
r[r==0] <- NA

plot(r, col=cols)

#create single raster for 2014 from the data

data <- readRDS("/home/shares/ohi/git-annex/impact_acceleration/stressors/comm_fish/int/catch_annual_data/catch_data_2014.rds")%>%
      rowwise() %>%
      mutate(catch = sum(LSF_CR, IUU_CR, Discards_CR, na.rm=TRUE)) %>%
                  group_by(Seq) %>%  #group by cell number
                  summarise(cell_catch = sum(catch))
  
ras_2014 <-raster::subs(saup_rast, data, by = 1, which = 2, subsWithNA=TRUE) %>%
  projectRaster(r, over=TRUE, method="ngb")

plot(ras_2014, col=cols)

#check the difference

diff = ras_2014 - r
plot(diff, col=cols)
#click(diff)
ras_2014
r
## very small differences...looks good!
```
