---
title: "Creating fishing pressure layers"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ohi-northeast/src/templates/ohi_hdr.html'
pdf_document:
  toc: true
---

This script standardizes the annual catch rasters by net primary production.

```{r setup, message=F,warning=F}

knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',message = FALSE, eval=F, warning = FALSE)

source('~/github/impact_acceleration/src/R/common.R')

library(seaaroundus)
library(raster)
library(readr)
library(dplyr)
library(data.table)
library(parallel)
library(foreach)
library(doParallel)

```


# Primary production data

Load NPP data

```{r,eval=F}
npp_files <- list.files(file.path(dir_M,'marine_threats/impact_acceleration/stressors/comm_fish/VGPM_primary_productivity/int/annual_npp'),full.names=T)[1:12]

```

# Standardize catch

`catch_npp_fun` is a function that reads in a catch raster, reprojects & resamples it then divides by npp for the same year and saves the output.

```{r}

catch_npp_fun <- function(file,layer){
  
  catch <- raster(file)
  
  yr <- substr(file,nchar(file)-7,nchar(file)-4)
  
  npp <- npp_files[str_detect(npp_files,yr)]%>%raster()
  
  c <- catch%>%
        projectRaster(crs = mollCRS,over=T)%>%
        resample(npp,method = 'ngb')%>%
        overlay(.,npp,fun=function(x,y){x/y})
  
  writeRaster(c,filename = paste0(file.path(dir_M),'/marine_threats/impact_acceleration/stressors/comm_fish/output/annual_catch_npp/',layer,'/',layer,'_',yr,'.tif'))  
  
}

```

## Apply the function
```{r, eval=F}

dem_d_files <- list.files(file.path(dir_M,'marine_threats/impact_acceleration/stressors/comm_fish/output/annual_catch/dem_dest'),full.names=T)[54:65]
dem_nd_hb    <-list.files(file.path(dir_M,'marine_threats/impact_acceleration/stressors/comm_fish/output/annual_catch/dem_nondest_hb'),full.names=T)[54:65]
dem_nd_lb    <-list.files(file.path(dir_M,'marine_threats/impact_acceleration/stressors/comm_fish/output/annual_catch/dem_nondest_lb'),full.names=T)[54:65]
pel_hb    <-list.files(file.path(dir_M,'marine_threats/impact_acceleration/stressors/comm_fish/output/annual_catch/pel_hb'),full.names=T)[54:65]
pel_lb    <-list.files(file.path(dir_M,'marine_threats/impact_acceleration/stressors/comm_fish/output/annual_catch/pel_lb'),full.names=T)[54:65]


lapply(dem_d_files,catch_npp_fun,layer = 'dem_dest')
lapply(dem_nd_hb,catch_npp_fun,layer = 'dem_nondest_hb')
lapply(dem_nd_lb,catch_npp_fun,layer = 'dem_nondest_lb')
lapply(pel_hb,catch_npp_fun,layer = 'pel_hb')
lapply(pel_lb,catch_npp_fun,layer = 'pel_lb')
```


# Rescale

Melanie's suggestion:

> For the reference point, it would be nice to link the fishing pressure to biological/ecological outcomes, but I don't know if these data are available. This fits in well with the nonlinearities project, and it might be worth discussing/researching this topic (I will add this as an issue for the nonlinearity project). But in the meantime, I think we should probably go with our standard approach of using the 99.99th max value, even though it is not ideal. I propose the following method (we should discuss this):

> - average 5 earliest years of data (2003-2007)  
> - from the 5 years average data: determine the value corresponding to the 99.99th quantile  
> - use this value as the reference point across all years  

```{r,eval=F}

#list of lists
all <- list.files(file.path(dir_M,'marine_threats/impact_acceleration/stressors/comm_fish/output/annual_catch_npp'),full.names=T,recursive = T)

layers <- c("dem_dest","dem_nondest_hb","dem_nondest_lb","pel_hb","pel_lb")

registerDoParallel(5)

foreach(layer = layers) %dopar% {    
  
  files = all[which(str_detect(all,as.character(layer)))]

  #average first five years
  first_5 <- files[1:5]%>%
            stack()%>%
            calc(.,fun=function(x){mean(x,na.rm=T)})
  
  #get 99.99th quantile
  quant <- getValues(files[1:5]%>%
            stack())%>%
          quantile(.,prob = 0.9999,na.rm=T)
  
  #rescale each year according to this reference point and save output
  for (i in 1:12){
    
    yr = substr(files[i],nchar(files[i])-7,nchar(files[i])-4)
    
          files[i]%>%
          raster()%>%
          calc(fun=function(x){ifelse(x<0,0,
                                      ifelse(x>quant,1,x/quant))})%>%
          resample(ocean,method = 'ngb')%>%
                     writeRaster(filename = paste0(file.path(dir_M),'/marine_threats/impact_acceleration/stressors/comm_fish/output/annual_catch_npp_rescale/', layer,'/',layer,'_',yr,'.tif'),format="GTiff", overwrite=TRUE)
  }
}
```
